### 2024-11-26 paper/paper.qmd
repo set up for submitting

---
title: "Where are Torontonians likely to get married"
subtitle: "Analysing the different marriage offices around GTA"
author: 
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/aamishi/marriages_weekly_reflection/tree/master"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
# bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(arrow)
library(knitr)

data <- read_parquet("../data/analysis_data/analysis_data.parquet")
```

# Introduction

In this week's reflection exercise, I will take a look at where people from GTA are most likely to register their marriage. I will be using @citeR to simulate, test and analyse this data. I will download the data from @opendatatoronto.

# Data

### point 1: general table to show a glimpse of the data
```{r}
#| label: fig-general-data-todo
#| fig-cap: General table TODO
#| echo: false

kable(data, caption = "Table: Sample Measurements of Subjects")
```

Talk more about it.

```{r}

```

Talk way more about it.

\newpage

# References

### 2024-11-26 scripts/01-download_data.R

@@ -7,26 +7,16 @@
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
 # [...UPDATE THIS...]
 
 #### Download data ####
-# get package
-package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# get all resources for this package
-resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
-  get_resource()
-
+# get package ; DATASET FOR NEW BODY FAT MEASUREMENT
+#### Clean data ####
+measurements_data <- read_csv("./data/raw_data/local-train/measurements.csv")
+subject_sex_data <- read_csv("./data/raw_data/local-train/hwg_metadata.csv")
 
 #### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
-
+write_parquet(measurements_data, "data/raw_data/measurements_data.parquet")
+write_parquet(subject_sex_data, "data/raw_data/subject_sex_data.parquet")
          

### 2024-11-26 scripts/02-data_cleaning.R

@@ -7,36 +7,33 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+#### Read data ####
+subject_sex_data <- read_parquet("data/raw_data/subject_sex_data.parquet")
+measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+combined_data <- merge(subject_sex_data, measurements_data, by = "subject_id")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+cleaned_data <- combined_data %>%
+  clean_names() %>%
+  mutate(waist_hip_ratio = waist / hip) %>%
+  mutate(original_bmi = weight_kg / (height_cm / 100)^2) %>%
+  mutate(subject_id = seq(1, nrow(combined_data)))
 
+cleaned_data$gender <- as.numeric(factor(combined_data$gender, levels = c("male", "female"))) - 1
+
+analysis_data <- cleaned_data %>%
+  select(subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi)
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_parquet(analysis_data, "data/analysis_data/analysis_data.parquet")
+
+
+
+
+
+
+
+

### 2024-11-26 scripts/04-model.R

@@ -13,7 +13,7 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/analysis_data/analysis_data.parquet")
 
 ### Model data ####
 first_model <-

### 2024-11-26 scripts/99-data.R

@@ -1,0 +1,206 @@
+#### Preamble ####
+# Purpose: Cleans the Marriage License data
+# Author: Aamishi Avarsekar
+# Date: 19th September 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+#### Read data ####
+analysis_data <- read_parquet("data/analysis_data/analysis_data.parquet")
+
+#### Commuicate a good sense of the data:
+# My idea:
+# 1. show the difference in gender and the bmi levels
+# 2. the ratios between waist-hip and bmi levels
+# 3. how wrist and ankle measurements could reflect ...
+# and the exclusively compare this to the MM-FM-FFM data
+
+# 1. general table
+kable(analysis_data, caption = "Table: Sample Measurements of Subjects")
+
+# 2. bmi vs gender
+# Create BMI bins with intervals of 5
+bmi_interval <- 5
+analysis_data <- analysis_data %>%
+  mutate(bmi_bin = cut(original_bmi, breaks = seq(0, 50, by = bmi_interval), right = FALSE, 
+                       labels = paste(seq(0, 45, by = bmi_interval), seq(5, 50, by = bmi_interval), sep = "-")))
+
+ggplot(analysis_data, aes(x = bmi_bin, fill = factor(gender))) +
+  geom_bar(position = "dodge", color = "black") +
+  labs(title = "BMI Distribution by Gender", x = "BMI Range", y = "Count") +
+  theme_minimal() +
+  scale_fill_manual(values = c("0" = "blue", "1" = "pink"))
+
+analysis_data %>%
+  group_by(bmi_bin, gender) %>%
+  summarise(count = n()) %>%
+  group_by(bmi_bin) %>%
+  mutate(prop = count / sum(count)) %>%
+  ggplot(aes(x = bmi_bin, y = prop, fill = factor(gender))) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Proportion of Gender within BMI Range", x = "BMI Range", y = "Proportion") +
+  scale_fill_manual(values = c("0" = "blue", "1" = "pink")) +
+  theme_minimal()
+
+
+ggplot(analysis_data, aes(x = factor(gender), y = original_bmi)) +
[…]

### 2024-11-26 paper/paper.qmd
peer reviews

@@
 ---
-title: "Where are Torontonians likely to get married"
-subtitle: "Analysing the different marriage offices around GTA"
+title: "How General Health Index (GHI) performs better to predict health levels"
+subtitle: "A comparitive analysis of GHI and BMI against body measurements"
 author: 
   - Aamishi Avarsekar
-thanks: "Code and data are available at: https://github.com/aamishi/marriages_weekly_reflection/tree/master"
+thanks: "Code and data are available at: https://github.com/aamishi/ImprovedGeneralHealthIndex/"
 date: today
@@
 number-sections: true
-# bibliography: references.bib
+bibliography: references.bib
 ---
@@
 
-library(tidyverse)
-library(dplyr)
-library(arrow)
-library(knitr)
-
-data <- read_parquet("../data/analysis_data/analysis_data.parquet")
 ```
@@
 
-In this week's reflection exercise, I will take a look at where people from GTA are most likely to register their marriage.
-I will be using @citeR to simulate, test and analyse this data.
-I will download the data from @opendatatoronto.
+BMI is a common indicator for general health but it has its limitations.
+It does not take into consideration the actual body composition of an individual and other important factors such as sex and age.
+Through this paper, I would like to propose a new way categorising people based on other body measurements such as the waist-to-hip ratio and wrist and ankle circumferences measurements as they are more indicative of a person's visceral fat, this proving a more accurate representation of general health.
+I would like to propose a newer formula for General Health Index (GHI) through the BodyM dataset.
+I plan on using the Isaac Kuzmar dataset that has accurate measurements of body fat to create a Confusion Matrix to see how my model compares in providing more accurate body fat estimations through easlily measurable data such as waist and hip circumference.
 
-# Data
+# Data {#sec-data}
 
-### point 1: general table to show a glimpse of the data
+## Overview
+
+Overview text
+
+## Measurement
+
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+
 ```{r}
-#| label: fig-general-data-todo
-#| fig-cap: General table TODO
+#| label: fig-bills
+#| fig-cap: Bills of penguins
 #| echo: false
-
-kable(data, caption = "Table: Sample Measurements of Subjects")
 ```
@@
 
+## Predictor variables
+
+Add graphs, tables and text.
+
+Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+# Model
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+
+```
+
+# Discussion
+
+## First discussion point {#sec-first-point}
+
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+
+## Second discussion point
+
+Please don't use these as sub-heading labels - change them to be what your point actually is.
+
+## Third discussion point
+
+## Weaknesses and next steps
+
+Weaknesses and next steps should also be included.
+
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+# Additional data details
+
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+```
+
 \newpage

### 2024-11-26 scripts/01-download_data.R

@@ -8,15 +8,24 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 # [...UPDATE THIS...]
 
 #### Download data ####
 # get package ; DATASET FOR NEW BODY FAT MEASUREMENT
 #### Clean data ####
 measurements_data <- read_csv("./data/raw_data/local-train/measurements.csv")
-subject_sex_data <- read_csv("./data/raw_data/local-train/hwg_metadata.csv")
+measurements_subject_data <- read_csv("./data/raw_data/local-train/hwg_metadata.csv")
+
+measurements_data_A <- read_csv("./data/raw_data/local-testA/measurements.csv")
+measurements_subject_data_A <- read_csv("./data/raw_data/local-testA/hwg_metadata.csv")
+body_mass_data <- read_csv2("./data/raw_data/DATASET FOR NEW BODY FAT MEASUREMENT.csv")
 
 #### Save data ####
 write_parquet(measurements_data, "data/raw_data/measurements_data.parquet")
-write_parquet(subject_sex_data, "data/raw_data/subject_sex_data.parquet")
+write_parquet(measurements_subject_data, "data/raw_data/measurements_subject_data.parquet")
+
+write_parquet(measurements_data_A, "data/raw_data/measurements_data_A.parquet")
+write_parquet(measurements_subject_data_A, "data/raw_data/measurements_subject_data_A.parquet")
+write_parquet(body_mass_data, "data/raw_data/body_mass_data.parquet")
          

### 2024-11-26 scripts/02-data_cleaning.R

@@ -11,11 +11,20 @@ library(janitor)
 library(arrow)
 library(dplyr)
 #### Read data ####
-subject_sex_data <- read_parquet("data/raw_data/subject_sex_data.parquet")
+# 1. MEASUREMENTS DATA (AWS - BodyM)
+measurements_subject_data <- read_parquet("data/raw_data/measurements_subject_data.parquet")
 measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
 
+measurements_subject_data_A <- read_parquet("data/raw_data/measurements_subject_data_A.parquet")
+measurements_data_A <- read_parquet("data/raw_data/measurements_data_A.parquet")
+
+# 2. BODY MASS DATA
+body_mass_data <- read_parquet("data/raw_data/body_mass_data.parquet")
+
 #### Clean data ####
-combined_data <- merge(subject_sex_data, measurements_data, by = "subject_id")
+
+# 1. MEASUREMENTS DATA (AWS - BodyM)
+combined_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
 
 cleaned_data <- combined_data %>%
   clean_names() %>%
@@ -27,8 +36,52 @@ cleaned_data$gender <- as.numeric(factor(combined_data$gender, levels = c("male"
 
 analysis_data <- cleaned_data %>%
   select(subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi)
+
+analysis_data$category <- rep(NA, nrow(analysis_data))  # Create a placeholder column with 2018 rows
+analysis_data$category <- factor(analysis_data$category, levels = c(
+  "Extremely Underweight", "Underweight", "Normal", "Overweight", "Extremely Overweight"
+))
+analysis_data$category <- ifelse(analysis_data$original_bmi < 18.5, "Underweight",
+                          ifelse(analysis_data$original_bmi < 25, "Normal",
+                          ifelse(analysis_data$original_bmi < 30, "Overweight", 
+                          ifelse(analysis_data$original_bmi < 30, "Obese I",
+                          ifelse(analysis_data$original_bmi < 30, "Obese II", "Obese III")))))
+
+# TEST A
+combined_data_A <- merge(measurements_subject_data_A, measurements_data_A, by = "subject_id")
+
+cleaned_data_A <- combined_data_A %>%
+  clean_names() %>%
+  mutate(waist_hip_ratio = waist / hip) %>%
[…]

### 2024-11-26 scripts/04-model.R

@@ -11,21 +11,21 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+### Model data ####
+library(nnet)
 
 #### Read data ####
 analysis_data <- read_parquet("data/analysis_data/analysis_data.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Multinomial logistic regression model
+# subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi
+multi_model <- multinom(category ~ weight_kg + height_cm + waist_hip_ratio + wrist + ankle, data = analysis_data)
+# Summary of the model
+summary(multi_model)
+
+
+analysis_data$predicted_category <- predict(multi_model, newdata = analysis_data)
+table(Predicted = analysis_data$predicted_category, Actual = body_mass_clean$category)
 
 
 #### Save model ####

### 2024-11-29 scripts/01-download_data.R
Data Download, +

@@ -6,26 +6,31 @@
 # License: MIT
 
 
-#### Workspace setup ####
+# Workspace setup
 library(tidyverse)
 library(arrow)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# get package ; DATASET FOR NEW BODY FAT MEASUREMENT
-#### Clean data ####
+# Download data
+# Model 1 Data: AWS BodyM Training Data
 measurements_data <- read_csv("./data/raw_data/local-train/measurements.csv")
 measurements_subject_data <- read_csv("./data/raw_data/local-train/hwg_metadata.csv")
 
-measurements_data_A <- read_csv("./data/raw_data/local-testA/measurements.csv")
-measurements_subject_data_A <- read_csv("./data/raw_data/local-testA/hwg_metadata.csv")
+# Model 2 Data: Isaac Kuzmar Et Al Data - Body Composition Data
+# Data uses ; as delimiter
 body_mass_data <- read_csv2("./data/raw_data/DATASET FOR NEW BODY FAT MEASUREMENT.csv")
 
+# Testing Data: AWS BodyM Test A Data
+test_data <- read_csv("./data/raw_data/local-testA/measurements.csv")
+test_subject_data <- read_csv("./data/raw_data/local-testA/hwg_metadata.csv")
+
 #### Save data ####
+# Model 1
 write_parquet(measurements_data, "data/raw_data/measurements_data.parquet")
 write_parquet(measurements_subject_data, "data/raw_data/measurements_subject_data.parquet")
 
-write_parquet(measurements_data_A, "data/raw_data/measurements_data_A.parquet")
-write_parquet(measurements_subject_data_A, "data/raw_data/measurements_subject_data_A.parquet")
+# Model 2
 write_parquet(body_mass_data, "data/raw_data/body_mass_data.parquet")
-         
+
+# Test
+write_parquet(test_data, "data/raw_data/test_data.parquet")
+write_parquet(test_subject_data, "data/raw_data/test_subject_data.parquet")       

### 2024-11-30 paper/paper.qmd
maybe working model

@@
 
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
 ```
@@
 ## Overview
+
+## BMI Classification:
+```{r}
+#| warning: false
+#| message: false
+bmi_canada_gov <- data.frame(
+  Category = c("Underweight", "Normal weight", "Overweight", "Obese - Class I", "Obese - Class II", "Obese - Class III"),
+  `BMI Range` = c("Below 18.5", "18.5–24.9", "25.0–29.9", "30.0–34.9", "35.0–39.9", "40.0 and above")
+)
+
+# Create a table using knitr::kable
+knitr::kable(
+  bmi_canada_gov,
+  col.names = c("Category", "BMI Range"),
+  caption = "BMI Categories as per the Government of Canada Guidelines"
+)
+```
+### The difference between an underweight and normal person:
+Let us take a closer look at the BMI levels below 25 for underweight and normal weight adults.
+
+```{r}
+#| warning: false
+#| message: false
+body_mass_data_raw <- read_parquet("../data/raw_data/body_mass_data.parquet")
+body_mass_clean <- body_mass_data_raw %>%
+  clean_names() %>%
+  select(gender_m1f2, age, height_cm, w_kg, bmi, fat_mass_kg, fm, ffm_kg, bone_mass_kg, muscle_mass_kg) %>%
+  mutate(
+    bmi_category = case_when(
+      bmi < 18.5 ~ "Underweight",
+      bmi >= 18.5 & bmi <= 24.9 ~ "Normal weight",
+      bmi >= 25.0 & bmi <= 29.9 ~ "Overweight",
+      bmi >= 30.0 & bmi <= 34.9 ~ "Obese - Class I",
+      bmi >= 35.0 & bmi <= 39.9 ~ "Obese - Class II",
+      bmi >= 40.0 ~ "Obese - Class III"
+    ))
+
+# Data Section: Part 1: What is the difference between an underweight / athletic person / normal person:
+bmi_under_25 <- body_mass_clean %>%
+  filter(bmi < 25)
+
+ggplot(bmi_under_25, aes(x = bmi, y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) +
+  labs(
+    title = "BMI vs. Fat Mass by BMI Category",
+    x = "BMI",
+    y = "Fat Mass (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+
+```
+
+### The difference between a normal and overweight person:
+```{r}
+bmi_normal_over <- body_mass_clean %>%
+  filter(18.5 < bmi & bmi < 30)
+
+# this is what i chose
+ggplot(bmi_normal_over, aes(x = bmi, y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) +
+  labs(
+    title = "BMI vs. Fat Mass by BMI Category",
+    x = "BMI",
+    y = "Fat Mass (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+```
+
+### All categories
+```{r}
+# this is what i chose
+ggplot(body_mass_clean, aes(x = bmi, y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) +
+  geom_vline(xintercept = c(18.5, 25, 30), linetype = "dashed", color = "black", linewidth = 0.5, alpha = 0.5) +
+  labs(
+    title = "BMI vs. Fat Mass by BMI Category",
+    x = "BMI",
+    y = "Fat Mass (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+```
 

### 2024-11-30 scripts/02-data_cleaning.R

@@ -4,22 +4,18 @@
 # Date: 19th September 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
+# Note: Aspects of this code was written with the aid of ChatGPT 4.0
 
-#### Workspace setup ####
+# Work space setup
 library(tidyverse)
 library(janitor)
 library(arrow)
 library(dplyr)
-#### Read data ####
+
+#Read data
 # 1. MEASUREMENTS DATA (AWS - BodyM)
 measurements_subject_data <- read_parquet("data/raw_data/measurements_subject_data.parquet")
 measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
-
-measurements_subject_data_A <- read_parquet("data/raw_data/measurements_subject_data_A.parquet")
-measurements_data_A <- read_parquet("data/raw_data/measurements_data_A.parquet")
-
-# 2. BODY MASS DATA
-body_mass_data <- read_parquet("data/raw_data/body_mass_data.parquet")
 
 #### Clean data ####
 
@@ -29,59 +25,33 @@ combined_data <- merge(measurements_subject_data, measurements_data, by = "subje
 cleaned_data <- combined_data %>%
   clean_names() %>%
   mutate(waist_hip_ratio = waist / hip) %>%
+  mutate(height_hip_ratio = height_cm / hip) %>%
   mutate(original_bmi = weight_kg / (height_cm / 100)^2) %>%
-  mutate(subject_id = seq(1, nrow(combined_data)))
+  mutate(subject_id = seq(1, nrow(combined_data))) %>%
+  select(-height) %>%
+  rename(
+    height = height_cm,  
+    weight = weight_kg,           
+    bmi = original_bmi                
+  )
 
-cleaned_data$gender <- as.numeric(factor(combined_data$gender, levels = c("male", "female"))) - 1
+cleaned_data$gender <- as.numeric(factor(combined_data$gender, levels = c("male", "female"))) 
 
-analysis_data <- cleaned_data %>%
-  select(subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi)
+# Written with the help of ChatGPT with modifications
+# Create a placeholder column with 2018 rows
+cleaned_data$bmi_category <- factor(NA, levels = c("Underweight", "Normal", "Overweight", "Obese"))
 
[…]

### 2024-11-30 scripts/02b-data_cleaning.R

@@ -1,0 +1,38 @@
+#### Preamble ####
+# Purpose: TODO
+# Author: Aamishi Avarsekar
+# Date: TODO
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+
+#### Read Data ####
+# 2. BODY MASS DATA
+body_mass_data_raw <- read_parquet("data/raw_data/body_mass_data.parquet")
+
+#### Clean Data ####
+body_mass_clean <- body_mass_data_raw %>%
+  clean_names() %>%
+  select(gender_m1f2, age, height_cm, w_kg, bmi, fat_mass_kg, fm, ffm_kg, bone_mass_kg, muscle_mass_kg, bmr_kcal, fm_trunk) %>%
+  mutate(
+    bmi_category = case_when(
+      bmi < 18.5 ~ "Underweight",
+      bmi >= 18.5 & bmi <= 24.9 ~ "Normal",
+      bmi >= 25.0 & bmi <= 29.9 ~ "Overweight",
+      bmi >= 30.0 ~ "Obese"
+    )) %>%
+  rename(
+    height = height_cm,  
+    weight = w_kg,           
+    gender = gender_m1f2                
+  )
+
+#body_mass_under_25 <-
+
+#### Save data ####
+write_parquet(body_mass_clean, "data/analysis_data/body_mass_data.parquet")

### 2024-11-30 scripts/02c-data_cleaning.R

@@ -1,0 +1,70 @@
+#### Preamble ####
+# Purpose: Cleans the Marriage License data
+# Author: Aamishi Avarsekar
+# Date: 19th September 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+# Note: Aspects of this code was written with the aid of ChatGPT 4.0
+
+# Work space setup
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+
+#Read data
+# 1. MEASUREMENTS DATA (AWS - BodyM)
+measurements_subject_data <- read_parquet("data/raw_data/test_subject_data.parquet")
+measurements_data <- read_parquet("data/raw_data/test_data.parquet")
+
+#### Clean data ####
+
+# 1. MEASUREMENTS DATA (AWS - BodyM)
+combined_test_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
+
+cleaned_test_data <- combined_test_data %>%
+  clean_names() %>%
+  mutate(waist_hip_ratio = waist / hip) %>%
+  mutate(original_bmi = weight_kg / (height_cm / 100)^2) %>%
+  mutate(subject_id = seq(1, nrow(combined_test_data))) %>%
+  select(-height) %>%
+  rename(
+    height = height_cm,  
+    weight = weight_kg,           
+    bmi = original_bmi                
+  )
+
+cleaned_test_data$gender <- as.numeric(factor(combined_test_data$gender, levels = c("male", "female"))) - 1
+
+# Written with the help of ChatGPT with modifications
+# Create a placeholder column with 2018 rows
+cleaned_test_data$bmi_category <- factor(NA, levels = c("Underweight", "Normal", "Overweight", "Obese"))
+
+# Assign category based on original_bmi values
+cleaned_test_data$bmi_category <- ifelse(cleaned_test_data$bmi < 18.5, "Underweight",
+                                    ifelse(cleaned_test_data$bmi < 25, "Normal",
+                                           ifelse(cleaned_test_data$bmi < 30, "Overweight", "Obese")))
+
+# Ensure 'category' is a factor with the correct levels
+cleaned_test_data$bmi_category <- factor(cleaned_test_data$bmi_category, levels = c("Underweight", "Normal", "Overweight", "Obese"))
+
+# Changing gender: 1 = male, 2 = female
+# Selecting only: subject_id, gender, height, weight, bmi, bmi_category
[…]

### 2024-11-30 scripts/04-model.R

@@ -8,30 +8,59 @@
 # Any other information needed? [...UPDATE THIS...]
 
 
-#### Workspace setup ####
+# Works pace setup - packages needed
 library(tidyverse)
 library(rstanarm)
-### Model data ####
 library(nnet)
+library(randomForest)
 
 #### Read data ####
-analysis_data <- read_parquet("data/analysis_data/analysis_data.parquet")
+cleaned_data <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
+body_mass_read <- read_parquet("data/analysis_data/body_mass_data.parquet")
 
 # Multinomial logistic regression model
 # subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi
-multi_model <- multinom(category ~ weight_kg + height_cm + waist_hip_ratio + wrist + ankle, data = analysis_data)
+# Variables selected:
+measurements_model <- multinom(bmi_category ~ height + gender + waist_hip_ratio + height_hip_ratio, data = cleaned_data)
 # Summary of the model
-summary(multi_model)
+summary(measurements_model)
 
 
-analysis_data$predicted_category <- predict(multi_model, newdata = analysis_data)
-table(Predicted = analysis_data$predicted_category, Actual = body_mass_clean$category)
+#### Read data ####
+# body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
+body_mass_model_bmi <- multinom(bmi_category ~ gender + fm + bmr_kcal + muscle_mass_kg + fm_trunk, data = body_mass_read)
+body_mass_read$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_read)
+
+
+body_mass_model <- randomForest(fm ~ height + weight + gender * predicted_category + muscle_mass_kg, data = body_mass_read)
+
+
+# model 2 - final work being done here
+body_mass_read$predicted_fm <- predict(body_mass_model, newdata = body_mass_read)
+plot(body_mass_read$fm, body_mass_read$predicted_fm,
+     xlab = "Actual Fat Percentage",
+     ylab = "Estimated Fat Percentage",
+     main = "Estimated vs Actual Fat Percentage",
+     pch = 16, col = "blue")
+abline(a = 0, b = 1, col = "red", lty = 2) # Add a 1:1 reference line
+
 
 
 #### Save model ####
 saveRDS(
-  first_model,
[…]

### 2024-11-30 scripts/04b-model.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+### Model data ####
+library(nnet)
+
+#### Read data ####
+body_mass_read <- read_parquet("data/analysis_data/body_mass_data.parquet")
+body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
+
+library(randomForest)
+body_mass_model <- randomForest(fm ~ weight  + bmi_category, data = body_mass_read)
+
+body_mass_model_bmi <- multinom(bmi_category ~ weight + fm , data = body_mass_read)
+
+
+
+#### Save model ####
+saveRDS(
+  body_mass_model,
+  file = "models/body_mass_model.rds"
+)
+
+saveRDS(
+  body_mass_model_bmi,
+  file = "models/body_mass_model_bmi.rds"
+)
+
+

### 2024-11-30 scripts/04c-model.R

@@ -1,0 +1,44 @@
+plot(body_mass_read$fm, body_mass_read$predicted_fm,
+     xlab = "Actual Fat Percentage",
+     ylab = "Estimated Fat Percentage",
+     main = "Estimated vs Actual Fat Percentage",
+     pch = 16, col = "blue")
+abline(a = 0, b = 1, col = "red", lty = 2) +
+  facet_wrap(~ bmi_category + gender)
+
+
+
+ggplot(body_mass_read) +
+  geom_density(aes(x = fm, fill = "Actual"), alpha = 0.5) +
+  geom_density(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5) +
+  labs(x = "Fat Percentage", y = "Density",
+       title = "Distribution of Actual vs Estimated Fat Percentage") +
+  scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
+  theme_minimal() +
+  facet_wrap(~ bmi_category)
+
+ggplot(body_mass_read) +
+  geom_histogram(aes(x = fm, fill = "Actual"), alpha = 0.5, position = "identity", bins = 30) +
+  geom_histogram(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5, position = "identity", bins = 30) +
+  labs(x = "Fat Percentage", y = "Count",
+       title = "Distribution of Actual vs Estimated Fat Percentage") +
+  scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
+  theme_minimal() +
+  facet_wrap(~ bmi_category)
+
+ggplot(body_mass_read) +
+  geom_boxplot(aes(x = bmi_category, y = fm, fill = "Actual"), alpha = 0.5, position = "dodge") +
+  geom_boxplot(aes(x = bmi_category, y = predicted_fm, fill = "Estimated"), alpha = 0.5, position = "dodge") +
+  labs(x = "BMI Category", y = "Fat Percentage",
+       title = "Boxplot of Actual vs Estimated Fat Percentage") +
+  scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
+  theme_minimal()
+
+ggplot(body_mass_read) +
+  geom_violin(aes(x = bmi_category, y = fm, fill = "Actual"), alpha = 0.5) +
+  geom_violin(aes(x = bmi_category, y = predicted_fm, fill = "Estimated"), alpha = 0.5) +
[…]

### 2024-11-30 scripts/101-model_1_EDA.R

@@ -1,0 +1,77 @@
+# Work space setup
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+# before graphs []
+ggplot(cleaned_data, aes(x = waist_hip_ratio, color = bmi_category, fill = bmi_category)) +
+  geom_density(alpha = 0.3) + 
+  facet_wrap(~ gender) +
+  labs(
+    title = "Overlapping Density Plot for Fat Mass by BMI OG",
+    x = "waist_hip_ratio",
+    y = "Density",
+    color = "BMI Category",
+    fill = "BMI Category"
+  ) +
+  theme_minimal()
+
+# analysis_data
+ggplot(cleaned_data, aes(x = waist_hip_ratio, color = predicted_category, fill = predicted_category)) +
+  geom_density(alpha = 0.3) + 
+  facet_wrap(~ gender) +
+  labs(
+    title = "Overlapping Density Plot for Fat Mass by predicted_category Category 4",
+    x = "waist_hip_ratio",
+    y = "Density",
+    color = "BMI Category",
+    fill = "BMI Category"
+  ) +
+  theme_minimal()
+
+library(ggplot2)
+
+ggplot(cleaned_data, aes(x = bmi, fill = interaction(bmi_category, predicted_category))) + 
+  geom_density(alpha = 0.5) + 
+  labs(title = "Comparison of Original and Improved BMI Distributions", 
+       x = "BMI Value", 
+       y = "Density") 
+
+df_long <- cleaned_data %>%
+  pivot_longer(cols = c(bmi_category, predicted_category), 
+               names_to = "category_type", 
+               values_to = "category")
+
+# Now you can plot the densities
+ggplot(df_long, aes(x = bmi, fill = category_type)) + 
+  geom_density(alpha = 0.5) + 
+  facet_wrap(~ category, scales = "free") + 
+  labs(title = "Distribution of Original vs Improved BMI by Category", 
+       x = "BMI Value", 
+       y = "Density") +
+  scale_fill_manual(values = c("blue", "green")) 
+
+ggplot(df_long, aes(x = weight_kg, y = height_cm)) +
+  geom_point(aes(color = category), alpha = 0.7) +
+  labs(
+    title = "Scatterplot with 2D Density Contours",
+    x = "Fat Mass (kg)",
+    y = "Weight (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+
[…]

### 2024-11-30 scripts/102-model_2_EDA.R

@@ -1,0 +1,83 @@
+# Work space setup
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+
+
+# Data Section: Part 1: What is the difference between an underweight / athletic person / normal person:
+bmi_under_25 <- body_mass_clean %>%
+  filter(bmi < 25)
+
+# this is what i chose
+ggplot(bmi_under_25, aes(x = , y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) 
+labs(
+  title = "BMI vs. Fat Mass by BMI Category",
+  x = "BMI",
+  y = "Fat Mass (kg)",
+  color = "BMI Category"
+) +
+  theme_minimal()
+
+
+# Data Section: Part 2: What is the difference between an athletic person / normal person / overweight person:
+bmi_normal_over <- body_mass_clean %>%
+  filter(18.5 < bmi & bmi < 30)
+
+# this is what i chose
+ggplot(bmi_normal_over, aes(x = bmi, y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) +
+  facet_wrap(~ bmi_category) +
+  labs(
+    title = "BMI vs. Fat Mass by BMI Category",
+    x = "BMI",
+    y = "Fat Mass (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+
+
+# all categories
+bmi_normal_over <- body_mass_clean %>%
+  filter(18.5 < bmi & bmi < 30)
+
+# this is what i chose
+ggplot(body_mass_clean, aes(x = bmi, y = fm)) +
+  geom_point(aes(color = bmi_category), alpha = 0.7) +
+  labs(
+    title = "BMI vs. Fat Mass by BMI Category",
+    x = "BMI",
+    y = "Fat Mass (kg)",
+    color = "BMI Category"
+  ) +
+  theme_minimal()
+
+
+#### FAT MASS VISUALS
+# all categories
+# i like
+# could add - mean vlines for each category
+ggplot(body_mass_clean, aes(x = fat_mass_kg, color = bmi_category, fill = bmi_category)) +
+  geom_density(alpha = 0.3) + 
+  facet_wrap(~ gender) +
+  labs(
+    title = "Overlapping Density Plot for Fat Mass by BMI Category",
+    x = "Fat Mass (kg)",
+    y = "Density",
+    color = "BMI Category",
+    fill = "BMI Category"
[…]

### 2024-11-30 scripts/103-test_data_EDA.R

@@ -1,0 +1,228 @@
+# Work space setup
+library(tidyverse)
+library(janitor)
+library(arrow)
+library(dplyr)
+install.packages("caret")
+
+library(caret)
+
+
+
+
+
+
+
+# Import data:
+# Model 1 Data
+measurements_data_predictions <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
+# Model 2 Data
+body_mass_data_predictions <- read_parquet("data/analysis_data/body_mass_data.parquet")
+
+# Test Data: 
+cleaned_test_data_predictions <- read_parquet("data/analysis_data/test_analysis.parquet")
+
+# Model 1 predictions:
+measurements_data_predictions$predicted_category <- predict(measurements_model, newdata = measurements_data_predictions)
+#table(Predicted = TODO$predicted_category, Actual = cleaned_data$bmi_category)
+
+
+
+
+# Model 2 predictions:
+#table(Predicted = measurements_data_predictions$model_1_predicted_categories, Actual = measurements_data_predictions$bmi_category)
+#table(Predicted = body_mass_data_predictions$model_2_predicted_categories, Actual = body_mass_data_predictions$bmi_category)
+body_mass_data_predictions$predicted_category <- predict(body_mass_model, newdata = body_mass_data_predictions)
+
+# Extract Categories:
+measurements_data_extracted <- measurements_data_predictions %>%
+  select(height, weight, bmi, bmi_category, predicted_category)
+
+body_mass_data_extracted <- body_mass_data_predictions %>%
+  select(height, weight, bmi, bmi_category, predicted_category)
+
+# visual to see results of the models
+# model 1:
+model_1_long <- measurements_data_extracted %>%
+  pivot_longer(cols = c(bmi_category, model_1_predicted_categories), 
+               names_to = "category_type", 
+               values_to = "category")
+
+# Now you can plot the densities
+ggplot(model_1_long, aes(x = bmi, fill = category_type)) + 
+  geom_density(alpha = 0.5) + 
+  facet_wrap(~ category, scales = "free") + 
+  labs(title = "Distribution of Original vs Improved BMI by Category", 
+       x = "BMI Value", 
+       y = "Density") +
+  scale_fill_manual(values = c("blue", "green")) 
+
+# model 2
+model_2_long <- body_mass_data_extracted %>%
+  pivot_longer(cols = c(bmi_category, model_2_predicted_categories), 
[…]

### 2024-11-30 paper/paper.qmd
commit to preserve ideas & clean up followed

@@
 
-# Data {#sec-data}
-
-## Overview
+- this paper uses 2 datasets to capture the difference in fat assessment for popular methods.
+BMI is a popular way of estimating health issues in an individual but it fails to capture details like body composition.
+In its initial design, BMI was designed using only male [] subjects and thus cannot be applied to females.
+However, due to excess popularity, it is used as a self-assessment to estimate your body composition and health risks.
+However, due to its inability to capture body composition, a lot of false positives and false negatives are produced.
+These indivuals often identify health risks [] later in the onset.
+My aim with this paper is to introduce simple measurements that can be recorded at home and help indivuals recognise when they should seek mdeical attention.
 
@@
 ```
+
 ### The difference between an underweight and normal person:
@@
 
+A new way of estimating general body health is to use WHR.
+This ratio aims to capture the excess fat deposit in a person's abdominal area which is were most people store adipose fat.
+A high WHR indicates high levels of fat.
+However, this is not enough.
+This is simply the most effective way of identyfing the most obvious and dangerous place of fat accumulation.
+
+# Data {#sec-data}
+
+## Overview
+
+### Body Composition Data - Isaac Kuzmar Et Al.
+#### talk about this data
+#### how was it collected
+#### what is collected
+#### graph what i will be using in my model
+
+### Measurements Data - BodyM - AWS.
+#### talk about this data
+#### how was it collected
+#### what is collected
+#### graph what i will be using in my model
+
+## what original BMI and WHR fail to capture
+
+## How are these data sets important are how they relate to each other
+#### desitination
+#### well as any relationships between the variables.
+
 Overview text
@@
 ### Model justification
+- I wanted to provide a categorisation of fat levels in an indivual based on simple measurements.
+Using measurements such as height and weight that are used in the formula and WHR and ankle and wrist measurements, I wanted to capture a person's body compositon as much as possible.
+i expected that there is a positve relationship between a person's excess body fat levels and their WHR.
+In addition to WHR, the height if a person is also important to capture the distribution of fat, a short person with a high WHR would be at a higher risk of obesity related diseases.
+- men are more likely to collect fat in their abdomen area and women in their hip area. To capture nature body shapes and leanness, i also considered....
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
+- height is not all, personality is more important
 ## Second discussion point
@@
 ## Weaknesses and next steps
+- the BodyM dataset does not include age. An older person has lower metabolism [] and a higher chance of collecting fat on their abdomen.
 

### 2024-11-30 scripts/02-data_cleaning.R

@@ -19,11 +19,40 @@ measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
 
 #### Clean data ####
 
+# Written with the help of ChatGPT with modifications
+# fat_categories with waist_hip_ratio as a proxy
+categorize_whr <- function(whr, sex) {
+  if (sex == 2) {
+    if (whr < 0.80) {
+      return("low")
+    } else if (whr >= 0.80 & whr < 0.85) {
+      return("moderate")
+    } else if (whr >= 0.85 & whr < 0.90) {
+      return("high")
+    } else {
+      return("extreme")
+    }
+  } else if (sex == 1) {
+    if (whr < 0.90) {
+      return("low")
+    } else if (whr >= 0.90 & whr < 0.95) {
+      return("moderate")
+    } else if (whr >= 0.95 & whr < 1.00) {
+      return("risk")
+    } else {
+      return("extreme")
+    }
+  } else {
+    return("Invalid sex")
+  }
+}
+
 # 1. MEASUREMENTS DATA (AWS - BodyM)
 combined_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
 
 cleaned_data <- combined_data %>%
   clean_names() %>%
+  mutate(gender = as.numeric(factor(combined_data$gender, levels = c("male", "female")))) %>%
   mutate(waist_hip_ratio = waist / hip) %>%
   mutate(height_hip_ratio = height_cm / hip) %>%
   mutate(original_bmi = weight_kg / (height_cm / 100)^2) %>%
@@ -33,21 +62,12 @@ cleaned_data <- combined_data %>%
     height = height_cm,  
     weight = weight_kg,           
     bmi = original_bmi                
-  )
+  ) %>%
+  select(subject_id, gender, height, weight, ankle, wrist, waist_hip_ratio, height_hip_ratio, bmi) %>%
+  mutate(fat_percentage_category = mapply(categorize_whr, waist_hip_ratio, gender))
+  
 
 cleaned_data$gender <- as.numeric(factor(combined_data$gender, levels = c("male", "female"))) 
-
-# Written with the help of ChatGPT with modifications
-# Create a placeholder column with 2018 rows
-cleaned_data$bmi_category <- factor(NA, levels = c("Underweight", "Normal", "Overweight", "Obese"))
-
-# Assign category based on original_bmi values
-cleaned_data$bmi_category <- ifelse(cleaned_data$bmi < 18.5, "Underweight",
[…]

### 2024-11-30 scripts/02b-data_cleaning.R

@@ -15,6 +15,93 @@ library(dplyr)
 # 2. BODY MASS DATA
 body_mass_data_raw <- read_parquet("data/raw_data/body_mass_data.parquet")
 
+library(dplyr)
+
+# Define thresholds in a function
+assign_female_fat_category <- function(age, fm) {
+  if (age >= 18 & age <= 29) {
+    if (fm < 14) return("low")
+    if (fm <= 16.5) return("excellent")
+    if (fm <= 19.4) return("good")
+    if (fm <= 22.7) return("fair")
+    if (fm <= 27.1) return("poor")
+    return("high")
+  } else if (age >= 30 & age <= 39) {
+    if (fm < 14) return("low")
+    if (fm <= 17.4) return("excellent")
+    if (fm <= 20.8) return("good")
+    if (fm <= 24.6) return("fair")
+    if (fm <= 29.1) return("poor")
+    return("high")
+  } else if (age >= 40 & age <= 49) {
+    if (fm < 14) return("low")
+    if (fm <= 19.8) return("excellent")
+    if (fm <= 23.8) return("good")
+    if (fm <= 27.6) return("fair")
+    if (fm <= 31.9) return("poor")
+    return("high")
+  } else if (age >= 50 & age <= 59) {
+    if (fm < 14) return("low")
+    if (fm <= 22.5) return("excellent")
+    if (fm <= 27.0) return("good")
+    if (fm <= 30.4) return("fair")
+    if (fm <= 34.5) return("poor")
+    return("high")
+  } else if (age >= 60 & age <= 69) {
+    if (fm < 14) return("low")
+    if (fm <= 23.2) return("excellent")
+    if (fm <= 27.9) return("good")
+    if (fm <= 31.3) return("fair")
+    if (fm <= 35.4) return("poor")
+    return("high")
+  } else {
+    return(NA)  # Return NA if age is outside the defined range
+  }
+}
+
+# Define thresholds in a function for men
+assign_male_fat_category <- function(age, fm) {
+  if (age >= 18 & age <= 29) {
+    if (fm < 8) return("low")
+    if (fm <= 10.5) return("excellent")
+    if (fm <= 14.8) return("good")
+    if (fm <= 18.6) return("fair")
+    if (fm <= 23.1) return("poor")
+    return("high")
+  } else if (age >= 30 & age <= 39) {
+    if (fm < 8) return("low")
+    if (fm <= 11.4) return("excellent")
[…]

### 2024-11-30 scripts/04-model.R

@@ -13,38 +13,45 @@ library(tidyverse)
 library(rstanarm)
 library(nnet)
 library(randomForest)
+library(tidyr)
 
 #### Read data ####
-cleaned_data <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
-body_mass_read <- read_parquet("data/analysis_data/body_mass_data.parquet")
+cleaned_data_model <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
+body_mass_model <- read_parquet("data/analysis_data/body_mass_data.parquet")
 
 # Multinomial logistic regression model
 # subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi
 # Variables selected:
-measurements_model <- multinom(bmi_category ~ height + gender + waist_hip_ratio + height_hip_ratio, data = cleaned_data)
+measurements_model <- multinom(fat_percentage_category ~ height + gender + waist_hip_ratio + height_hip_ratio, data = cleaned_data_model)
 # Summary of the model
 summary(measurements_model)
 
 
 #### Read data ####
 # body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
-body_mass_model_bmi <- multinom(bmi_category ~ gender + fm + bmr_kcal + muscle_mass_kg + fm_trunk, data = body_mass_read)
-body_mass_read$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_read)
+body_mass_model_bmi <- multinom(fat_percentage_category ~ gender + fm + bmr_kcal + muscle_mass_kg + fm_trunk, data = body_mass_model)
+body_mass_model$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_model)
 
 
-body_mass_model <- randomForest(fm ~ height + weight + gender * predicted_category + muscle_mass_kg, data = body_mass_read)
+# body_mass_model <- randomForest(fm ~ height + weight + gender * predicted_category + muscle_mass_kg, data = body_mass_read)
 
+data_long <- body_mass_model %>%
+  pivot_longer(cols = c(fat_percentage_category, predicted_category),
+               names_to = "BMI_Category_Type", values_to = "BMI_Category")
 
-# model 2 - final work being done here
[…]

### 2024-11-30 scripts/04c-model.R

@@ -8,37 +8,30 @@ abline(a = 0, b = 1, col = "red", lty = 2) +
 
 
 
-ggplot(body_mass_read) +
-  geom_density(aes(x = fm, fill = "Actual"), alpha = 0.5) +
-  geom_density(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5) +
-  labs(x = "Fat Percentage", y = "Density",
-       title = "Distribution of Actual vs Estimated Fat Percentage") +
-  scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
-  theme_minimal() +
-  facet_wrap(~ bmi_category)
+# fat mass predictor model
+# ggplot(body_mass_read) +
+#   geom_density(aes(x = fm, fill = "Actual"), alpha = 0.5) +
+#   geom_density(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5) +
+#   labs(x = "Fat Percentage", y = "Density",
+#        title = "Distribution of Actual vs Estimated Fat Percentage") +
+#   scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
+#   theme_minimal() +
+#   facet_wrap(~ bmi_category)
 
-ggplot(body_mass_read) +
-  geom_histogram(aes(x = fm, fill = "Actual"), alpha = 0.5, position = "identity", bins = 30) +
-  geom_histogram(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5, position = "identity", bins = 30) +
-  labs(x = "Fat Percentage", y = "Count",
-       title = "Distribution of Actual vs Estimated Fat Percentage") +
-  scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
-  theme_minimal() +
-  facet_wrap(~ bmi_category)
+# ggplot(body_mass_read) +
+#   geom_histogram(aes(x = fm, fill = "Actual"), alpha = 0.5, position = "identity", bins = 30) +
+#   geom_histogram(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5, position = "identity", bins = 30) +
+#   labs(x = "Fat Percentage", y = "Count",
+#        title = "Distribution of Actual vs Estimated Fat Percentage") +
+#   scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
+#   theme_minimal() +
[…]

### 2024-11-30 scripts/101-model_1_EDA.R

@@ -3,75 +3,51 @@ library(tidyverse)
 library(janitor)
 library(arrow)
 library(dplyr)
-# before graphs []
-ggplot(cleaned_data, aes(x = waist_hip_ratio, color = bmi_category, fill = bmi_category)) +
-  geom_density(alpha = 0.3) + 
-  facet_wrap(~ gender) +
+
+eda_measurements <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
+eda_measurements$height_category <- cut(eda_measurements$height, breaks = c(0, 160, 170, 190, 250), 
+                                          labels = c("Short", "Medium", "Tall", "Very Tall"))
+
+ggplot(eda_measurements, aes(x = height, y = waist_hip_ratio)) +
+  geom_point(color = 'blue', size = 2) +
   labs(
-    title = "Overlapping Density Plot for Fat Mass by BMI OG",
-    x = "waist_hip_ratio",
-    y = "Density",
-    color = "BMI Category",
-    fill = "BMI Category"
+    title = "Waist-to-Hip Ratio vs Height",
+    x = "Height (cm)",
+    y = "Waist-to-Hip Ratio"
   ) +
   theme_minimal()
 
-# analysis_data
-ggplot(cleaned_data, aes(x = waist_hip_ratio, color = predicted_category, fill = predicted_category)) +
-  geom_density(alpha = 0.3) + 
-  facet_wrap(~ gender) +
+ggplot(eda_measurements, aes(x = bmi, y = waist_hip_ratio)) +
+  geom_point(color = 'red', size = 2) +
   labs(
-    title = "Overlapping Density Plot for Fat Mass by predicted_category Category 4",
-    x = "waist_hip_ratio",
-    y = "Density",
-    color = "BMI Category",
-    fill = "BMI Category"
+    title = "Waist-to-Hip Ratio vs Weight",
+    x = "Weight (kg)",
+    y = "Waist-to-Hip Ratio"
   ) +
   theme_minimal()
 
-library(ggplot2)
+mean_value <- mean(eda_measurements$waist_hip_ratio, na.rm = TRUE)
+median_value <- median(eda_measurements$waist_hip_ratio, na.rm = TRUE)
 
-ggplot(cleaned_data, aes(x = bmi, fill = interaction(bmi_category, predicted_category))) + 
-  geom_density(alpha = 0.5) + 
-  labs(title = "Comparison of Original and Improved BMI Distributions", 
-       x = "BMI Value", 
[…]

### 2024-11-30 scripts/102-model_2_EDA.R

@@ -5,79 +5,83 @@ library(arrow)
 library(dplyr)
 
 
-# Data Section: Part 1: What is the difference between an underweight / athletic person / normal person:
-bmi_under_25 <- body_mass_clean %>%
-  filter(bmi < 25)
+body_mass_data_eda <- read_parquet("data/analysis_data/body_mass_data.parquet")
+body_mass_data_eda <- body_mass_data_eda %>%
+  select(-bone_mass_kg) %>%
+  filter(gender==2)
+head(body_mass_data_eda)
 
-# this is what i chose
-ggplot(bmi_under_25, aes(x = , y = fm)) +
-  geom_point(aes(color = bmi_category), alpha = 0.7) 
-labs(
-  title = "BMI vs. Fat Mass by BMI Category",
-  x = "BMI",
-  y = "Fat Mass (kg)",
-  color = "BMI Category"
-) +
-  theme_minimal()
-
-
-# Data Section: Part 2: What is the difference between an athletic person / normal person / overweight person:
-bmi_normal_over <- body_mass_clean %>%
-  filter(18.5 < bmi & bmi < 30)
-
-# this is what i chose
-ggplot(bmi_normal_over, aes(x = bmi, y = fm)) +
-  geom_point(aes(color = bmi_category), alpha = 0.7) +
-  facet_wrap(~ bmi_category) +
+ggplot(body_mass_data_eda, aes(x = weight, y = height, color = fat_percentage_category)) +
+  geom_point(size = 3) +
   labs(
-    title = "BMI vs. Fat Mass by BMI Category",
-    x = "BMI",
-    y = "Fat Mass (kg)",
-    color = "BMI Category"
+    title = "Fat Percentage Category vs Height and Weight",
+    x = "Height (cm)",
+    y = "Weight (kg)",
+    color = "Fat Percentage Category"
   ) +
   theme_minimal()
 
+body_mass_data_eda$height_category <- cut(body_mass_data_eda$height, breaks = c(0, 160, 170, 190, 250), 
+                          labels = c("Short", "Medium", "Tall", "Very Tall"))
 
-# all categories
-bmi_normal_over <- body_mass_clean %>%
-  filter(18.5 < bmi & bmi < 30)
-
-# this is what i chose
-ggplot(body_mass_clean, aes(x = bmi, y = fm)) +
-  geom_point(aes(color = bmi_category), alpha = 0.7) +
+# Create a bar plot of fat_percentage_category against height_category
[…]

### 2024-11-30 scripts/02c-data_cleaning.R
cleaning scripts

@@ -1,70 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the Marriage License data
-# Author: Aamishi Avarsekar
-# Date: 19th September 2024
-# Contact: aamishi.avarsekar@mail.utoronto.ca
-# License: MIT
-# Note: Aspects of this code was written with the aid of ChatGPT 4.0
-
-# Work space setup
-library(tidyverse)
-library(janitor)
-library(arrow)
-library(dplyr)
-
-#Read data
-# 1. MEASUREMENTS DATA (AWS - BodyM)
-measurements_subject_data <- read_parquet("data/raw_data/test_subject_data.parquet")
-measurements_data <- read_parquet("data/raw_data/test_data.parquet")
-
-#### Clean data ####
-
-# 1. MEASUREMENTS DATA (AWS - BodyM)
-combined_test_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
-
-cleaned_test_data <- combined_test_data %>%
-  clean_names() %>%
-  mutate(waist_hip_ratio = waist / hip) %>%
-  mutate(original_bmi = weight_kg / (height_cm / 100)^2) %>%
-  mutate(subject_id = seq(1, nrow(combined_test_data))) %>%
-  select(-height) %>%
-  rename(
-    height = height_cm,  
-    weight = weight_kg,           
-    bmi = original_bmi                
-  )
-
-cleaned_test_data$gender <- as.numeric(factor(combined_test_data$gender, levels = c("male", "female"))) - 1
-
-# Written with the help of ChatGPT with modifications
-# Create a placeholder column with 2018 rows
-cleaned_test_data$bmi_category <- factor(NA, levels = c("Underweight", "Normal", "Overweight", "Obese"))
-
-# Assign category based on original_bmi values
-cleaned_test_data$bmi_category <- ifelse(cleaned_test_data$bmi < 18.5, "Underweight",
-                                    ifelse(cleaned_test_data$bmi < 25, "Normal",
-                                           ifelse(cleaned_test_data$bmi < 30, "Overweight", "Obese")))
-
-# Ensure 'category' is a factor with the correct levels
-cleaned_test_data$bmi_category <- factor(cleaned_test_data$bmi_category, levels = c("Underweight", "Normal", "Overweight", "Obese"))
-
-# Changing gender: 1 = male, 2 = female
-# Selecting only: subject_id, gender, height, weight, bmi, bmi_category
[…]

### 2024-11-30 scripts/04b-model.R

@@ -1,39 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-### Model data ####
-library(nnet)
-
-#### Read data ####
-body_mass_read <- read_parquet("data/analysis_data/body_mass_data.parquet")
-body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
-
-library(randomForest)
-body_mass_model <- randomForest(fm ~ weight  + bmi_category, data = body_mass_read)
-
-body_mass_model_bmi <- multinom(bmi_category ~ weight + fm , data = body_mass_read)
-
-
-
-#### Save model ####
-saveRDS(
-  body_mass_model,
-  file = "models/body_mass_model.rds"
-)
-
-saveRDS(
-  body_mass_model_bmi,
-  file = "models/body_mass_model_bmi.rds"
-)
-
-

### 2024-11-30 scripts/04c-model.R

@@ -1,37 +1,0 @@
-plot(body_mass_read$fm, body_mass_read$predicted_fm,
-     xlab = "Actual Fat Percentage",
-     ylab = "Estimated Fat Percentage",
-     main = "Estimated vs Actual Fat Percentage",
-     pch = 16, col = "blue")
-abline(a = 0, b = 1, col = "red", lty = 2) +
-  facet_wrap(~ bmi_category + gender)
-
-
-
-# fat mass predictor model
-# ggplot(body_mass_read) +
-#   geom_density(aes(x = fm, fill = "Actual"), alpha = 0.5) +
-#   geom_density(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5) +
-#   labs(x = "Fat Percentage", y = "Density",
-#        title = "Distribution of Actual vs Estimated Fat Percentage") +
-#   scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
-#   theme_minimal() +
-#   facet_wrap(~ bmi_category)
-
-# ggplot(body_mass_read) +
-#   geom_histogram(aes(x = fm, fill = "Actual"), alpha = 0.5, position = "identity", bins = 30) +
-#   geom_histogram(aes(x = predicted_fm, fill = "Estimated"), alpha = 0.5, position = "identity", bins = 30) +
-#   labs(x = "Fat Percentage", y = "Count",
-#        title = "Distribution of Actual vs Estimated Fat Percentage") +
-#   scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
-#   theme_minimal() +
-#   facet_wrap(~ bmi_category)
-
-# ggplot(body_mass_read) +
-#   geom_violin(aes(x = bmi_category, y = fm, fill = "Actual"), alpha = 0.5) +
-#   geom_violin(aes(x = bmi_category, y = predicted_fm, fill = "Estimated"), alpha = 0.5) +
-#   labs(x = "BMI Category", y = "Fat Percentage",
-#        title = "Violin Plot of Actual vs Estimated Fat Percentage") +
-#   scale_fill_manual(values = c("Actual" = "blue", "Estimated" = "orange")) +
-#   theme_minimal()
-

### 2024-11-30 scripts/101-model_1_EDA.R

@@ -1,53 +1,0 @@
-# Work space setup
-library(tidyverse)
-library(janitor)
-library(arrow)
-library(dplyr)
-
-eda_measurements <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
-eda_measurements$height_category <- cut(eda_measurements$height, breaks = c(0, 160, 170, 190, 250), 
-                                          labels = c("Short", "Medium", "Tall", "Very Tall"))
-
-ggplot(eda_measurements, aes(x = height, y = waist_hip_ratio)) +
-  geom_point(color = 'blue', size = 2) +
-  labs(
-    title = "Waist-to-Hip Ratio vs Height",
-    x = "Height (cm)",
-    y = "Waist-to-Hip Ratio"
-  ) +
-  theme_minimal()
-
-ggplot(eda_measurements, aes(x = bmi, y = waist_hip_ratio)) +
-  geom_point(color = 'red', size = 2) +
-  labs(
-    title = "Waist-to-Hip Ratio vs Weight",
-    x = "Weight (kg)",
-    y = "Waist-to-Hip Ratio"
-  ) +
-  theme_minimal()
-
-mean_value <- mean(eda_measurements$waist_hip_ratio, na.rm = TRUE)
-median_value <- median(eda_measurements$waist_hip_ratio, na.rm = TRUE)
-
-ggplot(eda_measurements, aes(x = weight, y = waist_hip_ratio)) +
-  geom_point(size = 2) +
-  facet_wrap(~ height_category) +  # Facet by age group
-  labs(
-    title = "Waist-to-Hip Ratio vs Height by Age Group",
-    x = "Height (cm)",
-    y = "Waist-to-Hip Ratio",
-    color = "Age Group"
-  ) +
-  theme_minimal()
-
-ggplot(eda_measurements, aes(x = waist_hip_ratio)) +
-  geom_histogram(binwidth = 0.05, fill = "lightblue", color = "black", alpha = 0.7) +
-  geom_vline(aes(xintercept = mean_value), color = "red", linetype = "dashed", size = 1) +
-  geom_vline(aes(xintercept = median_value), color = "green", linetype = "dashed", size = 1) +
-  labs(
-    title = "Waist-to-Hip Ratio Distribution with Mean and Median",
-    x = "Waist-to-Hip Ratio",
-    y = "Frequency"
-  ) +
-  theme_minimal() + 
-  facet_wrap(~ height_category)

### 2024-12-02 paper/paper.qmd
final code stuff - improvements needed

@@
 #| message: false
+#| echo: false
 bmi_canada_gov <- data.frame(
@@
 #| message: false
+#| echo: false
 body_mass_data_raw <- read_parquet("../data/raw_data/body_mass_data.parquet")
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
 bmi_normal_over <- body_mass_clean %>%
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
 # this is what i chose
@@
 ## Overview
+#### General Statement
+- [is this data observational or sample] The data used for this paper has two different sources and serve two different purpose.
+The first dataset considered is the BodyM dataset from Amazon Web Services (AWS) [ref].
+This dataset serves as the training data for the model to predict the category of fat in a person.
+This dataset also included silhouettes to capture accurate body compositions of real test subjects.
+[the purpose of this].
+This data was collected to support the estimation of bodily measurements using Machine Learning [reference].
+The silhouettes are not being used for the purposes of this paper.
+- [is this data observational or sample] The second dataset is the Body Fat Distribution distribution collected for estimating a new body fat measurement technique by Isaac Kuzmar Et Al.
+[reference].
+For the purposes of this paper, this dataset would be refered to as the Body Compositon dataset as it used to calculate fat percentages based on actual body composition.
+It serves as a comparision for the model’s predictions.
+
+### BodyM Dataset - AWS Marketplace for Open Data [fix this heading]
+The BodyM Dataset, refered to teh Body Measurements dataset throughout this paper, was collected for [Ruiz2022] Ruiz Et Al.
+in their paper with a focus on underrepresented body types in estimation of fat and its susbquent health risks.
+The main data captured from this collection was the front and lateral silhouttes of [2000something] test subjects.
+These silhouttes were then converted to black and white images to be used by their augementation model.
+The subjects include X male and Y female aged A to B.
+The body measurements in this dataset were generated through their adversarial body simulator (ABS).
+The ABS was specifically modeled to capture underrespresented body types.
+The S3 [reference] package included 3 datasets: training, Test A and Test B.
+This paper uses the training dataset.
+The visual images of test subjects were photographed and 3D-scanned by lab technicians.
+- Tibble of this dataset after cleaning:
+
+#### The variables of my use:
+
+For this paper, this dataset is used to estimate the category of body fat that a person carries.
+To effectively differentiate the different body types, the model considers measurements that could indicate actual body compositions.
+I chose gender, height and weight and basic predictors for the model.
+However, according to [reference, WHO], waist-to-hip ratio is more accurate to predict levels of body fat in a person.
+Based on [reference], the two sexes have different tendencies of excess body fat collection also known as [adipose fat].
+In males, excess body fat tends to collect in the abdominal region, whereas in females, fat collects in the hips.
+This ratio is indicative of fat accumulation and a higher ratio can be used as a precursor indicator for obesity [reference].
+
+To consider body composition along the height of a person, a new variable called height-to-hip ratio, height_hip_ratio, was created.
+[referernce], something like two people with the same WHR could differ in the fat category that they belong to depending on the height.
+A tall person with a higher WHR is at a lower risk of developing cardiovascular risks than a shorter person with the same height [reference].
+
+Lastly, to improve the quality of the model, I also selected ankle and wrist circumference measurements. [reference], [this is important for boundary values].
+
+- include a table for WHR and who’s statement
+
+#### The estimand for the model
+
+The estimand for the model, fat_percentage_category, is created using WHR as a proxy initially.
+It is calculated based on the subjects gender and WHR as males and females have different fat accumulation patters.
+The ratio intervels for these categories was derived from [].
+The model then predicts a more accurate fat percentage category based on other predictors.
+
+- visualisations.
+
+#### graph what i will be using in my model
 
 ### Body Composition Data - Isaac Kuzmar Et Al.
-#### talk about this data
-#### how was it collected
-#### what is collected
-#### graph what i will be using in my model
+### Data Overview
 
-### Measurements Data - BodyM - AWS.
-#### talk about this data
-#### how was it collected
-#### what is collected
+The dataset compiles the body measurements of subjects aged, 18 and 60, and specifically with a desire to lose weight and improve body image.
+The participants resided in Barranquilla, Colombia and consisted of 234 males and 111 females.
+Medical exclusions were made while recruiting the subjects, such as pregnant women or people with medical pacemakers.
+The bodily measurements such as fat mass in kilograms and fat free mas in kilograms were determined using the Tanita MC-780 [reference] body composition analyzer.
+
+### The variables of my use:
+
+The main variables that I considered fat mass percentage fm.
+This variable is obtained by dividing the fat mass in kilogram by the total weight in kilograms.
+The fat mass percentage is considered the main predictor for fat mass categories.
+As the body composition is more accuarate at measuring the actual amount of excess fat a person carries and can differentiate it from other weights such as fat free muscle mass [reference the variable]
 #### graph what i will be using in my model
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| warning: false
+#| message: false
 #| echo: false
@@
 ```{r}
-
+#| warning: false
+#| message: false
+#| echo: false
 ```
@@
 Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+- multinomial logistic regression because we are predicting multiple categories
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+## Model General Overview:
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+This paper uses 2 models to model the two datasets.
+The first model, the Body Measurements Model, would be used to predict fat percetage categories using the measurements of person’s body that can be obtained using a measuring tape.
+The second model, the Body Composition Model, is also designed to access the relation between different bodiliy compositions such as total body weight, excess fat weight, lean muscle mass in kilograms.
+
+I used the packages [] to do []. Both models employ Multilevel Regression [] as the Fat Percentage Category has multiple categories including: essential, athlete, normal, high.
+
+### Model 1:
+
+In this model, the dependent variable is fat_percentage_category. The model is set up as follows
+
+\begin{align}
+\text{log}\left(\frac{P(y = j)}{P(y = \text{ref})}\right) = \beta_{0j} + \beta_{1j} \cdot \text{height} + \beta_{2j} \cdot \text{gender} + \beta_{3j} \cdot \text{waist\_hip\_ratio} + \beta_{4j} \cdot \text{height\_hip\_ratio}
+\end{align}
+
+Where:
+- \( y \) is the **fat percentage category** (e.g., categories like 'low', 'medium', 'high').
+- \( P(y = j) \) is the probability that the fat percentage falls into category \( j \).
+- \( P(y = \text{ref}) \) is the reference category.
+- \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \).
+
+### Model Justification
+
+It is incredibily challenging to differentiate the body composition of a person based soley on their total weight, height and weight only.
+However, to increase the [goodness] of the classification, this model uses the waist_to_hip ratio, which is especially an important indicator for fat accumualtion across both genders.
+The model also uses height_hip_ratio to account for different body types.
+[this sentence might go in the datasection] As height is an important factor to evaluate the collection of fat in latitdunial(how tf do you spell it) places on a person’s body, this is an important factor to consider.
+However, using height simply does not effectvely communicate how a person’s body is composed latitudinally.
+So I chose to employ the height_hip_ratio ratio to provide more classification between body types.
+Additional features such and ankle and wrist are used to further differentiate between the categories.
+
+### Assumptions and limitations
+
+Assumptions:
+- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.
+
+Limitations:
+
+- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
+- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
+-
+
+### Model validation
+
+---
+
+### Model 2:
+
+In this model, the dependent variable is fat_percentage_category. The model is set up as follows
+
+- fat_percentage_category = beta_0 + beta_1() + beta_() + beta_() + beta_()
+
+where
+
+- fatmass:
+- weight:
+- height_hip_ratio
+
+### Model Justification
+
+### Assumptions and limitations
+
+Assumptions:
+- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.
+
+Limitations:
+
+- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
+- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
+- This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error.
+
+### Model validation:
+
+Model out-of-sample testing:
+
+The dataset provided by BodyM already had a trainining and testing split.
+I used the Test A to test the accuracy of the model.
+As the method of collection was similar to the training data, the potential errors in the model could only arise from the fact that it is unseen data.
+
+More elaboare explanation, tables and figures can be found at [].
+
+RSME:
+I also computed the RSME to evaluate the model’s performance as a lower RSME indicates accuracy in predicted vs actual fat percentage categories.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-modelresults. Our model's results and interpretation:
 
@@
 #| message: false
+
+library(caret)
+
+# Ensure actual categories are factors
+# actual_categories <- factor(cleaned_data_model$fat_percentage_category, 
+#                             levels = c("high", "low", "moderate", "risk"))
+# 
+# # Ensure predicted categories are factors
+# predicted_categories <- factor(predict(measurements_model, newdata = cleaned_data_model, type = "class"),
+#                                 levels = c("high", "low", "moderate", "risk"))
+# 
+# # Now create the confusion matrix
+# 
+# 
+# 
+# conf_matrix <- confusionMatrix(predicted_categories, actual_categories)
+# print(conf_matrix)
+# accuracy <- mean(predicted_categories == actual_categories)
+# print(accuracy)
+# 
+# library(pROC)
+# 
+# probabilities <- predict(measurements_model, newdata = cleaned_data_model, type = "probs")
+# 
+# # Convert true categories to a binary format for ROC computation (one-vs-all approach)
+# roc_result <- multiclass.roc(actual_categories, probabilities)
+# 
+# # Plot ROC curve
+# plot(roc_result)
 

### 2024-12-02 scripts/02-data_cleaning.R

@@ -21,34 +21,49 @@ measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
 
 # Written with the help of ChatGPT with modifications
 # fat_categories with waist_hip_ratio as a proxy
-categorize_whr <- function(whr, sex) {
-  if (sex == 2) {
-    if (whr < 0.80) {
+categorize_bfp <- function(bmi_category) {
+    if (bmi_category == "Underweight") {
       return("low")
-    } else if (whr >= 0.80 & whr < 0.85) {
+    } else if (bmi_category == "Normal") {
       return("moderate")
-    } else if (whr >= 0.85 & whr < 0.90) {
+    } else if (bmi_category == "Overweight") {
       return("high")
     } else {
       return("extreme")
     }
-  } else if (sex == 1) {
-    if (whr < 0.90) {
-      return("low")
-    } else if (whr >= 0.90 & whr < 0.95) {
-      return("moderate")
-    } else if (whr >= 0.95 & whr < 1.00) {
-      return("risk")
-    } else {
-      return("extreme")
-    }
-  } else {
-    return("Invalid sex")
-  }
 }
+
+# categorize_bfp <- function(gender, whr) {
+#   if (gender == 1) {  # Male
+#     if (whr < 0.90) {
+#       return("low")
+#     } else if (whr < 0.95) {
+#       return("moderate")
+#     } else if (whr <= 1.00) {
+#       return("high")
+#     } else {
+#       return("extreme")
+#     }
+#   } else if (gender == 2) {  # Female
+#     if (whr < 0.80) {
+#       return("low")
+#     } else if (whr < 0.85) {
+#       return("moderate")
+#     } else if (whr <= 0.90) {
+#       return("high")
+#     } else {
+#       return("extreme")
+#     }
+#   } else {
+#     return(NA)  # Invalid gender
+#   }
+# }
 
 # 1. MEASUREMENTS DATA (AWS - BodyM)
 combined_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
+
+min_height <- 150
+max_height <- 180
 
 cleaned_data <- combined_data %>%
   clean_names() %>%
@@ -64,10 +79,42 @@ cleaned_data <- combined_data %>%
     bmi = original_bmi                
   ) %>%
   select(subject_id, gender, height, weight, ankle, wrist, waist_hip_ratio, height_hip_ratio, bmi) %>%
[…]

### 2024-12-02 scripts/02b-data_cleaning.R

@@ -18,89 +18,48 @@ body_mass_data_raw <- read_parquet("data/raw_data/body_mass_data.parquet")
 library(dplyr)
 
 # Define thresholds in a function
-assign_female_fat_category <- function(age, fm) {
-  if (age >= 18 & age <= 29) {
-    if (fm < 14) return("low")
-    if (fm <= 16.5) return("excellent")
-    if (fm <= 19.4) return("good")
-    if (fm <= 22.7) return("fair")
-    if (fm <= 27.1) return("poor")
-    return("high")
-  } else if (age >= 30 & age <= 39) {
-    if (fm < 14) return("low")
-    if (fm <= 17.4) return("excellent")
-    if (fm <= 20.8) return("good")
-    if (fm <= 24.6) return("fair")
-    if (fm <= 29.1) return("poor")
-    return("high")
-  } else if (age >= 40 & age <= 49) {
-    if (fm < 14) return("low")
-    if (fm <= 19.8) return("excellent")
-    if (fm <= 23.8) return("good")
-    if (fm <= 27.6) return("fair")
-    if (fm <= 31.9) return("poor")
-    return("high")
-  } else if (age >= 50 & age <= 59) {
-    if (fm < 14) return("low")
-    if (fm <= 22.5) return("excellent")
-    if (fm <= 27.0) return("good")
-    if (fm <= 30.4) return("fair")
-    if (fm <= 34.5) return("poor")
-    return("high")
-  } else if (age >= 60 & age <= 69) {
-    if (fm < 14) return("low")
-    if (fm <= 23.2) return("excellent")
-    if (fm <= 27.9) return("good")
-    if (fm <= 31.3) return("fair")
-    if (fm <= 35.4) return("poor")
-    return("high")
+assign_fat_category <- function(gender, age, fm) {
+  if (gender == 1) {  # Male
+    if (age < 40) {
+      if (fm < 8) return("low")
+      else if (fm <= 20) return("moderate")
+      else if (fm <= 25) return("high")
+      else return("extreme")
+    } else if (age < 60) {
+      if (fm < 11) return("low")
+      else if (fm <= 22) return("moderate")
+      else if (fm <= 28) return("high")
+      else return("extreme")
+    } else {
+      if (fm < 13) return("low")
+      else if (fm <= 25) return("moderate")
+      else if (fm <= 30) return("high")
+      else return("extreme")
+    }
[…]

### 2024-12-02 scripts/04-model.R

@@ -26,44 +26,58 @@ measurements_model <- multinom(fat_percentage_category ~ height + gender + waist
 # Summary of the model
 summary(measurements_model)
 
+cleaned_data_model$predicted_category <- predict(measurements_model, newdata = cleaned_data_model)
+
+
 
 #### Read data ####
 # body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
-body_mass_model_bmi <- multinom(fat_percentage_category ~ gender + fm + bmr_kcal + muscle_mass_kg + fm_trunk, data = body_mass_model)
+body_mass_model_bmi <- multinom(fat_percentage_category ~ height + gender + fm, data = body_mass_model)
+summary(body_mass_model_bmi)
 body_mass_model$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_model)
 
 
 # body_mass_model <- randomForest(fm ~ height + weight + gender * predicted_category + muscle_mass_kg, data = body_mass_read)
 
-data_long <- body_mass_model %>%
-  pivot_longer(cols = c(fat_percentage_category, predicted_category),
-               names_to = "BMI_Category_Type", values_to = "BMI_Category")
+##### FOR THE MODEL's RESULTS SECTION:
+measurements_results <- cleaned_data_model %>%
+  select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
+  mutate(source = "measurements")
 
-# Boxplot - i like 
-ggplot(data_long, aes(x = BMI_Category, y = fm, fill = BMI_Category_Type)) +
-  geom_boxplot(position = position_dodge(0.8), alpha = 0.7) +
-  labs(title = "Fat Mass Percentage by BMI Category", 
-       x = "BMI Category", y = "Fat Mass Percentage (%)") +
-  theme_minimal()
+body_composition_results <- body_mass_model %>%
+  select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
+  mutate(source = "body_composition")
 
-ggplot(data_long, aes(x = fm, fill = BMI_Category_Type)) +
+elongated_data <- bind_rows(measurements_results, body_composition_results)
+
+# View the resulting data
+head(elongated_data)
+
+ggplot(elongated_data, aes(x = fat_percentage_category, fill = source)) +
+  geom_bar(position = "fill") +  # Normalize counts to proportions
[…]

### 2024-12-02 scripts/103-test_data_EDA.R

@@ -6,223 +6,41 @@ library(dplyr)
 install.packages("caret")
 
 library(caret)
+combined_data <- bind_rows(measurements_results, body_composition_results)
+normalized_data <- combined_data %>%
+  group_by(source, fat_percentage_category, predicted_category) %>%
+  tally() %>%
+  mutate(norm_count = n / sum(n)) # Normalizing counts by category total
 
-
-
-
-
-
-
-# Import data:
-# Model 1 Data
-measurements_data_predictions <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
-# Model 2 Data
-body_mass_data_predictions <- read_parquet("data/analysis_data/body_mass_data.parquet")
-
-# Test Data: 
-cleaned_test_data_predictions <- read_parquet("data/analysis_data/test_analysis.parquet")
-
-# Model 1 predictions:
-measurements_data_predictions$predicted_category <- predict(measurements_model, newdata = measurements_data_predictions)
-#table(Predicted = TODO$predicted_category, Actual = cleaned_data$bmi_category)
-
-
-
-
-# Model 2 predictions:
-#table(Predicted = measurements_data_predictions$model_1_predicted_categories, Actual = measurements_data_predictions$bmi_category)
-#table(Predicted = body_mass_data_predictions$model_2_predicted_categories, Actual = body_mass_data_predictions$bmi_category)
-body_mass_data_predictions$predicted_category <- predict(body_mass_model, newdata = body_mass_data_predictions)
-
-# Extract Categories:
-measurements_data_extracted <- measurements_data_predictions %>%
-  select(height, weight, bmi, bmi_category, predicted_category)
-
-body_mass_data_extracted <- body_mass_data_predictions %>%
-  select(height, weight, bmi, bmi_category, predicted_category)
-
-# visual to see results of the models
-# model 1:
-model_1_long <- measurements_data_extracted %>%
-  pivot_longer(cols = c(bmi_category, model_1_predicted_categories), 
-               names_to = "category_type", 
-               values_to = "category")
-
-# Now you can plot the densities
-ggplot(model_1_long, aes(x = bmi, fill = category_type)) + 
-  geom_density(alpha = 0.5) + 
-  facet_wrap(~ category, scales = "free") + 
-  labs(title = "Distribution of Original vs Improved BMI by Category", 
-       x = "BMI Value", 
-       y = "Density") +
[…]

### 2024-12-03 paper/paper.qmd
everything so far

@@
 library(dplyr)
+library(caret)
+library(knitr)
+
+# reading the final datasets:
+cleaned_data_model <- read_parquet(file = here::here("data/analysis_data/measurements_analysis_data.parquet"))
+body_mass_model <- read_parquet(file = here::here("data/analysis_data/body_mass_data.parquet"))
 ```
@@
     ))
-
-# Data Section: Part 1: What is the difference between an underweight / athletic person / normal person:
-bmi_under_25 <- body_mass_clean %>%
-  filter(bmi < 25)
-
-ggplot(bmi_under_25, aes(x = bmi, y = fm)) +
-  geom_point(aes(color = bmi_category), alpha = 0.7) +
-  labs(
-    title = "BMI vs. Fat Mass by BMI Category",
-    x = "BMI",
-    y = "Fat Mass (kg)",
-    color = "BMI Category"
-  ) +
-  theme_minimal()
-
 ```
-
-### The difference between a normal and overweight person:
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-
-bmi_normal_over <- body_mass_clean %>%
-  filter(18.5 < bmi & bmi < 30)
-
-# this is what i chose
-ggplot(bmi_normal_over, aes(x = bmi, y = fm)) +
-  geom_point(aes(color = bmi_category), alpha = 0.7) +
-  labs(
-    title = "BMI vs. Fat Mass by BMI Category",
-    x = "BMI",
-    y = "Fat Mass (kg)",
-    color = "BMI Category"
-  ) +
-  theme_minimal()
-```
+# Data Section: Part 1: What is the difference between an underweight / athletic person /
 
@@
 # this is what i chose
-ggplot(body_mass_clean, aes(x = bmi, y = fm)) +
+ggplot(body_mass_model, aes(x = bmi, y = fm)) +
   geom_point(aes(color = bmi_category), alpha = 0.7) +
@@
 
+```{r}
+# measurements_tibble <- as_tibble(cleaned_data_model)
+# kable(measurements_tibble, caption = "Example Tibble Output")
+```
+
 #### The variables of my use:
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+cleaned_data_model$bmi_category <- factor(cleaned_data_model$bmi_category, 
+                            levels = c("Underweight", "Normal", "Overweight", "Obese"))
+ggplot(cleaned_data_model, aes(x = gender, fill = bmi_category)) +
+  geom_bar(position = "fill") +  # 'fill' makes it a proportion plot
+  ylab("Proportion") +
+  xlab("Gender") +
+  labs(title = "Proportions of BMI Categories by Gender") +
+  scale_y_continuous(labels = scales::percent) +  # Converts y-axis to percentage format
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
+ggplot(cleaned_data_model, aes(x = gender, y = waist_hip_ratio, fill = gender)) +
+  geom_boxplot() +
+  ylab("Waist-to-Hip Ratio") +
+  xlab("Gender") +
+  labs(title = "Distribution of Waist-to-Hip Ratio by Gender") +
+  theme_minimal() +
+  facet_wrap(~fat_percentage_category) + 
+  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
+
+cleaned_data_model$bmi_category <- factor(cleaned_data_model$bmi_category, 
+                            levels = c("Underweight", "Normal", "Overweight", "Obese"))
+ggplot(cleaned_data_model, aes(x = gender, y = height_hip_ratio, fill = gender)) +
+  geom_boxplot() +
+  ylab("Waist-to-Hip Ratio") +
+  xlab("Gender") +
+  labs(title = "Distribution of Waist-to-Hip Ratio by Gender") +
+  theme_minimal() +
+  facet_wrap(~bmi_category) + 
+  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# filtered_data <- subset(cleaned_data_model, bmi_category %in% c("Overweight", "Obese"))
+
+# Create a combined boxplot to visualize differences in ankle and wrist for bmi_category = Overweight and Obese
+
+cleaned_data_model$bmi_category <- factor(cleaned_data_model$bmi_category, 
+                            levels = c("Underweight", "Normal", "Overweight", "Obese"))
+
+long_data <- cleaned_data_model %>%
+  gather(key = "Measurement", value = "Value", ankle, wrist)
+
+# Create side-by-side boxplots for ankle and wrist measurements
+ggplot(long_data, aes(x = bmi_category, y = Value, fill = bmi_category)) +
+  geom_boxplot() +
+  facet_wrap(~ Measurement) +  # This creates separate panels for 'ankle' and 'wrist'
+  ylab("Measurement Value") +
+  xlab("BMI Category") +
+  labs(title = "Comparison of Ankle and Wrist Measurements for Overweight and Obese BMI Categories") +
+  theme_minimal() +
+  scale_fill_manual(values = c("Overweight" = "lightcoral", "Obese" = "lightsalmon")) +
+  theme(legend.position = "none")
+```
+
 #### graph what i will be using in my model
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Recode gender levels
+body_mass_model %>%
+  mutate(gender = factor(gender, levels = c(1, 2), labels = c("Male", "Female"))) %>%
+  mutate(fat_percentage_category = factor(fat_percentage_category,
+                                          levels = c("Low", "Moderate", "High", "Extreme"))) %>%
+  mutate(
+    gender = as.factor(gender),
+    fat_percentage_category = as.factor(fat_percentage_category)
+  )
+
+ggplot(body_mass_model, aes(x = gender, fill = fat_percentage_category)) +
+  geom_bar(position = "fill") +  # 'fill' makes it a proportion plot
+  ylab("Proportion") +
+  xlab("Gender") +
+  labs(title = "Proportions of BMI Categories by Gender") +
+  scale_y_continuous(labels = scales::percent) +  # Converts y-axis to percentage format
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+ggplot(body_mass_model, aes(x = weight, y = height, color = fm)) +
+  geom_point(size = 3) +
+  scale_color_viridis_c() + # Use a continuous color scale
+  labs(title = "Height vs Weight with Fat Mass",
+       x = "Weight",
+       y = "Height",
+       color = "Fat Mass (fm)") +
+  theme_minimal()
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+ggplot(body_mass_model, aes(x = weight, y = height, color = fm, size = fm)) +
+  geom_point(alpha = 0.7) +
+  scale_color_viridis_c() +
+  scale_size_continuous(range = c(1, 10)) +
+  labs(title = "Height vs Weight with Fat Mass",
+       x = "Weight",
+       y = "Height",
+       color = "Fat Mass (fm)",
+       size = "Fat Mass (fm)") +
+  theme_minimal()
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+ggplot(body_mass_model, aes(x = bmi, y = fm, color = gender)) +
+  geom_point(alpha = 0.7) +
+  facet_wrap(~ gender) +
+  labs(title = "Fat Mass vs BMI by Gender",
+       x = "BMI",
+       y = "Fat Mass (fm)",
+       color = "Gender") +
+  theme_minimal()
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+body_mass_long <- body_mass_model %>%
+  pivot_longer(cols = c(fm, bmi), names_to = "measure", values_to = "value")
+
+# Dual y-axis plot
+ggplot(body_mass_long, aes(x = weight, y = value, color = measure)) +
+  geom_point(alpha = 0.5) + # Optional: Add points for raw data
+  geom_smooth(se = FALSE, method = "loess", span = 0.5) + # Smoothed trendline
+  labs(title = "Smoothed Trends of FM and BMI by Age",
+       x = "weight",
+       y = "Value",
+       color = "Measure") +
+  theme_minimal() +
+  facet_wrap(~gender)
+
+```
+- talk about the crowding for men
+
 ## what original BMI and WHR fail to capture
@@
 #### well as any relationships between the variables.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Combine both datasets
+cleaned_test_extract <- cleaned_data_model %>%
+  select(gender, bmi_category, fat_percentage_category)
+body_mass_extracted <- body_mass_model %>%
+  select(gender, bmi_category, fat_percentage_category)
+cleaned_test_extract$dataset <- "Measurements"
+body_mass_extracted$dataset <- "Body Data"
+combined_data <- rbind(cleaned_test_extract, body_mass_extracted)
+
+# Count occurrences of each fat_percentage_category per dataset
+library(dplyr)
+category_counts <- combined_data %>%
+  group_by(dataset, fat_percentage_category) %>%
+  summarise(Count = n(), .groups = "drop")
+
+# Create a bar plot to visualize the differences
+library(ggplot2)
+ggplot(category_counts, aes(x = fat_percentage_category, y = Count, fill = dataset)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Comparison of Fat Percentage Categories Across Data1 and Data2",
+       x = "Fat Percentage Category", y = "Count") +
+  theme_minimal() +
+  scale_fill_manual(values = c("Measurements" = "lightblue", "Body Data" = "lightgreen"))
+
+proportions <- combined_data %>%
+  group_by(dataset, fat_percentage_category) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(Proportion = Count / sum(Count))
+
+# Create a bar plot to visualize the proportions
+ggplot(proportions, aes(x = fat_percentage_category, y = Proportion, fill = dataset)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Proportional Comparison of Fat Percentage Categories Across Data1 and Data2",
+       x = "Fat Percentage Category", y = "Proportion") +
+  scale_y_continuous(labels = scales::percent) +  # Show proportions as percentages
+  theme_minimal() +
+  scale_fill_manual(values = c("Measurements" = "lightblue", "Body Data" = "lightgreen"))
+```
 
@@
 
+<!-- Where: -->
+<!-- \begin{align} -->
+<!-- - \( y \) is the **fat percentage category** (e.g., categories like 'low', 'medium', 'high'). -->
+<!-- - \( P(y = j) \) is the probability that the fat percentage falls into category \( j \). -->
+<!-- - \( P(y = \text{ref}) \) is the reference category. -->
+<!-- - \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \). -->
+<!-- \end{align} -->
+
 Where:
-- \( y \) is the **fat percentage category** (e.g., categories like 'low', 'medium', 'high').
+- \( y \) is the **fat percentage category**. The categories include low, moderate, high, and extreme. These equate low, moderate, high and extreme levels of body fat.
 - \( P(y = j) \) is the probability that the fat percentage falls into category \( j \).
@@
 # Results
+## Model Results for Measurements Model
 
-Our results are summarized in @tbl-modelresults. Our model's results and interpretation:
+Our results are summarized in @tbl-measurementmodelresults. Our model's results and interpretation:
 
@@
 #| message: false
+#| label: tbl-measurementmodelresults
+#| tbl-cap: "TODO"
 
-library(caret)
+# # Code chunk to import models
+measurements_model_paper <- readRDS(file = here::here("models/measurements_model.rds"))
+measurements_model_summary <- summary(measurements_model_paper)
+measurements_summary_df <- as.data.frame(measurements_model_summary$coefficients)
 
-# Ensure actual categories are factors
-# actual_categories <- factor(cleaned_data_model$fat_percentage_category, 
-#                             levels = c("high", "low", "moderate", "risk"))
-# 
-# # Ensure predicted categories are factors
-# predicted_categories <- factor(predict(measurements_model, newdata = cleaned_data_model, type = "class"),
-#                                 levels = c("high", "low", "moderate", "risk"))
-# 
-# # Now create the confusion matrix
-# 
-# 
-# 
-# conf_matrix <- confusionMatrix(predicted_categories, actual_categories)
-# print(conf_matrix)
-# accuracy <- mean(predicted_categories == actual_categories)
-# print(accuracy)
-# 
-# library(pROC)
-# 
-# probabilities <- predict(measurements_model, newdata = cleaned_data_model, type = "probs")
-# 
-# # Convert true categories to a binary format for ROC computation (one-vs-all approach)
-# roc_result <- multiclass.roc(actual_categories, probabilities)
-# 
-# # Plot ROC curve
-# plot(roc_result)
+# Create and style the table
+kable(measurements_summary_df, caption = "Summary of Measurements Model")
+```
 
-```
+## Model Results for Body Mass Model
+
+Our results are summarized in @tbl-bodymass-modelresults. Our model's results and interpretation:
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
+#| label: tbl-bodymass-modelresults
+#| tbl-cap: "TODO"
 
+# # Code chunk to import models
+body_mass_model_paper <- readRDS(file = here::here("models/body_mass_model_bmi.rds"))
+body_mass_model_summary <- summary(body_mass_model_paper)
+body_mass_summary_df <- as.data.frame(body_mass_model_summary$coefficients)
+
+# Create and style the table
+kable(body_mass_summary_df, caption = "Summary of Measurements Model")
 ```
@@
 
-## Third discussion point
+## Second discussion point {#sec-second-point}
 
-## Weaknesses and next steps
+## Third discussion point {#sec-third-point}
+
+## Weaknesses and next steps {#sec-weaknesses-point}
 - the BodyM dataset does not include age. An older person has lower metabolism [] and a higher chance of collecting fat on their abdomen.
@@
 #| warning: false
+
+# predictions <- as.factor(predict(measurements_model_paper, newdata = cleaned_data_model))
+# reference <- as.factor(cleaned_data_model$fat_percentage_category)
+# measurements_cm <- confusionMatrix(predictions, cleaned_data_model$fat_percentage_category)
+# # Extract the confusion matrix
+# measurements_conf_matrix <- as.data.frame(measurements_cm$table)
+# measurements_conf_matrix_df <- as.data.frame(measurements_conf_matrix)
+# rownames(measurements_conf_matrix_df) <- paste("Observed:", rownames(measurements_conf_matrix))
+# colnames(measurements_conf_matrix_df) <- paste("Predicted:", colnames(measurements_conf_matrix))
+# kable(measurements_conf_matrix_df, caption = "Confusion Matrix")
+# 
+# # Extract overall statistics
+# overall_stats <- as.data.frame(t(measurements_cm$overall))
+# kable(overall_stats, caption = "Overall Statistics")
+# 
+# residuals <- residuals(measurements_model_paper, type = "deviance")
+# plot(residuals)
 ```

### 2024-12-03 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Cleans the Marriage License data
+# Purpose: Cleans the Measurements Data (BodyM)
 # Author: Aamishi Avarsekar
-# Date: 19th September 2024
+# Date: 3rd Decemeber 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
 # Note: Aspects of this code was written with the aid of ChatGPT 4.0
@@ -12,54 +12,25 @@ library(janitor)
 library(arrow)
 library(dplyr)
 
-#Read data
-# 1. MEASUREMENTS DATA (AWS - BodyM)
+# Read MEASUREMENTS DATA (AWS - BodyM)
 measurements_subject_data <- read_parquet("data/raw_data/measurements_subject_data.parquet")
 measurements_data <- read_parquet("data/raw_data/measurements_data.parquet")
 
 #### Clean data ####
 
 # Written with the help of ChatGPT with modifications
-# fat_categories with waist_hip_ratio as a proxy
+# fat categories with BMI as a proxy
 categorize_bfp <- function(bmi_category) {
     if (bmi_category == "Underweight") {
-      return("low")
+      return("Low")
     } else if (bmi_category == "Normal") {
-      return("moderate")
+      return("Moderate")
     } else if (bmi_category == "Overweight") {
-      return("high")
+      return("High")
     } else {
-      return("extreme")
+      return("Extreme")
     }
 }
-
-# categorize_bfp <- function(gender, whr) {
-#   if (gender == 1) {  # Male
-#     if (whr < 0.90) {
-#       return("low")
-#     } else if (whr < 0.95) {
-#       return("moderate")
-#     } else if (whr <= 1.00) {
-#       return("high")
-#     } else {
-#       return("extreme")
-#     }
-#   } else if (gender == 2) {  # Female
-#     if (whr < 0.80) {
-#       return("low")
-#     } else if (whr < 0.85) {
-#       return("moderate")
-#     } else if (whr <= 0.90) {
-#       return("high")
-#     } else {
-#       return("extreme")
-#     }
-#   } else {
-#     return(NA)  # Invalid gender
-#   }
-# }
-
-# 1. MEASUREMENTS DATA (AWS - BodyM)
 combined_data <- merge(measurements_subject_data, measurements_data, by = "subject_id")
 
 min_height <- 150
[…]

### 2024-12-03 scripts/02b-data_cleaning.R

@@ -1,9 +1,10 @@
 #### Preamble ####
-# Purpose: TODO
+# Purpose: Cleans the Body Mass Data (Isaaz Kuzmar Et Al)
 # Author: Aamishi Avarsekar
-# Date: TODO
+# Date: 3rd Decemeber 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
+# Note: Aspects of this code was written with the aid of ChatGPT 4.0
 
 #### Workspace setup ####
 library(tidyverse)
@@ -11,47 +12,44 @@ library(janitor)
 library(arrow)
 library(dplyr)
 
-#### Read Data ####
-# 2. BODY MASS DATA
+#### Read BODY MASS DATA
 body_mass_data_raw <- read_parquet("data/raw_data/body_mass_data.parquet")
 
-library(dplyr)
-
-# Define thresholds in a function
+# Define thresholds in a function ( written with ChapGPT)
 assign_fat_category <- function(gender, age, fm) {
   if (gender == 1) {  # Male
     if (age < 40) {
-      if (fm < 8) return("low")
-      else if (fm <= 20) return("moderate")
-      else if (fm <= 25) return("high")
-      else return("extreme")
+      if (fm < 8) return("Low")
+      else if (fm <= 20) return("Moderate")
+      else if (fm <= 25) return("High")
+      else return("Extreme")
     } else if (age < 60) {
-      if (fm < 11) return("low")
-      else if (fm <= 22) return("moderate")
-      else if (fm <= 28) return("high")
-      else return("extreme")
+      if (fm < 11) return("Low")
+      else if (fm <= 22) return("Moderate")
+      else if (fm <= 28) return("High")
+      else return("Extreme")
     } else {
-      if (fm < 13) return("low")
-      else if (fm <= 25) return("moderate")
-      else if (fm <= 30) return("high")
-      else return("extreme")
+      if (fm < 13) return("Low")
+      else if (fm <= 25) return("Moderate")
+      else if (fm <= 30) return("High")
+      else return("Extreme")
     }
   } else if (gender == 2) {  # Female
     if (age < 40) {
-      if (fm < 21) return("low")
-      else if (fm <= 33) return("moderate")
-      else if (fm <= 39) return("high")
-      else return("extreme")
+      if (fm < 21) return("Low")
+      else if (fm <= 33) return("Moderate")
[…]

### 2024-12-03 scripts/04-model.R

@@ -72,6 +72,25 @@ ggplot(elongated_data, aes(x = predicted_category, fill = source)) +
        fill = "Source") +
   theme_minimal()
 
+# Normalize the data to get proportions for each fat_percentage_category and predicted_category within each height_category
+heatmap_data <- elongated_data %>%
+  count(fat_percentage_category, predicted_category, source, height_category) %>%
+  group_by(height_category, fat_percentage_category) %>%
+  mutate(proportion = n / sum(n))  # Normalize by height_category and fat_percentage_category
+
+# Create the heatmap with normalized proportions and faceting by height_category
+ggplot(heatmap_data, aes(x = fat_percentage_category, y = predicted_category, fill = proportion)) +
+  geom_tile() +  # Create heatmap tiles
+  scale_fill_gradient(low = "white", high = "blue") +  # Color scale for proportions
+  labs(title = "Normalized Heatmap of Fat Percentage Category vs Predicted Category (Source Highlighted)",
+       x = "Fat Percentage Category",
+       y = "Predicted Category",
+       fill = "Proportion") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
+
+
+
 
 #### Save model ####
 saveRDS(

### 2024-12-04 paper/paper.qmd
abstract, llm, updated scripts, sim-test, updated paper

@@
 ---
-title: "How General Health Index (GHI) performs better to predict health levels"
-subtitle: "A comparitive analysis of GHI and BMI against body measurements"
+title: "Classification of Body Fat based on Physical Measurements and Body Composition"
+subtitle: "A comparitive analysis of methods of classifying body fat"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "BMI values fail to acknowledge the different body composition across both sexes.
+Using the Waist-Hip-Ratio is more accurate at estimating body fat levels using simple methods at home.
+This paper compares how this model performs against actual body composition such as isolated fat mass.
+Using this method makes self-identification of obsesity more accessible and preventive measures can be taken earlier."
 format: pdf
@@
 #| message: false
-cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
+# cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
 
@@
 # Recode gender levels
-body_mass_model %>%
+body_mass_model_data_section <- body_mass_model %>%
   mutate(gender = factor(gender, levels = c(1, 2), labels = c("Male", "Female"))) %>%
@@
 
-ggplot(body_mass_model, aes(x = gender, fill = fat_percentage_category)) +
+ggplot(body_mass_model_data_section, aes(x = gender, fill = fat_percentage_category)) +
   geom_bar(position = "fill") +  # 'fill' makes it a proportion plot
@@
 # Results
-## Model Results for Measurements Model
-
-Our results are summarized in @tbl-measurementmodelresults. Our model's results and interpretation:
+Our results are summarized in @tbl-measurementmodelresults and @tbl-bodymass-modelresults. Our model's results and interpretation:
 
@@
 ```
-
-## Model Results for Body Mass Model
-
-Our results are summarized in @tbl-bodymass-modelresults. Our model's results and interpretation:
+From the table {} we can see that the intercepts are very low and negative.
+The probability of the model making a classification is very unlikely with the catgegory with the highest probability being high.
+As our data is composed of people with a desire to improve body image, this makes sense, as the people with higher fat percentages are more likely to participate in such a study.
+`waist_hip_ratio` has a negative correlation with the levels of the fat percentage category prediction.
+High `waist_hip_ratio`s are associated with higher body fat percentage.
+So the lower a person's `waist_hip_ratio` is, the less likely the model classifies them in the `high` category.
+On the other hand, `height_hip_ratio` has a positive relation with the classification levels.
+`height_hip_ratio` takes into a consideration the overall mass (and fat) distribution of a person vertically.
+So if a person has a more evenly distributed composition, they would have a higher `height_hip_ratio` ratio with a higher chance of being classified in a lower fat category.
+Despite the model having very few subjects that were underweight, the classification between Underweight and Normal BMI values has improved with more diverse factors.
+It is likely that Normal BMI categories are also being correctly identified with lower fat levels.
+There is a negative correlation with the heights of the subjects suggesting that taller people are less likely to be classified with higher levels of fat.
+Detailed interpretation of this fact is later discussed in the Results Section {}.
 
@@
 # Create and style the table
-kable(body_mass_summary_df, caption = "Summary of Measurements Model")
+kable(body_mass_summary_df, caption = "Summary of Body Mass Model")
+```
+
+The negative intercepts indicate that in the absence of any predictors, the model has very low estimates for the body fat classifications.
+However, the model is most likeyly to classify a person in the `High` category owing to the distribution of the dataset.
+The positive intercept for a `Normal` classification mean that the taller a person is, the more likely they are to be classified as `Normal`.
+As this model is directly based on the actual body composition of a person, the increasing levels of fat percentage indicated by `fm` show that the higher percentage of fat that a person has, the more likely they would be higher in the classification levels.
+
+## Predictions based body circumferences overgeneralize the population
+
+```{r}
+
+# measurements_results <- cleaned_data_model %>%
+#   select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
+#  mutate(source = "Measurements")
+# 
+# body_composition_results <- body_mass_model %>%
+#   select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
+# mutate(source = "Body Composition")
+# 
+# elongated_data <- bind_rows(measurements_results, body_composition_results)
+# 
+# ggplot(elongated_data, aes(x = predicted_category, fill = source)) +
+#   geom_density(alpha = 0.5) +
+#   facet_wrap(~ predicted_category) +
+#   labs(title = "Density Plot of Predicted Categories by Source",
+#        x = "Predicted Category",
+#        y = "Density",
+#        fill = "Source") +
+#   theme_minimal()
+```
+From the above prediction density chart by each model per fat percentage category, we see that the Measurements model has higher peaks at the each classification.
+This suggests that more people are being classified into a similar category only based on their physical traits such as height and Waist-to-Hip ratio.
+The peaks of each category from the Body Composition is accurate in identifying the major classification.
+However, the distribution not being as concentrated as from the Measurements model suggest that the diversification of different body types is being identified properly.
+
+```{r chunk-name}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+levels_results <- union(levels(cleaned_data_model$predicted_category), levels(body_mass_model$predicted_category))
+
+measurements_results <- cleaned_data_model %>%
+  select(height, weight, height_category, fat_percentage_category, predicted_category) %>%
+  mutate(source = "Measurements")
+
+body_composition_results <- body_mass_model %>%
+  select(height, weight, height_category, fat_percentage_category, predicted_category) %>%
+ mutate(source = "BodyComposition")
+
+elongated_data <- bind_rows(measurements_results, body_composition_results)
+
+# Calculate the percentage of each predicted category
+category_percentage_by_source <- elongated_data %>%
+  group_by(source, predicted_category) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100) %>%
+  select(source, predicted_category, percentage)
+
+category_percentage_by_source_height <- elongated_data %>%
+  group_by(source, height_category, predicted_category) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100) %>%
+  select(source, height_category, predicted_category, percentage)
+
+# Print the result
+print(category_percentage_by_source)
+
+ggplot(category_percentage_by_source, aes(x = predicted_category, y = percentage, fill = source)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Percentage of Predicted Categories by Source",
+       x = "Predicted Category",
+       y = "Percentage (%)") +
+  theme_minimal() +
+  scale_fill_manual(values = c("Measurements" = "skyblue", "BodyComposition" = "grey")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels if needed
+```
+## The Body Composition Model has diverse classifcation for different heights groups
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Calculate proportions for each combination of predicted_category, height_category, and source
+category_proportions <- elongated_data %>%
+  group_by(height_category, predicted_category, source) %>%
+  summarise(count = n()) %>%
+  group_by(height_category, source) %>%
+  mutate(percentage = (count / sum(count)) * 100)
+
+# Create the stacked bar plot
+ggplot(category_proportions, aes(x = height_category, y = percentage, fill = predicted_category)) +
+  geom_bar(stat = "identity", position = "stack") +
+  facet_wrap(~ source) +
+  labs(title = "Proportion of Predicted Categories by Height Category and Source",
+       x = "Height Category",
+       y = "Percentage (%)",
+       fill = "Predicted Category") +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
+```
+
+The primary shortfall of the BMI categories was that it failed to differententiate the body types based on other factors such as where the fat was collected.
+From the above graph, we can see that across different heights, the Measurements Model is pretty consistent with providing similar proportions for different fat categories.
+This suggest that it was group people on too broad categories that did not necessarily capture the right effect of height of body compostion.
+Specifically, BMI fails to recognize the vertical spread of fat accumulation and so taller people get naturally classified as having lower fat levels.
+However, the body composition model, has a better ability a higher percentage of people actually have a higher classification of high and extreme body fat across all height groups.
+
+## Differentiating between different levels of Obeseity.
+
+## What BMI fails to capture
+```{r}
+
+  # Function to categorize weight
+
+cleaned_data_model <- cleaned_data_model %>%
+  mutate(weight_category = cut(weight, 
+                               breaks = seq(40, 150, by = 10), # Specify your range
+                               labels = c("40-50", "50-60", "60-70", "70-80", "80-90", "90-100", "100-110", "110-120", "120-130", "130-140", "140-150"),
+                               include.lowest = TRUE))
+
+# Apply the same logic for body_data
+body_mass_model <- body_mass_model %>%
+  mutate(weight_category = cut(weight, 
+                               breaks = seq(40, 150, by = 10), # Specify your range
+                               labels = c("40-50", "50-60", "60-70", "70-80", "80-90", "90-100", "100-110", "110-120", "120-130", "130-140", "140-150"),
+                               include.lowest = TRUE))
+
+# Assuming measurements_data and body_data have the same structure
+measurements_long <- cleaned_data_model %>%
+  select(weight_category, predicted_category, bmi_category) %>%
+  pivot_longer(cols = c(predicted_category, bmi_category),
+               names_to = "category_type",
+               values_to = "category") %>%
+  mutate(source = "Measurements")
+
+body_data_long <- body_mass_model %>%
+  select(weight_category, predicted_category, bmi_category) %>%
+  pivot_longer(cols = c(predicted_category, bmi_category),
+               names_to = "category_type",
+               values_to = "category") %>%
+  mutate(source = "Body Composition")
+
+# Combine both datasets into one for easy plotting
+combined_data <- bind_rows(measurements_long, body_data_long)
+
+  ggplot(combined_data, aes(x = weight_category, fill = category)) +
+  geom_bar(position = "fill", alpha = 0.6) +
+  facet_grid(category_type ~ source, scales = "free_y") +  # Facet by category_type and source
+  labs(title = "Proportions of Categories by Weight Range and Dataset",
+       x = "Weight Range (kg)",
+       y = "Proportion",
+       fill = "Category") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
 ```
@@
 
-## First discussion point {#sec-first-point}
+## Obesity Risk Identification {#sec-first-point}
 
-- height is not all, personality is more important
+This paper's main goal was to compare how accurate are body measurements to estimate fat percentage at home.
+Other ways of estimating a person's body fat include scans like DEXA {source}.
+These scans are expensive and inaccessible.
+Most people turn to measuring BMI's to estimate their general level of health.
+The Government of Cananda also advices that adults use BMI to estimate their health [].
+As seen in this paper, this inherently flawed as BMI fails to capture the composition of a body.
+However, even the introduction of simple measurement like Waist-to-Hip ratio, shows that there is a general improvement in differentiating fat levels.
+The Waist-to-Hip ratio is especially important between the two sexes as fat accumulation pattern differs, which is not even considered in the BMI formula.
+Additionally, as a higher proportion of people are classified as Normal using BMI, without failing to recognise that having a high level of fat despite having a lower BMI is still dangerous as it could potentially prevent them from seeking early care against illnesses like insulin resistance.
+Insulin resistance occurs when there is a body fat percentage of {}.
+By popularising a more accurate system, we could urge people to take earlier preventive care.
+
 ## Second discussion point
@@
 
+# ```{r}
+# modelplot(political_preferences, conf_level = 0.9) +
+#   labs(x = "90 per cent credibility interval")
+# ```
+
 \newpage

### 2024-12-04 scripts/00-simulate_data.R

@@ -1,40 +1,51 @@
 #### Preamble ####
-# Purpose: Simulates Marriage Licence data
+# Purpose: Simulates Measurements (BodyM) data
 # Author: Aamishi Avarsekar
-# Date: 19th September 2024
+# Date: 3rd December 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
-
-### Install packages
-install.packages("tidyverse")
 
 #### Workspace setup ####
 library(tidyverse)
 
-
-#### Simulate data ####
+#### Simulate data #### This code was written by ChatGPT
 set.seed(213)
 
-# Define the start and end date
-start_date <- as.Date("2023-06-01")
-end_date <- as.Date("2023-12-31")
+# Define the number of rows to simulate
+num_rows <- 1000
 
-# Set the number of random dates you want to generate
-number_of_dates <- 200
+# Simulate the data
+subject_id <- paste0("ID_", 1:num_rows)
+gender <- sample(c("Male", "Female"), num_rows, replace = TRUE)
+height <- runif(num_rows, 150, 200)  # Height in cm
+weight <- runif(num_rows, 40, 100)  # Weight in kg
+ankle <- runif(num_rows, 20, 30)  # Ankle circumference in cm
+wrist <- runif(num_rows, 12, 20)  # Wrist circumference in cm
+waist_hip_ratio <- runif(num_rows, 0.6, 1.0)  # Waist-to-hip ratio
+height_hip_ratio <- height / (waist_hip_ratio * 2)  # Example height-to-hip ratio formula
 
-#### Simulate data ####
-sample_marriage_data <-
-  tibble(
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "1970-01-01"
-    ),
-    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
-  )
+# Simulate BMI
+bmi <- weight / (height / 100)^2
+
+# Create BMI categories based on common ranges
+bmi_category <- cut(bmi, breaks = c(-Inf, 18.5, 24.9, 29.9, Inf),
+                    labels = c("Underweight", "Normal", "Overweight", "Obese"))
+
+# Simulate fat percentage categories
+fat_percentage_category <- sample(c("Low", "Moderate", "High", "Extreme"), num_rows, replace = TRUE)
+
+# Height categories (e.g., Low, Medium, High)
[…]

### 2024-12-04 scripts/00b-simulate_data.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Simulates Body Composition data
+# Author: Aamishi Avarsekar
+# Date: 3rd December 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+# This code was written by ChatGPT
+
+set.seed(123)  # For reproducibility
+
+# Number of rows to simulate
+n <- 1000
+
+# Simulate the columns
+body_mass_testing <- data.frame(
+  gender = sample(c("Male", "Female"), n, replace = TRUE),
+  age = sample(18:80, n, replace = TRUE),
+  height = rnorm(n, mean = 170, sd = 10),  # average height with some variability
+  weight = rnorm(n, mean = 70, sd = 15),   # average weight with some variability
+  bmi = rnorm(n, mean = 25, sd = 5),       # typical BMI range
+  fat_mass_kg = rnorm(n, mean = 15, sd = 5), 
+  fm = rnorm(n, mean = 20, sd = 8),        # fat mass in kg
+  ffm_kg = rnorm(n, mean = 50, sd = 10),   # fat-free mass in kg
+  bone_mass_kg = rnorm(n, mean = 3, sd = 1),
+  muscle_mass_kg = rnorm(n, mean = 30, sd = 7),
+  bmr_kcal = rnorm(n, mean = 1500, sd = 300),  # basal metabolic rate
+  fm_trunk = rnorm(n, mean = 7, sd = 2),   # fat mass in trunk area
+  bmi_category = sample(c("Underweight", "Normal", "Overweight", "Obese"), n, replace = TRUE),
+  fat_percentage_category = sample(c("Low", "Moderate", "High", "Extreme"), n, replace = TRUE),
+  height_category = sample(c("Short", "Average", "Tall"), n, replace = TRUE),
+  predicted_category = sample(c("Low", "Moderate", "High", "Extreme"), n, replace = TRUE),
+  weight_category = sample(c("Underweight", "Normal", "Overweight", "Obese"), n, replace = TRUE)
+)
+
+# View first few rows of the simulated data
+head(body_mass_testing)
+

### 2024-12-04 scripts/03-test_data.R

@@ -1,19 +1,94 @@
 #### Preamble ####
-# Purpose: Tests raw marriage license data
+# Purpose: Tests Body Composition data
 # Author: Aamishi Avarsekar
-# Date: 19th September 2024
+# Date: 3rd December 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
+# This was written by ChatGPT
 
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Test data ####
-data <- read_csv("data/raw_data/simulated.csv")
+measurement_data_testing<- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
 
-# Test for negative numbers
-data$number_of_marriage |> min() <= 0
+library(testthat)
 
-# Test for NAs
-all(is.na(data$number_of_marriage))
+# Function to simulate the data
+simulate_data <- function(num_rows = 1000) {
+  subject_id <- paste0("ID_", 1:num_rows)
+  gender <- sample(c("Male", "Female"), num_rows, replace = TRUE)
+  height <- runif(num_rows, 150, 200)  # Height in cm
+  weight <- runif(num_rows, 40, 100)  # Weight in kg
+  ankle <- runif(num_rows, 20, 30)  # Ankle circumference in cm
+  wrist <- runif(num_rows, 12, 20)  # Wrist circumference in cm
+  waist_hip_ratio <- runif(num_rows, 0.6, 1.0)  # Waist-to-hip ratio
+  height_hip_ratio <- height / (waist_hip_ratio * 2)  # Example height-to-hip ratio formula
+  
+  # Simulate BMI
+  bmi <- weight / (height / 100)^2
+  
+  # Create BMI categories based on common ranges
+  bmi_category <- cut(bmi, breaks = c(-Inf, 18.5, 24.9, 29.9, Inf),
+                      labels = c("Underweight", "Normal", "Overweight", "Obese"))
+  
+  # Simulate fat percentage categories
+  fat_percentage_category <- sample(c("Low", "Moderate", "High", "Extreme"), num_rows, replace = TRUE)
+  
+  # Height categories (e.g., Low, Medium, High)
+  height_category <- cut(height, breaks = c(-Inf, 160, 180, Inf), 
+                         labels = c("Short", "Average", "Tall"))
+  
+  # Combine all variables into a data frame
+  data.frame(
+    subject_id, gender, height, weight, ankle, wrist, waist_hip_ratio, 
[…]

### 2024-12-04 scripts/03b-test_data.R

@@ -1,0 +1,68 @@
+#### Preamble ####
+# Purpose: Tests Body Composition data
+# Author: Aamishi Avarsekar
+# Date: 3rd December 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+# This was written by ChatGPT
+
+# Load necessary libraries for testing
+library(testthat)
+
+body_mass_for_testing<- read_parquet("data/analysis_data/body_mass_data.parquet")
+
+# 3. Test if age column is within the expected range (18 to 80 years)
+test_that("Age column is within the expected range", {
+  expect_true(all(body_mass_for_testing$age >= 18 & body_mass_for_testing$age <= 80))
+})
+
+# 4. Test if height column has values greater than zero
+test_that("Height column has positive values", {
+  expect_true(all(body_mass_for_testing$height > 0))
+})
+
+# 5. Test if weight column has values greater than zero
+test_that("Weight column has positive values", {
+  expect_true(all(body_mass_for_testing$weight > 0))
+})
+
+# 6. Test if fat_mass_kg column has non-negative values
+test_that("Fat mass column has non-negative values", {
+  expect_true(all(body_mass_for_testing$fat_mass_kg >= 0))
+})
+
+# 8. Test if fat percentage categories are valid
+test_that("Fat percentage category column contains valid categories", {
+  expect_true(all(body_mass_for_testing$fat_percentage_category %in% c("Low", "Moderate", "High", "Extreme")))
+})
+
+# 9. Test if bmi_category column contains valid categories
+test_that("BMI category column contains valid categories", {
+  expect_true(all(body_mass_for_testing$bmi_category %in% c("Underweight", "Moderate", "Overweight", "Obese")))
+})
+
+# 11. Test if predicted_category column contains valid categories
+test_that("Predicted category column contains valid categories", {
+  expect_true(all(body_mass_for_testing$predicted_category %in% c("Low", "Moderate", "High", "Extreme")))
+})
+
+# 12. Test if weight_category column contains valid categories
+test_that("Weight category column contains valid categories", {
+  expect_true(all(body_mass_for_testing$weight_category %in% c("Underweight", "Normal", "Overweight", "Obese")))
+})
+
+# 13. Test if there are no missing values in the data frame
+test_that("No missing values in the data", {
+  expect_true(all(!is.na(body_mass_for_testing)))
+})
+
[…]

### 2024-12-04 scripts/04-model.R

@@ -1,12 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Model for Measurements Data and Body Composition Data
+# Author: Aamishi Avarsekar
+# Date: 3rd Decemeber 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Note: Aspects of this code was written with the aid of ChatGPT 4.0
 
 # Works pace setup - packages needed
 library(tidyverse)
@@ -20,83 +18,21 @@ cleaned_data_model <- read_parquet("data/analysis_data/measurements_analysis_dat
 body_mass_model <- read_parquet("data/analysis_data/body_mass_data.parquet")
 
 # Multinomial logistic regression model
-# subject_id, gender, height_cm, weight_kg, height, ankle, wrist, waist_hip_ratio, original_bmi
-# Variables selected:
 measurements_model <- multinom(fat_percentage_category ~ height + gender + waist_hip_ratio + height_hip_ratio, data = cleaned_data_model)
-# Summary of the model
-summary(measurements_model)
-
 cleaned_data_model$predicted_category <- predict(measurements_model, newdata = cleaned_data_model)
 
-
-
-#### Read data ####
-# body_mass_model <- multinom(fm ~ bmi_category , data = body_mass_read)
 body_mass_model_bmi <- multinom(fat_percentage_category ~ height + gender + fm, data = body_mass_model)
-summary(body_mass_model_bmi)
 body_mass_model$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_model)
-
-
-# body_mass_model <- randomForest(fm ~ height + weight + gender * predicted_category + muscle_mass_kg, data = body_mass_read)
-
-##### FOR THE MODEL's RESULTS SECTION:
-measurements_results <- cleaned_data_model %>%
-  select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
-  mutate(source = "measurements")
-
-body_composition_results <- body_mass_model %>%
-  select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
-  mutate(source = "body_composition")
-
[…]

### 2024-12-04 scripts/103-test_data_EDA.R

@@ -1,3 +1,12 @@
+#### Preamble ####
+# Purpose: EDA
+# Author: Aamishi Avarsekar
+# Date: 3rd December 2024
+# Contact: aamishi.avarsekar@mail.utoronto.ca
+# License: MIT
+# Note: this code may not work and is simply my eda work
+
+
 # Work space setup
 library(tidyverse)
 library(janitor)

### 2024-12-04 scripts/99-data.R

@@ -1,9 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the Marriage License data
+# Purpose: EDA
 # Author: Aamishi Avarsekar
-# Date: 19th September 2024
+# Date: 3rd December 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca
 # License: MIT
+# Note: this code may not work and is simply my eda work
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-12-04 paper/paper.qmd
intro + citeR

@@
 
-# Introduction
+# Introduction {#paper-intro}
 
@@
 It does not take into consideration the actual body composition of an individual and other important factors such as sex and age.
-Through this paper, I would like to propose a new way categorising people based on other body measurements such as the waist-to-hip ratio and wrist and ankle circumferences measurements as they are more indicative of a person's visceral fat, this proving a more accurate representation of general health.
-I would like to propose a newer formula for General Health Index (GHI) through the BodyM dataset.
-I plan on using the Isaac Kuzmar dataset that has accurate measurements of body fat to create a Confusion Matrix to see how my model compares in providing more accurate body fat estimations through easlily measurable data such as waist and hip circumference.
+In its initial design, BMI was designed using only male [] subjects and thus cannot be applied to females.
+Due to its inability to capture body composition, a lot of false positives and false negatives are produced.
+Through this paper, I would like to compare how using the Waist-to-Hip Ratio (WHR) serves as a more accurate way of estimating your body fat levels at home.
+The inspiration for using the WHR stems from fat accumulation patterns and distribution of fat vertically.
+This also helps us further differentiate between closer classifications of BMI such as Underweight and Normal Categories and Overweight and Obese Categories.
+BMI is notorious for classifications.
+A popular example of this would be body builders being classfied as Overweight according to BMI despite having lower levels for body fat and being at a lower risk of obesity health risks.
+On the other had, taller people are often excused as false negatives and are failed to be recognized as having higher body fat levels due to their height.
+The table {} below includes the BMI categories as endorsed by the Government of Canada.
 
-- this paper uses 2 datasets to capture the difference in fat assessment for popular methods.
-BMI is a popular way of estimating health issues in an individual but it fails to capture details like body composition.
-In its initial design, BMI was designed using only male [] subjects and thus cannot be applied to females.
-However, due to excess popularity, it is used as a self-assessment to estimate your body composition and health risks.
-However, due to its inability to capture body composition, a lot of false positives and false negatives are produced.
-These indivuals often identify health risks [] later in the onset.
-My aim with this paper is to introduce simple measurements that can be recorded at home and help indivuals recognise when they should seek mdeical attention.
-
-## BMI Classification:
 ```{r}
@@
 ```
+
+The second prong of this paper is finding the relation between age, height and weight against actual body fat that is separate from a person's total body weight.
+This serves as a comparison between the classification of body fat levels based on physical measurements versus more anabolic storage of fat.
+Actual body fat is estimated using several methods such as skinfold calipers or DEXA {} scans.
+Both of these methods require professionals to record and experts to decipher.
+It is unusual for common folk to opt for these methods without medication intervention or expert domain knowledge such as athletic coaches.
+
+Through this paper, I analyse how common factors such as sex, height and weight compare against WHR and Body Scan classifications and how imporant are the main predictors for the respective models.
+I employ two multilevel logistical regression models to categorize people based primarily on the WHR and Body Fat Percentages.
+The paper follows a two pronged approach using two different data sets that estimate the classification in two separate ways.
+I then compare their accuracy against their respective data set.
+
+It is a known fact that BMI generalizes several body types based only on height and weight considerations.
+However, through the work of this paper, it can be noted that High and Extreme levels of body fat are completely disregarded through WHR calculations.
+Both data sets and models show an increasing trend in body fat levels as weight increases.
+The model on measurements data predicts a smooth increase of body fat as a person's weight increases.
+However, the model based on body compositions predicts a steeper increase in body fat as a person's weight increases.
+This fact is more obvious when data is faceted by sex.
+
+This paper uses the R Programming Language {@citeR}, to simulate, download, clean and test the data, and create the model.
+More information on packages and specific methods is included later in the paper.
+This paper is structured as follows: The Data Section @ref introduces the data sets, their variables and how they were tailored to obtain the needed information.
+The Model section discusses in greater detail the structure of both models and what information they convey.
+The results and analysis of the models and discussed in the Results section.
+Lastly, the Discussion section discusses real-life implications from the findings of this paper, limitations faced and next steps.
+
+# Data Section {#paper-data-section}
+
+This pap
 
@@
 ```
-# Data Section: Part 1: What is the difference between an underweight / athletic person /
+ Part 1: What is the difference between an underweight / athletic person /
 

### 2024-12-04 scripts/03-test_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests Body Composition data
+# Purpose: Tests Measurements (BodyM) data
 # Author: Aamishi Avarsekar
 # Date: 3rd December 2024
 # Contact: aamishi.avarsekar@mail.utoronto.ca

### 2024-12-04 paper/paper.qmd
more updates to the paper

@@
 More information on packages and specific methods is included later in the paper.
-This paper is structured as follows: The Data Section @ref introduces the data sets, their variables and how they were tailored to obtain the needed information.
+This paper is structured as follows: The Data Section [@paper-data-section] introduces the data sets, their variables and how they were tailored to obtain the needed information.
 The Model section discusses in greater detail the structure of both models and what information they convey.
@@
 
-# Data Section {#paper-data-section}
+# Data Section [#paper-data-section]
 
-This pap
+This paper uses R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot], `arrow` [@citearrow], `dplyr` [@R-dplyr] packages throughout the analysis to clean both data sets and create visualizations.
+`nnet` [@R-nnet] is employed to fit and apply the models.
 
-### The difference between an underweight and normal person:
-Let us take a closer look at the BMI levels below 25 for underweight and normal weight adults.
-
+The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category.
+The graph [@tbl-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
 ```{r}
@@
 #| echo: false
-body_mass_data_raw <- read_parquet("../data/raw_data/body_mass_data.parquet")
-body_mass_clean <- body_mass_data_raw %>%
-  clean_names() %>%
-  select(gender_m1f2, age, height_cm, w_kg, bmi, fat_mass_kg, fm, ffm_kg, bone_mass_kg, muscle_mass_kg) %>%
-  mutate(
-    bmi_category = case_when(
-      bmi < 18.5 ~ "Underweight",
-      bmi >= 18.5 & bmi <= 24.9 ~ "Normal weight",
-      bmi >= 25.0 & bmi <= 29.9 ~ "Overweight",
-      bmi >= 30.0 & bmi <= 34.9 ~ "Obese - Class I",
-      bmi >= 35.0 & bmi <= 39.9 ~ "Obese - Class II",
-      bmi >= 40.0 ~ "Obese - Class III"
-    ))
-```
- Part 1: What is the difference between an underweight / athletic person /
+#| label: fig-bmi-cats-bf
+#| tbl-cap: "BMI vs. Fat Mass by BMI Category"
 
-### All categories
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-
-# this is what i chose
 ggplot(body_mass_model, aes(x = bmi, y = fm)) +
@@
 
-A new way of estimating general body health is to use WHR.
-This ratio aims to capture the excess fat deposit in a person's abdominal area which is were most people store adipose fat.
-A high WHR indicates high levels of fat.
-However, this is not enough.
-This is simply the most effective way of identyfing the most obvious and dangerous place of fat accumulation.
-
 # Data {#sec-data}
@@
 #### General Statement
-- [is this data observational or sample] The data used for this paper has two different sources and serve two different purpose.
-The first dataset considered is the BodyM dataset from Amazon Web Services (AWS) [ref].
-This dataset serves as the training data for the model to predict the category of fat in a person.
-This dataset also included silhouettes to capture accurate body compositions of real test subjects.
-[the purpose of this].
-This data was collected to support the estimation of bodily measurements using Machine Learning [reference].
-The silhouettes are not being used for the purposes of this paper.
-- [is this data observational or sample] The second dataset is the Body Fat Distribution distribution collected for estimating a new body fat measurement technique by Isaac Kuzmar Et Al.
-[reference].
-For the purposes of this paper, this dataset would be refered to as the Body Compositon dataset as it used to calculate fat percentages based on actual body composition.
-It serves as a comparision for the model’s predictions.
+The data used in this paper comes from two distinct sources, each serving a specific purpose:
 
-### BodyM Dataset - AWS Marketplace for Open Data [fix this heading]
-The BodyM Dataset, refered to teh Body Measurements dataset throughout this paper, was collected for [Ruiz2022] Ruiz Et Al.
-in their paper with a focus on underrepresented body types in estimation of fat and its susbquent health risks.
-The main data captured from this collection was the front and lateral silhouttes of [2000something] test subjects.
-These silhouttes were then converted to black and white images to be used by their augementation model.
-The subjects include X male and Y female aged A to B.
-The body measurements in this dataset were generated through their adversarial body simulator (ABS).
-The ABS was specifically modeled to capture underrespresented body types.
-The S3 [reference] package included 3 datasets: training, Test A and Test B.
-This paper uses the training dataset.
-The visual images of test subjects were photographed and 3D-scanned by lab technicians.
-- Tibble of this dataset after cleaning:
+2. **Body Composition Dataset**  
+This dataset, referred to as the Body Composition dataset for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al.
+[@Kuzmar2020].
+It is used to calculate fat percentages based on actual body composition and serves as a comparison for the model’s predictions.
 
-```{r}
-# measurements_tibble <- as_tibble(cleaned_data_model)
-# kable(measurements_tibble, caption = "Example Tibble Output")
-```
+1. **BodyM Dataset**  
+   This dataset, obtained from Amazon Web Services (AWS) [@AWS_Bodym], is used as training data for the model to predict body fat categories.
+It includes silhouettes of real test subjects to capture accurate body compositions.
+These silhouettes were collected to support the estimation of bodily measurements using Machine Learning techniques.
+However, the silhouettes are not used in this paper.
+
+The BodyM Dataset, referred to as the Body Measurements dataset throughout this paper, was collected by Ruiz et al.
+[@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks.
+The primary data captured in this collection consists of the front and lateral silhouettes of approximately [2000] test subjects.
+These silhouettes were then converted into black-and-white images for use in their augmentation model.
+The dataset includes X male and Y female subjects, aged A to B.
+The body measurements in this dataset were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
+The S3 package [@reference] included three datasets: Training, Test A, and Test B.
+For the purposes of this paper, only the Training dataset is used.
+The visual images of the test subjects were photographed and 3D-scanned by lab technicians.
 
@@
 
-For this paper, this dataset is used to estimate the category of body fat that a person carries.
-To effectively differentiate the different body types, the model considers measurements that could indicate actual body compositions.
-I chose gender, height and weight and basic predictors for the model.
-However, according to [reference, WHO], waist-to-hip ratio is more accurate to predict levels of body fat in a person.
-Based on [reference], the two sexes have different tendencies of excess body fat collection also known as [adipose fat].
-In males, excess body fat tends to collect in the abdominal region, whereas in females, fat collects in the hips.
-This ratio is indicative of fat accumulation and a higher ratio can be used as a precursor indicator for obesity [reference].
+For this paper, the BodyM Dataset is used to estimate the category of body fat that a person carries.
+To effectively differentiate between the various body types, the model takes into account measurements that could indicate actual body compositions.
+Gender, height, and weight are used as basic predictors for the model.
+However, according to [@reference, WHO], the waist-to-hip ratio (WHR) is a more accurate predictor of body fat levels.
+Based on [@reference], the two sexes exhibit different tendencies for excess body fat accumulation, also known as adipose fat.
+In males, excess fat tends to accumulate in the abdominal region, whereas in females, fat is more commonly stored in the hips.
+This ratio is indicative of fat distribution, and a higher WHR can be a precursor indicator of obesity [@reference].
 
-To consider body composition along the height of a person, a new variable called height-to-hip ratio, height_hip_ratio, was created.
-[referernce], something like two people with the same WHR could differ in the fat category that they belong to depending on the height.
-A tall person with a higher WHR is at a lower risk of developing cardiovascular risks than a shorter person with the same height [reference].
+To better account for body composition in relation to height, a new variable called height-to-hip ratio (height_hip_ratio) was introduced.
+As noted by [@reference], two individuals with the same WHR could differ in their fat category depending on height.
+For instance, a taller person with a higher WHR may face a lower risk of developing cardiovascular issues than a shorter person with the same WHR [@reference].
 
-Lastly, to improve the quality of the model, I also selected ankle and wrist circumference measurements. [reference], [this is important for boundary values].
+Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included.
+These measurements are particularly important for determining boundary values for fat classification [@reference].
 
-- include a table for WHR and who’s statement
+#### The estimand for the models
 
-#### The estimand for the model
-
-The estimand for the model, fat_percentage_category, is created using WHR as a proxy initially.
-It is calculated based on the subjects gender and WHR as males and females have different fat accumulation patters.
-The ratio intervels for these categories was derived from [].
-The model then predicts a more accurate fat percentage category based on other predictors.
-
-- visualisations.
+Both models employ `fat_percentage_category` as their independent variable or estimand.
+The reason for chosing a category over an actual value was the difference in predictors across the two models.
+Both models use different predictors to estimate `fat_percentage_category` so in order to produce useful comparisons against common predictors such as sex, height and weight, I used a categorical variable.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+I aimed to provide a categorization of body fat levels in an individual based on simple measurements.
+By using measurements such as height, weight, waist-to-hip ratio (WHR), and ankle and wrist circumference, I sought to capture a person’s body composition as accurately as possible.
+I expected a positive relationship between excess body fat levels and WHR.
+Additionally, height is an important factor to consider in capturing fat distribution—specifically, a shorter person with a high WHR would be at a higher risk of obesity-related diseases.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-- multinomial logistic regression because we are predicting multiple categories
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-- I wanted to provide a categorisation of fat levels in an indivual based on simple measurements.
-Using measurements such as height and weight that are used in the formula and WHR and ankle and wrist measurements, I wanted to capture a person's body compositon as much as possible.
-i expected that there is a positve relationship between a person's excess body fat levels and their WHR.
-In addition to WHR, the height if a person is also important to capture the distribution of fat, a short person with a high WHR would be at a higher risk of obesity related diseases.
-- men are more likely to collect fat in their abdomen area and women in their hip area. To capture nature body shapes and leanness, i also considered....
-
+Men are more likely to accumulate fat in the abdominal region, while women tend to store fat in the hip area.
+To capture natural body shapes and leanness, I also considered other factors that influence body composition, such as height-hip ratio that would capture the distribution of fat
+.
 ## Model General Overview:
@@
 
-In this model, the dependent variable is fat_percentage_category. The model is set up as follows
+The model fits a multinomial logistic regression model to predict the fat percentage category (`fat_percentage_category`) based on the following predictors:
 
-- fat_percentage_category = beta_0 + beta_1() + beta_() + beta_() + beta_()
+- `height` (a continuous variable representing the individual's height),
+- `gender` (a categorical variable representing the individual's gender),
+- `fm` (a continuous variable representing fat mass or a related measure).
 
-where
+The model is defined as follows:
 
-- fatmass:
-- weight:
-- height_hip_ratio
+\begin{align}
+P(Y = k \mid X_1, X_2, X_3) = \frac{e^{\beta_{0k} + \beta_{1k} X_1 + \beta_{2k} X_2 + \beta_{3k} X_3}}{1 + \sum_{m=1}^{K-1} e^{\beta_{0m} + \beta_{1m} X_1 + \beta_{2m} X_2 + \beta_{3m} X_3}}
+\end{align}
+
+Where:
+
+- \( P(Y = k \mid X_1, X_2, X_3) \) is the probability of observing the \( k \)-th category of `fat_percentage_category` given the values of `height` (\( X_1 \)), `gender` (\( X_2 \)), and `fm` (\( X_3 \)).
+- \( \beta_{0k}, \beta_{1k}, \beta_{2k}, \beta_{3k} \) are the estimated coefficients for each category \( k \) of the outcome variable, where the reference category is typically set to 0.
+- \( K \) is the total number of categories in the dependent variable `fat_percentage_category`.
+
+This model allows us to assess the relationship between the categorical outcome (`fat_percentage_category`) and the continuous and categorical predictors (`height`, `gender`, `fm`), estimating the probabilities for each category of fat percentage based on the input predictors.
 
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 

### 2024-12-04 scripts/04-model.R

@@ -10,7 +10,6 @@
 library(tidyverse)
 library(rstanarm)
 library(nnet)
-library(randomForest)
 library(tidyr)
 
 #### Read data ####

### 2024-12-04 paper/paper.qmd
before deadline submission

@@
   - Aamishi Avarsekar
-thanks: "Code and data are available at: https://github.com/aamishi/ImprovedGeneralHealthIndex/"
+thanks: "Code and data are available at: https://github.com/aamishi/ImprovedBodyFatClassification"
 date: today
@@
 It does not take into consideration the actual body composition of an individual and other important factors such as sex and age.
-In its initial design, BMI was designed using only male [] subjects and thus cannot be applied to females.
+In its initial design, BMI was designed using only male subjects and thus cannot be applied to females.
 Due to its inability to capture body composition, a lot of false positives and false negatives are produced.
@@
 BMI is notorious for classifications.
-A popular example of this would be body builders being classfied as Overweight according to BMI despite having lower levels for body fat and being at a lower risk of obesity health risks.
+A popular example of this would be body builders being classified as Overweight according to BMI despite having lower levels for body fat and being at a lower risk of obesity health risks.
 On the other had, taller people are often excused as false negatives and are failed to be recognized as having higher body fat levels due to their height.
-The table {} below includes the BMI categories as endorsed by the Government of Canada.
+The table [@tbl-gov-can-bmi] below includes the BMI categories as endorsed by the Government of Canada.
 
@@
 #| echo: false
+#| label: tbl-gov-can-bmi
+#| tbl-cap: "BMI vs. Fat Mass by BMI Category"
 bmi_canada_gov <- data.frame(
@@
 ```
-
-Talk more about it.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-```
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
 

### 2024-12-14 paper/paper.qmd
model section without references

@@
 ---
-title: "Classification of Body Fat based on Physical Measurements and Body Composition"
-subtitle: "A comparitive analysis of methods of classifying body fat"
+title: "Classification of Body Fat based on Measurements and Body Composition"
+subtitle: "Comparitive analysis of Body Fat classification based on BMI and Fat Percentage"
 author: 
@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-# Data Section [#paper-data-section]
+# Data Section {#paper-data-section}
 
@@
 
-The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category.
-The graph [@tbl-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
+The primary motivator of this paper is the misclassification of diverse body types into the same BMI categories.
+The graph [@fig-bmi-cats-bf] is a reference point for the main argument of the paper and {shows} how true body fat percentages differ within the same BMI category.
+As both data sets have different methods of estimating fitness levels, the aim of this paper is to generate a connection between the two data sets.
+In both data sets, I have created the variable `fat_percentage_category` that the model predict based on entirely different predictors measured in different scenarios.
+
+[x] The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category.
+The graph [@fig-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
 ```{r}
@@
 
-# Data {#sec-data}
-
 ## Overview
-#### General Statement
+### General Statement
 The data used in this paper comes from two distinct sources, each serving a specific purpose:
 
-2. **Body Composition Dataset**  
-This dataset, referred to as the Body Composition dataset for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al.
+## **Body Measurements Dataset**  {refer}
+This data set, referred to as the Body Measurements data set for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al.
 [@Kuzmar2020].
+The original paper ....
+
 It is used to calculate fat percentages based on actual body composition and serves as a comparison for the model’s predictions.
 
-1. **BodyM Dataset**  
+This dataset was
+
+## BodyM Data Set
    This dataset, obtained from Amazon Web Services (AWS) [@AWS_Bodym], is used as training data for the model to predict body fat categories.
@@
 These silhouettes were then converted into black-and-white images for use in their augmentation model.
-The dataset includes X male and Y female subjects, aged A to B.
-The body measurements in this dataset were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
+The data set includes X male and Y female subjects, aged A to B.
+The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
 The S3 package [@reference] included three datasets: Training, Test A, and Test B.
-For the purposes of this paper, only the Training dataset is used.
+For the purposes of this paper, only the Training data set is used.
 The visual images of the test subjects were photographed and 3D-scanned by lab technicians.
@@
 
-For this paper, the BodyM Dataset is used to estimate the category of body fat that a person carries.
-To effectively differentiate between the various body types, the model takes into account measurements that could indicate actual body compositions.
-Gender, height, and weight are used as basic predictors for the model.
-However, according to [@reference, WHO], the waist-to-hip ratio (WHR) is a more accurate predictor of body fat levels.
-Based on [@reference], the two sexes exhibit different tendencies for excess body fat accumulation, also known as adipose fat.
+For this paper, the BodyM Dataset is used to estimate the category of body fat that a person carries using circumference measurements.
+This means the circumferential length of body parts such as the waist and the hips.
+To build the model, I chose standard predictors such as gender, height and the weight of subjects as the base predictors.
+However, these predictors are not effective at differentiating body types.
+Therefore, to effectively construct the distribution of body fat in human body only through measurements, I created two variables called `waist_hip_ration` and `height_hip_ratio`.
+These variables are used to model the differences in the lateral and horizontal fat distributions.
+According to WHO [@reference], the waist-to-hip ratio, WHR, is a more accurate predictor of body fat around the abdominal area.
+Based on the study done in [@reference], the two sexes exhitbit difference tendenices for excess body fat accumulation, also known as the adipose fat.
 In males, excess fat tends to accumulate in the abdominal region, whereas in females, fat is more commonly stored in the hips.
-This ratio is indicative of fat distribution, and a higher WHR can be a precursor indicator of obesity [@reference].
-
-To better account for body composition in relation to height, a new variable called height-to-hip ratio (height_hip_ratio) was introduced.
-As noted by [@reference], two individuals with the same WHR could differ in their fat category depending on height.
-For instance, a taller person with a higher WHR may face a lower risk of developing cardiovascular issues than a shorter person with the same WHR [@reference].
-
+This ratio is indicative of fat distribution, and a higher WHR can be a precursing indicator of obesity [@reference].
+In the study done in [place], it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
 Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included.
 These measurements are particularly important for determining boundary values for fat classification [@reference].
+Excess fat accumulation at the wrists and ankles could indicate other health risks as seen in [].
 
@@
 
-Both models employ `fat_percentage_category` as their independent variable or estimand.
-The reason for chosing a category over an actual value was the difference in predictors across the two models.
-Both models use different predictors to estimate `fat_percentage_category` so in order to produce useful comparisons against common predictors such as sex, height and weight, I used a categorical variable.
+Using the above predictor variables, I contructed the variable `fat_percentage_category` as I recruit a multilevel linear regression model to classify the body fat.
+Initially, I wanted to predict the actual fat percentage, but this proved difficult as ...
+I decided to use a categorical variable as the estimand to produced useful comparisons against other categorical variables such as sex, height categories and weight categories.
+Additionally, `fat_percentage_category` was constructed using BMI as a proxy to gauge fat levels in a person.
+BMI is still widely used in medical scenarios to provide general diagnosis of health levels so I used that as a base to study the interactions between current systems and new offerings of WHR and height_hip_ratio.
+The relationship between `fat_percentage_category` and BMI was considered to be relatively straightforward to mimick the nature of the current applications of BMI.
+For BMI, I chose to consider the categories, `underweight`, `normal`, `overweight` and `obese`.
+These were calulated using the BMI formula:
+[insert BMI formula on a new line].
+
+No difference between the genders was considered.
+Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, [ii] and [iii].
+However, due to the lack of data points for Obesity II and III (BMI X - X), I chose to categorize all levels of obesity within the same `obesity` category.
+
+The following table shows the relationship between the new `fat_percentage_category` against existing the `bmi_category`.
 
@@
 ```
+
 ```{r}
@@
 
-## Measurement
+## Measurement of Data
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+### Measurements Data Set
+
+### Body Composition Data Set
 
@@
 
-# Model
+\newpage
+# Models
 
+Both models, the Measurements Model and the Body Composition Model have the same independent variable that is based on different proxies.
+For a gross estimation, the primary proxy for each model against Fat Percentage Category is included in [@tbl-gross-estimation] below.
+Reasons behind this decision are discussed in greater detail in each model section.
+Both models are Multilevel Regression Models to predict `fat_percentage_category` into categories of "Low", "Moderate", "High" and "Extreme" levels.
+I used the `nnet` (@R-nnet) package in R (@citeR) to create both multilevel regression models.
+Model diagnostics including error checks is included in Appendix [@sec-model-details] along with more elaborate explanations, tables and figures.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-gross-estimation
+#| tbl-cap: "Fat Percentage Category against proxies: BMI Category and Actual Fat Percentage"
+
+# Data for the table
+gross_estimation <- data.frame(
+  Fat_Percentage_Category = c("Low", "Moderate", "High", "Extreme"),
+  BMI_Category = c("Underweight", "Normal", "Overweight", "Obese"),
+  Male = c("<8%", "8-20%", "21-25%", ">25%"),
+  Female = c("<21%", "21-33%", "34-39%", ">39%")
+)
+
+# Create the Kable table
+kable(
+  gross_estimation,
+  col.names = c("Fat Percentage Category", "BMI", "Fat Percentage (Male)", "Fat Percentage (Female)"))
+```
+
+## Measurements Model
 I aimed to provide a categorization of body fat levels in an individual based on simple measurements.
-By using measurements such as height, weight, waist-to-hip ratio (WHR), and ankle and wrist circumference, I sought to capture a person’s body composition as accurately as possible.
+By using measurements such as height, weight, waist-hip-ratio (WHR), and ankle and wrist circumferences, I sought to capture a person’s body composition as accurately as possible.
 I expected a positive relationship between excess body fat levels and WHR.
-Additionally, height is an important factor to consider in capturing fat distribution—specifically, a shorter person with a high WHR would be at a higher risk of obesity-related diseases.
+Additionally, height is an important factor to consider in capturing fat distribution — specifically, a shorter person with a high WHR would be at a higher risk of obesity-related diseases.
+As men are more likely to accumulate fat in the abdominal region, while women tend to store fat in the hip area.
+To capture natural body shapes and contours, the height-hip-ratio is also a dominant predictor in the model.
 
-Men are more likely to accumulate fat in the abdominal region, while women tend to store fat in the hip area.
-To capture natural body shapes and leanness, I also considered other factors that influence body composition, such as height-hip ratio that would capture the distribution of fat
-.
-## Model General Overview:
+In this model, the dependent variable is `fat_percentage_category`. The model is set up as follows
 
-This paper uses 2 models to model the two datasets.
-The first model, the Body Measurements Model, would be used to predict fat percetage categories using the measurements of person’s body that can be obtained using a measuring tape.
-The second model, the Body Composition Model, is also designed to access the relation between different bodiliy compositions such as total body weight, excess fat weight, lean muscle mass in kilograms.
-
-I used the packages [] to do []. Both models employ Multilevel Regression [] as the Fat Percentage Category has multiple categories including: essential, athlete, normal, high.
-
-### Model 1:
-
-In this model, the dependent variable is fat_percentage_category. The model is set up as follows
-
+<!-- my model eq -->
 \begin{align}
-\text{log}\left(\frac{P(y = j)}{P(y = \text{ref})}\right) = \beta_{0j} + \beta_{1j} \cdot \text{height} + \beta_{2j} \cdot \text{gender} + \beta_{3j} \cdot \text{waist\_hip\_ratio} + \beta_{4j} \cdot \text{height\_hip\_ratio}
+\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{baseline} \mid X)} \right) = \beta_{0k} + \beta_{1k} \cdot \text{height} + \beta_{2k} \cdot \text{gender} + \beta_{3k} \cdot \text{waist\_hip\_ratio} + \beta_{4k} \cdot \text{height\_hip\_ratio}
 \end{align}
 
-<!-- Where: -->
-<!-- \begin{align} -->
-<!-- - \( y \) is the **fat percentage category** (e.g., categories like 'low', 'medium', 'high'). -->
-<!-- - \( P(y = j) \) is the probability that the fat percentage falls into category \( j \). -->
-<!-- - \( P(y = \text{ref}) \) is the reference category. -->
-<!-- - \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \). -->
-<!-- \end{align} -->
+Where
 
-Where:
-- \( y \) is the **fat percentage category**. The categories include low, moderate, high, and extreme. These equate low, moderate, high and extreme levels of body fat.
-- \( P(y = j) \) is the probability that the fat percentage falls into category \( j \).
-- \( P(y = \text{ref}) \) is the reference category.
-- \( \beta_{0j}, \beta_{1j}, \dots \) are the coefficients to be estimated for each category \( j \).
+TODO !!!!!! change this to normal
+
+- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where \( k \in \{1, 2, \dots, K\} \)) given the predictors $X$.
+- $\text{baseline}$ is the reference category (the category of the outcome variable that is omitted in the model, typically the first category).
+- $\beta_{0k}$, $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the intercept and the predictors for category \( k \).
+- The predictors in this model are `height`, `gender`, `waist_hip_ratio`, and `height_hip_ratio`.
 
@@
 
-It is incredibily challenging to differentiate the body composition of a person based soley on their total weight, height and weight only.
-However, to increase the [goodness] of the classification, this model uses the waist_to_hip ratio, which is especially an important indicator for fat accumualtion across both genders.
-The model also uses height_hip_ratio to account for different body types.
-[this sentence might go in the datasection] As height is an important factor to evaluate the collection of fat in latitdunial(how tf do you spell it) places on a person’s body, this is an important factor to consider.
-However, using height simply does not effectvely communicate how a person’s body is composed latitudinally.
-So I chose to employ the height_hip_ratio ratio to provide more classification between body types.
+It is incredibly challenging to differentiate the body composition of a person based solely on their total weight and height only.
+However, to increase the accuracy of the classification, this model uses the `waist_to_hip` ratio, which is especially an important indicator for fat accumulation across both genders.
+The model also uses `height_hip_ratio` to account for different body types.
+While height is an important physical attribute, simply using the height variable does not effectively capture how a person’s body is composed horizontally.
+So, I chose to employ the `height_hip_ratio` ratio to provide more classification between body types.
 Additional features such and ankle and wrist are used to further differentiate between the categories.
+As this model uses BMI as a proxy for body fat, using both height and weight as interacting predictors produced no difference in the new predicted categories of `fat_percentage_category` as the same variables were indirectly used as proxies to assign `fat_percentage_category` using the original BMI formula.
+Therefore, I chose to only consider height.
+Another reason for opting out of weight as a predictor, was again, the inability of the model to differentiate between fat and muscle mass.
+So removing weight altogether was successful in the new classification of fat level.
 
@@
 Assumptions:
-- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.
+
+his model assumes that there is a positive relationship with the higher classification of the fat percentage category and the fat storing tendencies in males and females.
+Other physical ailments that would cause an increase in the circumference measurements without increasing the overall fat in the body is not considered.
+This model assumes that females with a higher WHR and males with a higher height-hip ratio are classified higher in the fat categories.
 
@@
 
-- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
-- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
--
+The training data set for this model does not include the age of the participants.
+This is a huge drawback when it comes to assigning risk labels as aging changes how fat is perceived in the bod.
+Hence, this model would not be able to differentiate between age groups and would work poorly if this information is included.
+I also considered estimating actual the fat percentage instead of fat percentage **category**.
+However, to increase the comparability across both models, I chose to use a categorical variable instead.
 
-### Model validation
+## Body Composition Model
 
----
-
-### Model 2:
-
-The model fits a multinomial logistic regression model to predict the fat percentage category (`fat_percentage_category`) based on the following predictors:
-
-- `height` (a continuous variable representing the individual's height),
-- `gender` (a categorical variable representing the individual's gender),
-- `fm` (a continuous variable representing fat mass or a related measure).
-
-The model is defined as follows:
+The model fits a multinomial logistic regression model to predict the fat percentage category (`fat_percentage_category`).
+The model is defined as follows under the assumption that "Normal" is the baseline category of fat classification:
 
 \begin{align}
-P(Y = k \mid X_1, X_2, X_3) = \frac{e^{\beta_{0k} + \beta_{1k} X_1 + \beta_{2k} X_2 + \beta_{3k} X_3}}{1 + \sum_{m=1}^{K-1} e^{\beta_{0m} + \beta_{1m} X_1 + \beta_{2m} X_2 + \beta_{3m} X_3}}
+\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{Normal} \mid X)} \right) = \beta_{1k} \cdot \text{age} + \beta_{2k} \cdot \text{height} + \beta_{3k} \cdot \text{gender} + \beta_{4k} \cdot \text{fm}
 \end{align}
@@
 
-- \( P(Y = k \mid X_1, X_2, X_3) \) is the probability of observing the \( k \)-th category of `fat_percentage_category` given the values of `height` (\( X_1 \)), `gender` (\( X_2 \)), and `fm` (\( X_3 \)).
-- \( \beta_{0k}, \beta_{1k}, \beta_{2k}, \beta_{3k} \) are the estimated coefficients for each category \( k \) of the outcome variable, where the reference category is typically set to 0.
-- \( K \) is the total number of categories in the dependent variable `fat_percentage_category`.
+- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where (k \in \{\text{Low}, \text{High}, \text{Extreme) given the predictors $X$.
+- $\text{Normal}$ is the baseline category (the omitted category), meaning the model compares all other categories to "Normal".
+- $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the predictors for category \( k \), relative to the baseline category "Normal".
+- The predictors in this model are `age`, `height`, `gender`, and `fm` (which stands for fat mass percentage).
 
-This model allows us to assess the relationship between the categorical outcome (`fat_percentage_category`) and the continuous and categorical predictors (`height`, `gender`, `fm`), estimating the probabilities for each category of fat percentage based on the input predictors.
+This model allows us to assess the relationship between the categorical outcome (`fat_percentage_category`) and the continuous and categorical predictors (`age`, `height`, `gender`, `fm`), estimating the probabilities for each category of fat percentage based on the input predictors.
 
 ### Model Justification
+
+The equation for the model expresses the **log-odds** of being in a given category of `fat_percentage_category` with respect to the baseline category "Normal".
+For each non-baseline category, the model estimates a separate set of coefficients that describe the log-odds of being in that category compared to being in the "Normal" category.
+The existence of excess fat in a person has been associated with increased obesity related risks.
+This model primarily focuses on actual fat in a person and is the main predictor variable of interest.
+I have considered only two other predictors: `gender` and `height` of a person as there is physiological difference in how access fat affects males and females of persons of different height across different ages.
+So, `age` is also considered as a predictor variable.
+Additionally, there are different thresholds to classify a person based on the same fat percentage.
+As seen in the previous model, I want to study the precise difference in estimation of fat levels when weight is not considered holistically.
+Hence, `gender` and `height` were considered to be comparable with the Body Measurements Models.
+I had also considered using muscle mass of the subjects.
+However, the data was skewed toward obese people as this data sets was specifically consisted of people who wanted to improve body image.
+I decided to omit muscle mass as it interacted with the fat mass variable and could be indirectly interpreted as total body weight, suggesting over fitting.
+
+Similar to the Body Measurements Model, this model also considers `fat_percentage_category` to be the estimand.
+According to the study [], the two sexes have different thresholds for different ages, suggesting linkages to different metabolic needs based on different phases of life.
+Hence, `fat_percentage_category` is initially classified using `gender`, `age` and `fm`.
 
@@
 Assumptions:
-- this model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.
+
+This model assumes that there is an obvious difference in body measurements and fat storing capacities in males and females, hence gender is an important predictor and also affects the proxy values of fat percentage owing to the calculation based on WHR.
 
@@
 
-- The dataset does not include the age of the particapants. while a general age range is provided, such as 18-X years old, different age groups could differ in how they are categorised.
-- This model would not be able to differentiate between age groups and hence would work poorly if this information is included.
-- This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error.
+This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error.
+While this is an indirect form of observed data, the original data set claims [x] accuracy.
+However, the error that it could cause in this paper is not accounted for and is considered the true measurement of the test subject.
 
-### Model validation:
-
-Model out-of-sample testing:
-
-The dataset provided by BodyM already had a trainining and testing split.
-I used the Test A to test the accuracy of the model.
-As the method of collection was similar to the training data, the potential errors in the model could only arise from the fact that it is unseen data.
-
-More elaboare explanation, tables and figures can be found at [].
-
-RSME:
-I also computed the RSME to evaluate the model’s performance as a lower RSME indicates accuracy in predicted vs actual fat percentage categories.
-
+\newpage
 # Results
@@
 
-# Model details {#sec-model-details}
+# Models Validation {#sec-model-details}
 

### 2024-12-14 scripts/02-data_cleaning.R

@@ -61,7 +61,7 @@ cleaned_data <- combined_data %>%
   # mutate(fat_percentage_category = mapply(categorize_bfp, bmi_category)) %>%
   mutate(fat_percentage_category = factor(
     mapply(categorize_bfp, bmi_category),
-    levels = c("Low", "Moderate", "High", "Extreme") 
+    levels = c("Moderate", "Low", "High", "Extreme") 
   )) %>%
   mutate( # This part was written with the help of ChatGPT
     height_category = cut(
@@ -72,6 +72,7 @@ cleaned_data <- combined_data %>%
       right = FALSE  # Make intervals left-closed, so 150 goes into the "150-160" range
     )
   )
+
 
 #### Save data ####
 write_parquet(cleaned_data, "data/analysis_data/measurements_analysis_data.parquet")

### 2024-12-14 scripts/02b-data_cleaning.R

@@ -79,7 +79,7 @@ body_mass_clean <- body_mass_data_raw %>%
   # mutate(fat_percentage_category = mapply(assign_fat_category, gender, age, fm)) %>%
   mutate(fat_percentage_category = factor(
     mapply(assign_fat_category, gender, age, fm),
-    levels = c("Low", "Moderate", "High", "Extreme") 
+    levels = c("Moderate", "Low", "High", "Extreme") 
   )) %>%
   mutate(
     height_category = cut(

### 2024-12-14 scripts/04-model.R

@@ -8,9 +8,12 @@
 
 # Works pace setup - packages needed
 library(tidyverse)
-library(rstanarm)
+library(arrow)
 library(nnet)
 library(tidyr)
+library(brms)
+
+
 
 #### Read data ####
 cleaned_data_model <- read_parquet("data/analysis_data/measurements_analysis_data.parquet")
@@ -20,10 +23,10 @@ body_mass_model <- read_parquet("data/analysis_data/body_mass_data.parquet")
 measurements_model <- multinom(fat_percentage_category ~ height + gender + waist_hip_ratio + height_hip_ratio, data = cleaned_data_model)
 cleaned_data_model$predicted_category <- predict(measurements_model, newdata = cleaned_data_model)
 
-body_mass_model_bmi <- multinom(fat_percentage_category ~ height + gender + fm, data = body_mass_model)
+body_mass_model_bmi <- multinom(fat_percentage_category ~ age + height + gender + fm, data = body_mass_model)
 body_mass_model$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_model)
 
-
+summary(measurements_model)
 
 #### Save model ####
 write_parquet(cleaned_data_model, "data/analysis_data/measurements_analysis_data.parquet")

### 2024-12-15 paper/paper.qmd
more formal paper with intro refs

@@
 On the other had, taller people are often excused as false negatives and are failed to be recognized as having higher body fat levels due to their height.
-The table [@tbl-gov-can-bmi] below includes the BMI categories as endorsed by the Government of Canada.
+The table [@tbl-gov-can-bmi] below includes the BMI categories as endorsed by the Government of Canada (@CanadaBMI2024).
 
@@
 This serves as a comparison between the classification of body fat levels based on physical measurements versus more anabolic storage of fat.
-Actual body fat is estimated using several methods such as skinfold calipers or DEXA {} scans.
+Actual body fat is estimated using several methods such as skinfold calipers or DEXA scans (@Kuzmar2020).
 Both of these methods require professionals to record and experts to decipher.
@@
 
-This paper uses the R Programming Language {@citeR}, to simulate, download, clean and test the data, and create the model.
+This paper uses @citeR, to simulate, download, clean and test the data, and create the model.
 More information on packages and specific methods is included later in the paper.
-This paper is structured as follows: The Data Section [@paper-data-section] introduces the data sets, their variables and how they were tailored to obtain the needed information.
+This paper is structured as follows: The \[Data Section\](@paper-data-section) introduces the data sets, their variables and how they were tailored to obtain the needed information.
 The Model section discusses in greater detail the structure of both models and what information they convey.
@@
 
-This paper uses R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot], `arrow` [@citearrow], `dplyr` [@R-dplyr] packages throughout the analysis to clean both data sets and create visualizations.
+This paper uses R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeGgplot], `arrow`(@citeArrow), `dplyr` [@R-dplyr] packages throughout the analysis to clean both data sets and create visualizations.
 `nnet` [@R-nnet] is employed to fit and apply the models.
@@
 The primary motivator of this paper is the misclassification of diverse body types into the same BMI categories.
-The graph [@fig-bmi-cats-bf] is a reference point for the main argument of the paper and {shows} how true body fat percentages differ within the same BMI category.
+The graph [@fig-bmi-cats-bf] is a reference point for the main argument of the paper and shows how true body fat percentages differ within the same BMI category.
 As both data sets have different methods of estimating fitness levels, the aim of this paper is to generate a connection between the two data sets.
 In both data sets, I have created the variable `fat_percentage_category` that the model predict based on entirely different predictors measured in different scenarios.
+For each of the two data sets and their respective models, I have used the proxies, BMI and Body Fat Percentage, to determine the initial value for `fat_percentage_category`.
+The models then predict the new categories based on factors such as WHR and Body Fat Percentage.
+More discussion is included in the respective model section.
 
-[x] The primary motivator of this paper was the diverse body compositions that were classified under the same BMI category.
-The graph [@fig-bmi-cats-bf] shows the diverse values of Body Fat Percentages classified in the same BMI category.
 ```{r}
@@
 
-## Overview
-### General Statement
-The data used in this paper comes from two distinct sources, each serving a specific purpose:
+## BodyM Data Set
 
-## **Body Measurements Dataset**  {refer}
-This data set, referred to as the Body Measurements data set for the purposes of this paper, was collected for estimating a new body fat measurement technique by Isaac Kuzmar et al.
-[@Kuzmar2020].
-The original paper ....
-
-It is used to calculate fat percentages based on actual body composition and serves as a comparison for the model’s predictions.
-
-This dataset was
-
-## BodyM Data Set
 This dataset, obtained from Amazon Web Services (AWS) [@AWS_Bodym], is used as training data for the model to predict body fat categories.
@@
 
-The BodyM Dataset, referred to as the Body Measurements dataset throughout this paper, was collected by Ruiz et al.
+The BodyM Data Set, referred to as the Body Measurements data set throughout this paper, was collected by Ruiz et al.
 [@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks.
-The primary data captured in this collection consists of the front and lateral silhouettes of approximately [2000] test subjects.
+The primary data captured in this collection consists of the front and lateral silhouettes of approximately \[2000\] test subjects.
 These silhouettes were then converted into black-and-white images for use in their augmentation model.
@@
 
-#### The variables of my use:
+### The variables of my use:
 
@@
 This ratio is indicative of fat distribution, and a higher WHR can be a precursing indicator of obesity [@reference].
-In the study done in [place], it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
+In the study done in \[place\], it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
 Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included.
 These measurements are particularly important for determining boundary values for fat classification [@reference].
-Excess fat accumulation at the wrists and ankles could indicate other health risks as seen in [].
+Excess fat accumulation at the wrists and ankles could indicate other health risks as seen in \[\].
 
-#### The estimand for the models
+### The estimand for the model
 
@@
 For BMI, I chose to consider the categories, `underweight`, `normal`, `overweight` and `obese`.
-These were calulated using the BMI formula:
-[insert BMI formula on a new line].
+These were calulated using the BMI formula: \[insert BMI formula on a new line\].
 
 No difference between the genders was considered.
-Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, [ii] and [iii].
+Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, \[ii\] and \[iii\].
 However, due to the lack of data points for Obesity II and III (BMI X - X), I chose to categorize all levels of obesity within the same `obesity` category.
@@
 
-#### graph what i will be using in my model
+### Measurement of Data
 
-### Body Composition Data - Isaac Kuzmar Et Al.
-### Data Overview
+## Body Composition Data - Isaac Kuzmar Et Al.
 
-The dataset compiles the body measurements of subjects aged, 18 and 60, and specifically with a desire to lose weight and improve body image.
+The data set compiles the body measurements of subjects aged, 18 and 60, and specifically with a desire to lose weight and improve body image.
 The participants resided in Barranquilla, Colombia and consisted of 234 males and 111 females.
 Medical exclusions were made while recruiting the subjects, such as pregnant women or people with medical pacemakers.
-The bodily measurements such as fat mass in kilograms and fat free mas in kilograms were determined using the Tanita MC-780 [reference] body composition analyzer.
+The bodily measurements such as fat mass in kilograms and fat free mas in kilograms were determined using the Tanita MC-780 \[reference\] body composition analyzer.
 
@@
 The fat mass percentage is considered the main predictor for fat mass categories.
-As the body composition is more accuarate at measuring the actual amount of excess fat a person carries and can differentiate it from other weights such as fat free muscle mass [reference the variable]
-#### graph what i will be using in my model
+As the body composition is more accuarate at measuring the actual amount of excess fat a person carries and can differentiate it from other weights such as fat free muscle mass \[reference the variable\] \#### graph what i will be using in my model
+
+### The estimand for the model
 
@@
 ```
+
 -   talk about the crowding for men
-
-## what original BMI and WHR fail to capture
-
-## How are these data sets important are how they relate to each other
-#### desitination
-#### well as any relationships between the variables.
 
@@
 
-## Measurement of Data
-
-### Measurements Data Set
-
-### Body Composition Data Set
-
 ```{r}
@@
 
+### Measurement of Data
+
 \newpage
+
 # Models
@@
 ## Measurements Model
+
 I aimed to provide a categorization of body fat levels in an individual based on simple measurements.
@@
 <!-- my model eq -->
+
+```{=tex}
 \begin{align}
-\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{baseline} \mid X)} \right) = \beta_{0k} + \beta_{1k} \cdot \text{height} + \beta_{2k} \cdot \text{gender} + \beta_{3k} \cdot \text{waist\_hip\_ratio} + \beta_{4k} \cdot \text{height\_hip\_ratio}
+\log \left( \frac{P(Y = k \mid X)}{P(Y = \text{Normal} \mid X)} \right) = \beta_{0k} + \beta_{1k} \cdot \text{height} + \beta_{2k} \cdot \text{gender} + \beta_{3k} \cdot \text{waist\_hip\_ratio} + \beta_{4k} \cdot \text{height\_hip\_ratio}
 \end{align}
+```
 
@@
 
-TODO !!!!!! change this to normal
-
-- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where \( k \in \{1, 2, \dots, K\} \)) given the predictors $X$.
-- $\text{baseline}$ is the reference category (the category of the outcome variable that is omitted in the model, typically the first category).
-- $\beta_{0k}$, $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the intercept and the predictors for category \( k \).
+-   $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where ( k \in {1, 2, \dots, K} )) given the predictors $X$.
+-   $\text{Normal}$ is the reference category
+-   $\beta_{0k}$, $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the intercept and the predictors for category ( k ).
 -   The predictors in this model are `height`, `gender`, `waist_hip_ratio`, and `height_hip_ratio`.
@@
 Hence, this model would not be able to differentiate between age groups and would work poorly if this information is included.
-I also considered estimating actual the fat percentage instead of fat percentage **category**.
+I also considered estimating actual the fat percentage instead of fat percentage *category*.
 However, to increase the comparability across both models, I chose to use a categorical variable instead.
@@
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 Where:
 
-- $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where (k \in \{\text{Low}, \text{High}, \text{Extreme) given the predictors $X$.
+-   $P(Y = k \mid X)$ is the probability of the response variable $Y$ being in category $k$ (where ($k$ $\in$ {$\text{Low}$, $\text{High}$, $\text{Extreme}$) given the predictors $X$.
 -   $\text{Normal}$ is the baseline category (the omitted category), meaning the model compares all other categories to "Normal".
-- $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the predictors for category \( k \), relative to the baseline category "Normal".
+-   $\beta_{1k}$, $\beta_{2k}$, $\beta_{3k}$, and $\beta_{4k}$ are the coefficients corresponding to the predictors for category ( k ), relative to the baseline category "Normal".
 -   The predictors in this model are `age`, `height`, `gender`, and `fm` (which stands for fat mass percentage).
@@
 Similar to the Body Measurements Model, this model also considers `fat_percentage_category` to be the estimand.
-According to the study [], the two sexes have different thresholds for different ages, suggesting linkages to different metabolic needs based on different phases of life.
+According to the study \[\], the two sexes have different thresholds for different ages, suggesting linkages to different metabolic needs based on different phases of life.
 Hence, `fat_percentage_category` is initially classified using `gender`, `age` and `fm`.
@@
 
-This model is also built on data that was generated by the ABS [] simulator so it may be inherently working with an error.
-While this is an indirect form of observed data, the original data set claims [x] accuracy.
-However, the error that it could cause in this paper is not accounted for and is considered the true measurement of the test subject.
+This model is also built on data that was generated by the ABS  simulator so it may be inherently working with an error.
+The error that it could cause in this paper is not accounted for and is considered the true measurement of the test subject.
 
 \newpage
+
 # Results
+
 Our results are summarized in @tbl-measurementmodelresults and @tbl-bodymass-modelresults. Our model's results and interpretation:
@@
 #| label: tbl-measurementmodelresults
-#| tbl-cap: "TODO"
+#| tbl-cap: "Summary of Measurements Model"
 
@@
 # Create and style the table
-kable(measurements_summary_df, caption = "Summary of Measurements Model")
+kable(measurements_summary_df)
 ```
-From the table {} we can see that the intercepts are very low and negative.
-The probability of the model making a classification is very unlikely with the catgegory with the highest probability being high.
-As our data is composed of people with a desire to improve body image, this makes sense, as the people with higher fat percentages are more likely to participate in such a study.
+
+From @tbl-measurementmodelresults we can see that the intercept for the Low category is negative.
+Without considering any additional predictors, the model has the highest probability of classifying a person in the Extreme category as its interest is high and positive.
+As all classifications are considered against the Normal category, the model is most likely to classify a person as having Extreme fat levels and very unlikely to classify them as Low.
+This data set has poor representation of underweight people and as it is composed of people with a desire to improve body image, this makes sense, as the people with higher fat percentages are more likely to participate in such a study.
 `waist_hip_ratio` has a negative correlation with the levels of the fat percentage category prediction.
@@
 There is a negative correlation with the heights of the subjects suggesting that taller people are less likely to be classified with higher levels of fat.
-Detailed interpretation of this fact is later discussed in the Results Section {}.
+Detailed interpretation of this fact is later discussed in the [Results Section](#results-height-point).
 
@@
 #| label: tbl-bodymass-modelresults
-#| tbl-cap: "TODO"
+#| tbl-cap: "Summary of Body Mass Model"
 
@@
 # Create and style the table
-kable(body_mass_summary_df, caption = "Summary of Body Mass Model")
+kable(body_mass_summary_df)
 ```
 
-The negative intercepts indicate that in the absence of any predictors, the model has very low estimates for the body fat classifications.
-However, the model is most likeyly to classify a person in the `High` category owing to the distribution of the dataset.
-The positive intercept for a `Normal` classification mean that the taller a person is, the more likely they are to be classified as `Normal`.
-As this model is directly based on the actual body composition of a person, the increasing levels of fat percentage indicated by `fm` show that the higher percentage of fat that a person has, the more likely they would be higher in the classification levels.
+For the low category relative to Normal, age and height did not have a significant influence despite initial speculation.
+Both variables have the $\beta$ coefficients of $\beta=1.889$ and $\beta=-0.129$ respectively.
+However, gender had significant influence on the log-odds of being classified into this category.
+As males and females are regarded as 1 and 2 in the binary category, `gender`, the $\beta = 69.183$ coefficient indicates that males are more likely to fall into the low category.
+Naturally, there exists a negative coefficient between fat mass percentage with $\beta= -9.494$
+
+For the High and Extreme categories, age and height had weaker influences.
+However, gender has a lower log-odd ($\beta=-43.068$ and $\beta=-83.059$ respectively) .Suggesting a negative association between the binary `gender` variable.
+Hence, females have a higher influence of being classified into these categories over males.
+
+These model results highlight the nuanced relationships between the adipose and anthropocentric measurements. For the Body Composition Model, `gender` and `fm` exhibit the highest influence.
 
@@
 #| message: false
+#| label: fig-density-compare
+#| fig-cap: "Density Plot of Predicted Categories by Source"
 
-# measurements_results <- cleaned_data_model %>%
-#   select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
-#  mutate(source = "Measurements")
-# 
-# body_composition_results <- body_mass_model %>%
-#   select(gender, height, weight, height_category, fat_percentage_category, predicted_category) %>%
-# mutate(source = "Body Composition")
-# 
-# elongated_data <- bind_rows(measurements_results, body_composition_results)
-# 
-# ggplot(elongated_data, aes(x = predicted_category, fill = source)) +
-#   geom_density(alpha = 0.5) +
-#   facet_wrap(~ predicted_category) +
-#   labs(title = "Density Plot of Predicted Categories by Source",
-#        x = "Predicted Category",
-#        y = "Density",
-#        fill = "Source") +
-#   theme_minimal()
-```
-From the above prediction density chart by each model per fat percentage category, we see that the Measurements model has higher peaks at the each classification.
-This suggests that more people are being classified into a similar category only based on their physical traits such as height and Waist-to-Hip ratio.
-The peaks of each category from the Body Composition is accurate in identifying the major classification.
-However, the distribution not being as concentrated as from the Measurements model suggest that the diversification of different body types is being identified properly.
-
-```{r chunk-name}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
 levels_results <- union(levels(cleaned_data_model$predicted_category), levels(body_mass_model$predicted_category))
@@
 
-# Calculate the percentage of each predicted category
-category_percentage_by_source <- elongated_data %>%
-  group_by(source, predicted_category) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100) %>%
-  select(source, predicted_category, percentage)
+ggplot(elongated_data, aes(x = predicted_category, fill = source)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(~ predicted_category) +
+  labs(x = "Predicted Category",
+       y = "Density",
+       fill = "Source") +
+  theme_minimal()
+```
 
-category_percentage_by_source_height <- elongated_data %>%
-  group_by(source, height_category, predicted_category) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100) %>%
-  select(source, height_category, predicted_category, percentage)
+From [the prediction density chart](@fig-density-compare) by each model per fat percentage category, we see that the Measurements model has higher peaks at the each classification.
+This suggests that more people are being classified into a similar category only based on their physical traits such as height and Waist-to-Hip ratio.
+The peaks of each category from the Body Composition is accurate in identifying the major classification.
+However, the distribution not being as concentrated as from the Measurements model suggest that the diversification of different body types is being identified properly.
 
-# Print the result
-print(category_percentage_by_source)
-
-ggplot(category_percentage_by_source, aes(x = predicted_category, y = percentage, fill = source)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Percentage of Predicted Categories by Source",
-       x = "Predicted Category",
-       y = "Percentage (%)") +
-  theme_minimal() +
-  scale_fill_manual(values = c("Measurements" = "skyblue", "BodyComposition" = "grey")) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels if needed
-```
-## The Body Composition Model has diverse classifcation for different heights groups
 ```{r}
@@
 #| message: false
+#| label: fig-conf-matrix
+#| tbl-cap: "Confusion Matrix for Both Sources (Proportions)"
+levels_results <- union(levels(cleaned_data_model$predicted_category), levels(body_mass_model$predicted_category))
+
+measurements_results <- cleaned_data_model %>%
+  select(height, weight, height_category, fat_percentage_category, predicted_category) %>%
+  mutate(source = "Measurements")
+
+body_composition_results <- body_mass_model %>%
+  select(height, weight, height_category, fat_percentage_category, predicted_category) %>%
+ mutate(source = "BodyComposition")
+
+elongated_data <- bind_rows(measurements_results, body_composition_results)
+
+# Combine both sources into one dataset
+combined_results <- bind_rows(
+  measurements_results %>% mutate(source = "Measurements"),
+  body_composition_results %>% mutate(source = "BodyComposition")
+)
+
+# Create the confusion matrix and calculate proportions
+combined_confusion_matrix <- combined_results %>%
+  group_by(source, predicted_category, fat_percentage_category) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(source) %>%
+  mutate(proportion = count / sum(count)) %>%  # Calculate proportions
+  ungroup()
+
+# Visualize the confusion matrices as faceted heatmaps based on proportions
+ggplot(combined_confusion_matrix, aes(x = predicted_category, y = fat_percentage_category, fill = proportion)) +
+  geom_tile(color = "black") +  # Tiles for the matrix
+  scale_fill_gradient(low = "white", high = "blue") +  # Color gradient for proportion values
+  labs(x = "Predicted Category", 
+       y = "Fat Percentage Category", 
+       fill = "Proportion") +
+  facet_wrap(~ source, scales = "free") +  # Create separate heatmaps for each source
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
+
+```
+
+The [Confusion Matrix](@fig-conf-matrix) shows the difference in the prediction of the fat percentage category by each model.
+Here, we can see that the Body Composition Data is mostly consistent with its prediction apart from a few key nuances.
+The diagonal values in the Body Compositions show high correlation between the body fat percentage and the classification.
+This is expected as this model serves as a comparison to the Measurements Model.
+However, it can be noted that there is a spill between Moderate and High levels of body fat.
+The model predicts the Extreme category the closest.
+On the other hand, there is more obvious spill over between categories using the measurements model.
+The diagonal values have lighter colours based on the Confusion Matrix Colour Legend.
+This could suggest that the introduction of adipose and anthropocentric differs in predicting fat in the body as compared to its proxy, BMI.
+The same hate map [faceted by gender](@fig-heat-map-gender), highlights that the BMI proxy in Measurements Model seems to be under performing for females causing more classifications.
+This could be alluded from the fact that BMI was created with only male subjects under consideration \[todo reference\].
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-heat-map-gender
+#| tbl-cap: "Heatmap of Predicted Categories by Source and Gender"
+measurements_results_gender <- cleaned_data_model %>%
+  select(gender, fat_percentage_category, predicted_category) %>%
+  mutate(source = "Measurements")
+
+body_composition_results_gender <- body_mass_model %>%
+  select(gender, fat_percentage_category, predicted_category) %>%
+ mutate(source = "BodyComposition")
+
+measurements_results_gender <- measurements_results_gender %>%
+  mutate(gender = recode(gender, "1" = "Male", "2" = "Female"))
+
+body_composition_results_gender <- body_composition_results_gender %>%
+  mutate(gender = recode(as.character(gender), "1" = "Male", "2" = "Female"))
+
+# Combine the datasets
+combined_results_gender <- bind_rows(measurements_results_gender, body_composition_results_gender)
+
+proportion_data_gender <- combined_results_gender %>%
+  group_by(source, gender, fat_percentage_category) %>%
+  count(predicted_category) %>%
+  mutate(proportion = n / sum(n) * 100)
+
+  ggplot(proportion_data_gender, aes(x = fat_percentage_category, y = predicted_category, fill = proportion)) +
+  geom_tile() +
+  facet_grid(gender ~ source) +
+  scale_fill_gradient(low = "white", high = "blue") +
+  labs(
+    x = "Actual Fat Percentage Category",
+    y = "Predicted Category",
+    fill = "Proportion (%)"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    strip.text = element_text(size = 12, face = "bold")
+  )
+```
+
+## What BMI fails to capture
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-what-bmi-doesnt-show
+#| tbl-cap: "Weight Distribution by Predicted Category, Source, and BMI Category"
+
+# Assuming measurements_data and body_data have the same structure
+measurements_long <- cleaned_data_model %>%
+select(weight, predicted_category, bmi_category) %>%
+  mutate(source = "Measurements")
+
+body_data_long <- body_mass_model %>%
+  select(weight, predicted_category, bmi_category) %>%
+  mutate(source = "Body Composition")
+
+# Combine both datasets into one for easy plotting
+combined_data <- bind_rows(measurements_long, body_data_long) %>%
+ mutate(predicted_category = factor(predicted_category,
+                                          levels = c("Low", "Moderate", "High", "Extreme"))) %>%
+  mutate(bmi_category = factor(bmi_category,
+                                          levels = c("Underweight", "Normal", "Overweight", "Obese")))
+
+ggplot(combined_data, aes(x = predicted_category, y = weight, fill = source)) +
+  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
+  geom_jitter(width = 0.2, alpha = 0.4, aes(color = source)) +
+  facet_grid(~ bmi_category) +
+  labs(
+    x = "Predicted Category",
+    y = "Weight",
+    fill = "Source",
+    color = "Source"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    strip.text = element_text(size = 10, face = "bold")
+  )
+
+```
+
+From @fig-what-bmi-doesnt-show, we see that for a wider range in weight, there is a higher proportion of people who are classifed in the same BMI category.
+From the Body Composition Model, these categories like on the lower range of the BMI weight thresholds.
+Increase in fat weight moves people between Fat Classification but they remain in the same BMI category for a longer range of their weight.
+This shows the lack of factors involved in the BMI classification.
+Additionally,
+
+## Body Fat classifcation for different heights classes {#results-height-point}
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-height-graphs
+#| tbl-cap: "Proportion of Predicted Categories by Height Category and Source"
+#| 
 # Calculate proportions for each combination of predicted_category, height_category, and source
@@
    facet_wrap(~ source) +
-  labs(title = "Proportion of Predicted Categories by Height Category and Source",
-       x = "Height Category",
+   labs(x = "Height Category",
         y = "Percentage (%)",
@@
 
-The primary shortfall of the BMI categories was that it failed to differententiate the body types based on other factors such as where the fat was collected.
-From the above graph, we can see that across different heights, the Measurements Model is pretty consistent with providing similar proportions for different fat categories.
-This suggest that it was group people on too broad categories that did not necessarily capture the right effect of height of body compostion.
-Specifically, BMI fails to recognize the vertical spread of fat accumulation and so taller people get naturally classified as having lower fat levels.
-However, the body composition model, has a better ability a higher percentage of people actually have a higher classification of high and extreme body fat across all height groups.
+Building on the previous point of BMI failing to correctly capture physical attributes of body composition, this graph emphasises how the Moderate Category is persistently overclassiffied across height groups.
+In the Measurements Model graph, all height groups share similar proportions of Fat Category classifications.
+More than 50% of the people below 170CM and slightly less than 50% people taller than 170CM are classified as Normal Fat Category.
+However, this overclassication is debunked by the Body Composition Model Graph as almost all of the height classes have a major classification in the Normal category.
+In fact, height classes above 160CM have a collective proportion of greater than 75% for High and Extreme fat levels.
+As the Body Composition Model specifically disregards weight holistically and considers pure fat only, we can see that tall people have an increased chance of being classified as Normal fat category.
+Subsequent health risk identifications are delayed for these groups of people and is later discussed in greater detail \[here todo.\] Additionally, following the original BMI formula, taller people are at advantage as their weight is spread further along their height.
+However, the difference in proportions of the classifications across both models highlights this specific downfall of the BMI formula.
 
-## Differentiating between different levels of Obeseity.
+## Difference in Gender Performance
 
-## What BMI fails to capture
 ```{r}
+#| eval: true
 #| echo: false
-#| eval: true
+#| message: false
 #| warning: false
-#| message: false
+#| label: fig-gender-point
+#| tbl-cap: "Comparison of Predicted Categories by Source for Each Gender"
 
-  # Function to categorize weight
+proportion_data_2 <- proportion_data_gender %>%
+ mutate(predicted_category = factor(predicted_category,
+                                          levels = c("Low", "Moderate", "High", "Extreme")))
 
-cleaned_data_model <- cleaned_data_model %>%
-  mutate(weight_category = cut(weight, 
-                               breaks = seq(40, 150, by = 10), # Specify your range
-                               labels = c("40-50", "50-60", "60-70", "70-80", "80-90", "90-100", "100-110", "110-120", "120-130", "130-140", "140-150"),
-                               include.lowest = TRUE))
-
-# Apply the same logic for body_data
-body_mass_model <- body_mass_model %>%
-  mutate(weight_category = cut(weight, 
-                               breaks = seq(40, 150, by = 10), # Specify your range
-                               labels = c("40-50", "50-60", "60-70", "70-80", "80-90", "90-100", "100-110", "110-120", "120-130", "130-140", "140-150"),
-                               include.lowest = TRUE))
-
-# Assuming measurements_data and body_data have the same structure
-measurements_long <- cleaned_data_model %>%
-  select(weight_category, predicted_category, bmi_category) %>%
-  pivot_longer(cols = c(predicted_category, bmi_category),
-               names_to = "category_type",
-               values_to = "category") %>%
-  mutate(source = "Measurements")
-
-body_data_long <- body_mass_model %>%
-  select(weight_category, predicted_category, bmi_category) %>%
-  pivot_longer(cols = c(predicted_category, bmi_category),
-               names_to = "category_type",
-               values_to = "category") %>%
-  mutate(source = "Body Composition")
-
-# Combine both datasets into one for easy plotting
-combined_data <- bind_rows(measurements_long, body_data_long)
-
-  ggplot(combined_data, aes(x = weight_category, fill = category)) +
-  geom_bar(position = "fill", alpha = 0.6) +
-  facet_grid(category_type ~ source, scales = "free_y") +  # Facet by category_type and source
-  labs(title = "Proportions of Categories by Weight Range and Dataset",
-       x = "Weight Range (kg)",
-       y = "Proportion",
-       fill = "Category") +
+#Visualize with a grouped bar chart showing percentages for both sources in the same graph
+ggplot(proportion_data_2, aes(x = predicted_category, y = proportion, fill = source)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  facet_wrap(~ gender, scales = "free", ncol = 2) +
+  labs(
+    x = "Predicted Percentage Category",
+    y = "Proportion (%)",
+    fill = "Source"
+  ) +
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    strip.text = element_text(size = 12, face = "bold")
+  )
 
@@
 
+@fig-gender-point shows the performance of each of the models against the `gender` predictor.
+For females, the proporitons for all categories are higher in the Body Composition Model suggesting that the model is able to correctly classify the fat percentage category based on the predictors `gender` and `fm`.
+However, the Measurements Model has lower classifications suggesting that there is different classification based on the BMI proxy and the predictors `gender`, `waist_hip_ratio` and `height_hip_ratio.`
+
 # Discussion
 
-## Obesity Risk Identification {#sec-first-point}
+## Potential Health Risk Identification {#sec-first-point}
 
 This paper's main goal was to compare how accurate are body measurements to estimate fat percentage at home.
-Other ways of estimating a person's body fat include scans like DEXA {source}.
+Other ways of estimating a person's body fat include scans like DEXA (Kuzmar2020).
 These scans are expensive and inaccessible.
 Most people turn to measuring BMI's to estimate their general level of health.
-The Government of Cananda also advices that adults use BMI to estimate their health [].
+The Government of Canada (@CanadaBMI2024) also advices that adults use BMI to estimate their health.
 As seen in this paper, this inherently flawed as BMI fails to capture the composition of a body.
@@
 Additionally, as a higher proportion of people are classified as Normal using BMI, without failing to recognise that having a high level of fat despite having a lower BMI is still dangerous as it could potentially prevent them from seeking early care against illnesses like insulin resistance.
-Insulin resistance occurs when there is a body fat percentage of {}.
-By popularising a more accurate system, we could urge people to take earlier preventive care.
+By popularizing a more accurate system, we could urge people to take earlier preventive care.
+This opens up the path for the need for more sophisticated medical interventions.
+According to @Adabk1274, ratios such WHR are beneficial to identify cardiovascular health risks due to adipose tissue fat accumulation.
+However, due to the lack of standardization in current medical care, this ratio is yet to gain popularity over the BMI formula.
 
-## Second discussion point
+It also maybe understood that WHR shows a better performance for certain demographics only.
+In the study done by @Li2006, the WHR showed no increased association with cardiovascular diseases in Overweight and Obese men.
+However, for other demographics such as women, there is parallel increase in the risk of cardiovascular diseases with increase in WHR given the same BMI category.
+Nearly a double association with cardiovascular diseases was seen in Normal and Overweight women who had a higher WHR than women who had a lower WHR.
+This relation could applied to Normal BMI only for men.
+As the WHR is non-standard and of value to a smaller demographic of the general population, a wide clinical acceptance has not been seen for it.
+
+## What should athletes use?
+
+BMI is a crude way of classifying people. Athletes are especially at risk at being mis-classified despite a lower body mass weight but a higher muscle mass. According to \[\], Athletes have
 
@@
 
-## Second discussion point {#sec-second-point}
+## Weaknesses and next steps {#sec-weaknesses-point}
 
-## Third discussion point {#sec-third-point}
-
-## Weaknesses and next steps {#sec-weaknesses-point}
-- the BodyM dataset does not include age. An older person has lower metabolism [] and a higher chance of collecting fat on their abdomen.
+-   the BodyM dataset does not include age. An older person has lower metabolism \[\] and a higher chance of collecting fat on their abdomen.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 
-# ```{r}
-# modelplot(political_preferences, conf_level = 0.9) +
-#   labs(x = "90 per cent credibility interval")
-# ```
-
 \newpage

### 2024-12-15 scripts/02b-data_cleaning.R

@@ -66,7 +66,7 @@ body_mass_clean <- body_mass_data_raw %>%
   mutate(
     bmi_category = case_when(
       bmi < 18.5 ~ "Underweight",
-      bmi >= 18.5 & bmi <= 24.9 ~ "Moderate",
+      bmi >= 18.5 & bmi <= 24.9 ~ "Normal",
       bmi >= 25.0 & bmi <= 29.9 ~ "Overweight",
       bmi >= 30.0 ~ "Obese"
     )) %>%

### 2024-12-15 scripts/04-model.R

@@ -26,7 +26,7 @@ cleaned_data_model$predicted_category <- predict(measurements_model, newdata = c
 body_mass_model_bmi <- multinom(fat_percentage_category ~ age + height + gender + fm, data = body_mass_model)
 body_mass_model$predicted_category <- predict(body_mass_model_bmi, newdata = body_mass_model)
 
-summary(measurements_model)
+summary(body_mass_model_bmi)
 
 #### Save model ####
 write_parquet(cleaned_data_model, "data/analysis_data/measurements_analysis_data.parquet")

### 2024-12-15 paper/paper.qmd
sections 1, 2, 3 and 4 completed

@@
 More information on packages and specific methods is included later in the paper.
-This paper is structured as follows: The \[Data Section\](@paper-data-section) introduces the data sets, their variables and how they were tailored to obtain the needed information.
+This paper is structured as follows: The Data Section introduces the data sets, their variables and how they were tailored to obtain the needed information.
 The Model section discusses in greater detail the structure of both models and what information they convey.
@@
 [@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks.
-The primary data captured in this collection consists of the front and lateral silhouettes of approximately \[2000\] test subjects.
+The primary data captured in this collection consists of the front and lateral silhouettes of approximately \[TODO\] test subjects.
 These silhouettes were then converted into black-and-white images for use in their augmentation model.
-The data set includes X male and Y female subjects, aged A to B.
-The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
-The S3 package [@reference] included three datasets: Training, Test A, and Test B.
+The data set includes TODO male and TODO female subjects, aged TODO to TODO The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
+The S3 package  included three datasets: Training, Test A, and Test B.
 For the purposes of this paper, only the Training data set is used.
@@
 However, these predictors are not effective at differentiating body types.
-Therefore, to effectively construct the distribution of body fat in human body only through measurements, I created two variables called `waist_hip_ration` and `height_hip_ratio`.
+Therefore, to effectively construct the distribution of body fat in human body only through measurements, I created two variables called `waist_hip_ratio` and `height_hip_ratio`.
 These variables are used to model the differences in the lateral and horizontal fat distributions.
-According to WHO [@reference], the waist-to-hip ratio, WHR, is a more accurate predictor of body fat around the abdominal area.
-Based on the study done in [@reference], the two sexes exhitbit difference tendenices for excess body fat accumulation, also known as the adipose fat.
+According to WHO [@WHO2008], the waist-to-hip ratio, WHR, is a more accurate predictor of body fat around the abdominal area.
+Based on the study done in [@Wells2012], the two sexes exhibit difference tendencies for excess body fat accumulation, also known as the adipose fat.
 In males, excess fat tends to accumulate in the abdominal region, whereas in females, fat is more commonly stored in the hips.
-This ratio is indicative of fat distribution, and a higher WHR can be a precursing indicator of obesity [@reference].
-In the study done in \[place\], it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
+This ratio is indicative of fat distribution, and a higher WHR can be a precursing indicator of obesity.
+In the study, it is shown that an increase in height lowers the decreases the probability of being classified into a higher BMI category for both genders.
 Finally, to enhance the model’s quality, ankle and wrist circumference measurements were also included.
-These measurements are particularly important for determining boundary values for fat classification [@reference].
-Excess fat accumulation at the wrists and ankles could indicate other health risks as seen in \[\].
+These measurements are particularly important for determining boundary values for fat classification [@Mitchell1993].
+Excess fat accumulation at the wrists and ankles could indicate other health risks.
 
-### The estimand for the model
-
-Using the above predictor variables, I contructed the variable `fat_percentage_category` as I recruit a multilevel linear regression model to classify the body fat.
-Initially, I wanted to predict the actual fat percentage, but this proved difficult as ...
-I decided to use a categorical variable as the estimand to produced useful comparisons against other categorical variables such as sex, height categories and weight categories.
-Additionally, `fat_percentage_category` was constructed using BMI as a proxy to gauge fat levels in a person.
-BMI is still widely used in medical scenarios to provide general diagnosis of health levels so I used that as a base to study the interactions between current systems and new offerings of WHR and height_hip_ratio.
-The relationship between `fat_percentage_category` and BMI was considered to be relatively straightforward to mimick the nature of the current applications of BMI.
-For BMI, I chose to consider the categories, `underweight`, `normal`, `overweight` and `obese`.
-These were calulated using the BMI formula: \[insert BMI formula on a new line\].
-
-No difference between the genders was considered.
-Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, \[ii\] and \[iii\].
-However, due to the lack of data points for Obesity II and III (BMI X - X), I chose to categorize all levels of obesity within the same `obesity` category.
-
-The following table shows the relationship between the new `fat_percentage_category` against existing the `bmi_category`.
+The figures: @fig-prop-bmi-gender, @fig-prop-whr-gender and @fig-prop-hhr-gender show the graphs for the dataset's proportion for BMI for each gender as well as the Ankle and Wrist Measurements in @fig-prop-ankle-wrist.
 
@@
 #| message: false
+#| label: fig-prop-bmi-gender
+#| fig-cap: "Proportions of BMI Categories by Gender"
 
@@
 ggplot(cleaned_data_model, aes(x = gender, fill = bmi_category)) +
-  geom_bar(position = "fill") +  # 'fill' makes it a proportion plot
+  geom_bar(position = "fill") +  
   ylab("Proportion") +
   xlab("Gender") +
-  labs(title = "Proportions of BMI Categories by Gender") +
-  scale_y_continuous(labels = scales::percent) +  # Converts y-axis to percentage format
+  scale_y_continuous(labels = scales::percent) +
   theme_minimal()
@@
 #| message: false
+#| label: fig-prop-whr-gender
+#| fig-cap: "Distribution of Waist-to-Hip Ratio by Gender and Proxy Fat Percentage"
+
 cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
@@
   xlab("Gender") +
-  labs(title = "Distribution of Waist-to-Hip Ratio by Gender") +
   theme_minimal() +
   facet_wrap(~fat_percentage_category) + 
-  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
+  scale_fill_manual(values = c("Female" = "lightblue", "Male" = "lightgreen"))
 ```
@@
 #| message: false
-# cleaned_data_model$gender <- factor(cleaned_data_model$gender, labels = c("Female", "Male"))
+#| label: fig-prop-hhr-gender
+#| fig-cap: "Distribution of Height-Hip-Ratio by Gender and Proxy BMI"
 
@@
   facet_wrap(~bmi_category) + 
-  scale_fill_manual(values = c("female" = "lightblue", "male" = "lightgreen"))
+  scale_fill_manual(values = c("Female" = "lightblue", "Male" = "lightgreen"))
 ```
+
+### The estimand for the model
+
+Using the above predictor variables, I constructed the variable `fat_percentage_category` as I recruit a multilevel linear regression model to classify the body fat.
+I decided to use a categorical variable as the estimand to produced useful comparisons against other categorical variables such as sex, height categories and weight categories.
+Additionally, `fat_percentage_category` was constructed using BMI as a proxy to gauge fat levels in a person.
+BMI is still widely used in medical scenarios to provide general diagnosis of health levels so I used that as a base to study the interactions between current systems and new offerings of WHR and height_hip_ratio.
+The relationship between `fat_percentage_category` and BMI was considered to be relatively straightforward to mimick the nature of the current applications of BMI.
+For BMI, I chose to consider the categories, `underweight`, `normal`, `overweight` and `obese`.
+These were calculated using the BMI formula:
+
+\begin{align}
+\text{BMI} = \frac{\text{weight (kg)}}{\left(\frac{\text{height (cm)}}{100}\right)^2}
+\end{align}
+
+No difference between the genders was considered.
+Obesity categories using BMI consists of Obese I, Obese II and Obese III as they cause the increasing obesity relating such as diabetes, fatty liver and Osteoarthritis [@NIDDK2024].
+However, due to the lack of data points for Obesity II and III, I chose to categorize all levels of obesity within the same `obesity` category.
 
@@
 #| message: false
-# filtered_data <- subset(cleaned_data_model, bmi_category %in% c("Overweight", "Obese"))
-
-# Create a combined boxplot to visualize differences in ankle and wrist for bmi_category = Overweight and Obese
+#| label: fig-prop-ankle-wrist
+#| fig-cap: "Comparison of Ankle and Wrist Measurements for BMI Categories"
 
@@
   xlab("BMI Category") +
-  labs(title = "Comparison of Ankle and Wrist Measurements for Overweight and Obese BMI Categories") +
   theme_minimal() +
-  scale_fill_manual(values = c("Overweight" = "lightcoral", "Obese" = "lightsalmon")) +
   theme(legend.position = "none")
@@
 ### Measurement of Data
+
+From the 3 data sets, I used the training data set in which the subjects were 3D scanned and photographed by lab technicians.
+These photographs where then put through the ABS simulator for the orginal purposes of this study.
+All test subjects were required to wear tight body-fitting clothing and asked to stand in an "A" pose to maintain uniformity.
 
@@
 Medical exclusions were made while recruiting the subjects, such as pregnant women or people with medical pacemakers.
-The bodily measurements such as fat mass in kilograms and fat free mas in kilograms were determined using the Tanita MC-780 \[reference\] body composition analyzer.
+The bodily measurements such as fat mass in kilograms and fat free mas in kilograms were determined using the Tanita MC-780 [@Kuzmar2020] body composition analyzer.
 
@@
 
-The main variables that I considered fat mass percentage fm.
+The main variables that I considered fat mass percentage `fm`.
 This variable is obtained by dividing the fat mass in kilogram by the total weight in kilograms.
 The fat mass percentage is considered the main predictor for fat mass categories.
-As the body composition is more accuarate at measuring the actual amount of excess fat a person carries and can differentiate it from other weights such as fat free muscle mass \[reference the variable\] \#### graph what i will be using in my model
+As the body composition is more accurate at measuring the actual amount of excess fat a person carries and can differentiate it from other weights such as fat free muscle mass and total body weight.
+`age`, `height` and `gender` are also considered as predictor variables.
+As a person ages, their fat storing capabilities as well as their metabolism changes.
+While this cannot be directly measured through the data set, I used age and gender of a person to define the inital classification of the independant variable.
 
 ### The estimand for the model
+
+To make this model comparable to the Measurements Model, `fat_percentage_category` is the estimand as well.
+I considered a different proxy for `fat_percentage_category`.
+Afterwards, after fitting other variables such `age`, `height` and `gender`, the model predicted new categories.
 
@@
 #| message: false
-# Recode gender levels
+#| label: fig-bmi-2-gender
+#| fig-cap: "Proportions of BMI Categories by Gender"
+#| 
 body_mass_model_data_section <- body_mass_model %>%
@@
   xlab("Gender") +
-  labs(title = "Proportions of BMI Categories by Gender") +
   scale_y_continuous(labels = scales::percent) +  # Converts y-axis to percentage format
@@
 #| message: false
-ggplot(body_mass_model, aes(x = weight, y = height, color = fm)) +
-  geom_point(size = 3) +
-  scale_color_viridis_c() + # Use a continuous color scale
-  labs(title = "Height vs Weight with Fat Mass",
-       x = "Weight",
-       y = "Height",
-       color = "Fat Mass (fm)") +
-  theme_minimal()
-```
+#| label: fig-prop-fm-ht-wt
+#| fig-cap: "Relationship between Fat Mass and Height vs Weight"
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
 ggplot(body_mass_model, aes(x = weight, y = height, color = fm, size = fm)) +
@@
   scale_size_continuous(range = c(1, 10)) +
-  labs(title = "Height vs Weight with Fat Mass",
+  labs(
        x = "Weight",
@@
 #| message: false
-ggplot(body_mass_model, aes(x = bmi, y = fm, color = gender)) +
-  geom_point(alpha = 0.7) +
-  facet_wrap(~ gender) +
-  labs(title = "Fat Mass vs BMI by Gender",
-       x = "BMI",
-       y = "Fat Mass (fm)",
-       color = "Gender") +
-  theme_minimal()
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+#| label: fig-prop-smoothed-bmi
+#| fig-cap: "Smoothed Trends of FM and BMI by Gender"
 body_mass_long <- body_mass_model %>%
@@
   geom_smooth(se = FALSE, method = "loess", span = 0.5) + # Smoothed trendline
-  labs(title = "Smoothed Trends of FM and BMI by Age",
-       x = "weight",
-       y = "Value",
+  labs(
+       x = "Weight in kg",
+       y = "BMI",
        color = "Measure") +
@@
 
--   talk about the crowding for men
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-# Combine both datasets
-cleaned_test_extract <- cleaned_data_model %>%
-  select(gender, bmi_category, fat_percentage_category)
-body_mass_extracted <- body_mass_model %>%
-  select(gender, bmi_category, fat_percentage_category)
-cleaned_test_extract$dataset <- "Measurements"
-body_mass_extracted$dataset <- "Body Data"
-combined_data <- rbind(cleaned_test_extract, body_mass_extracted)
-
-# Count occurrences of each fat_percentage_category per dataset
-library(dplyr)
-category_counts <- combined_data %>%
-  group_by(dataset, fat_percentage_category) %>%
-  summarise(Count = n(), .groups = "drop")
-
-# Create a bar plot to visualize the differences
-library(ggplot2)
-ggplot(category_counts, aes(x = fat_percentage_category, y = Count, fill = dataset)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Comparison of Fat Percentage Categories Across Data1 and Data2",
-       x = "Fat Percentage Category", y = "Count") +
-  theme_minimal() +
-  scale_fill_manual(values = c("Measurements" = "lightblue", "Body Data" = "lightgreen"))
-
-proportions <- combined_data %>%
-  group_by(dataset, fat_percentage_category) %>%
-  summarise(Count = n(), .groups = "drop") %>%
-  mutate(Proportion = Count / sum(Count))
-
-# Create a bar plot to visualize the proportions
-ggplot(proportions, aes(x = fat_percentage_category, y = Proportion, fill = dataset)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Proportional Comparison of Fat Percentage Categories Across Data1 and Data2",
-       x = "Fat Percentage Category", y = "Proportion") +
-  scale_y_continuous(labels = scales::percent) +  # Show proportions as percentages
-  theme_minimal() +
-  scale_fill_manual(values = c("Measurements" = "lightblue", "Body Data" = "lightgreen"))
-```
-
-Overview text
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-```
+The figures: @fig-bmi-2-gender, @fig-prop-fm-ht-wt and @fig-prop-smoothed-bmi graph the 4 predictor variables of my interest for the subsequent model.
+@fig-prop-smoothed-bmi is a motivating factor in recognizing that gender plays an important role in conducting analysis for Body Fat with relevance to current clinical diagnostic tools such as BMI.
 
 ### Measurement of Data
+Height in centimeters and fat mass values were obtained using the Tanita MC-780MA medical equipment.
 
@@
 The same hate map [faceted by gender](@fig-heat-map-gender), highlights that the BMI proxy in Measurements Model seems to be under performing for females causing more classifications.
-This could be alluded from the fact that BMI was created with only male subjects under consideration \[todo reference\].
+This could be alluded from the fact that BMI was created with only male subjects under consideration [@Pray2023].
 

### 2024-12-15 paper/paper.qmd
final paper, sketches to be included

@@
 #| label: tbl-gov-can-bmi
-#| tbl-cap: "BMI vs. Fat Mass by BMI Category"
+#| tbl-cap: "BMI Categories as per the Government of Canada Guidelines"
 bmi_canada_gov <- data.frame(
@@
   bmi_canada_gov,
-  col.names = c("Category", "BMI Range"),
-  caption = "BMI Categories as per the Government of Canada Guidelines"
+  col.names = c("Category", "BMI Range")
 )
@@
 [@Ruiz2022] with a focus on underrepresented body types in the estimation of fat and its subsequent health risks.
-The primary data captured in this collection consists of the front and lateral silhouettes of approximately \[TODO\] test subjects.
+The primary data captured in this collection consists of the front and lateral silhouettes of 2,779 test subjects.
 These silhouettes were then converted into black-and-white images for use in their augmentation model.
-The data set includes TODO male and TODO female subjects, aged TODO to TODO The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
-The S3 package  included three datasets: Training, Test A, and Test B.
+The body measurements in this data set were generated using their Adversarial Body Simulator (ABS), which was specifically designed to capture underrepresented body types.
+The S3 package  included three data sets: Training, Test A, and Test B.
 For the purposes of this paper, only the Training data set is used.
@@
 ### Measurement of Data
+Medical records obtained through medical–nutritional consultations of BiomediKcal–Advanced Medical Nutrition & Lifestyle Center of Barranquilla were used to construct this data set.
 Height in centimeters and fat mass values were obtained using the Tanita MC-780MA medical equipment.
-
-\newpage
 
@@
 As the Body Composition Model specifically disregards weight holistically and considers pure fat only, we can see that tall people have an increased chance of being classified as Normal fat category.
-Subsequent health risk identifications are delayed for these groups of people and is later discussed in greater detail \[here todo.\] Additionally, following the original BMI formula, taller people are at advantage as their weight is spread further along their height.
+Subsequent health risk identifications are delayed for these groups of people and is later discussed in greater detail.
+Additionally, following the original BMI formula, taller people are at advantage as their weight is spread further along their height.
 However, the difference in proportions of the classifications across both models highlights this specific downfall of the BMI formula.
@@
 
-BMI is a crude way of classifying people. Athletes are especially at risk at being mis-classified despite a lower body mass weight but a higher muscle mass. According to \[\], Athletes have
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+BMI is a crude way of classifying people. Athletes are especially at risk at being mis-classified despite a lower body mass weight but a higher muscle mass.
 
@@
 
--   the BodyM dataset does not include age. An older person has lower metabolism \[\] and a higher chance of collecting fat on their abdomen.
+Neither of the dataset include children and cannot be applied to anyone under the age of 18.
+It is important to consider age as a primary factor while diagnosing any medical condition.
+Additionally, fat deposits affect different age demographics differently.
+For example, an obese child can develop early symptoms of adult obesity if not corrected in the younger years.
+Likewise, menopausal women face hormonal changes that cause increased abdominal obesity (@Davis2012).
+Information such as pregnancy status should also be considered if WHR is to be introduced as a replacement for BMI in clinical settings.
 
-Weaknesses and next steps should also be included.
+My models are recorded for ethically different groups of people.
+While the AWS BodyM Data Set is collected for a varied group of ethnicity, the Body Composition data set is specifically collected only in Barranquilla, Columbia.
+The way humans across different races store and respond to excess fat is also largely different [@Deurenberg1998].
+Hence, the models should have accounted for potential confounders that could arise due to enthicity.
+Furthermore, the models does not consider the difference in data sizes.
+While I did use proportions as much as possible, the data sets are skewed in slightly different directions.
+Both datasets have low representation for Underweight (by BMI) people.
+More analysis can be done in this regard to adjust for the lack of representation.
+The above factors make the data sets not precisely comparable.
 
+A suitable next step for this research paper would be to include age and other demographic details such as lifestyle category.
+Additional information about a person's lifestyle could further improve the diagnosis of any obesity related disease.
+It would also be significant to include whether the specific ethnic information of a person or to use data set from the same geographical location.
+
+Lastly, the models can improve the proxies for `fat_percentage_category` by studying current literature on the relationship between BMI and fat percentage.
+The model bases the categories based on the existances of similar trends without taking into consideration any adjustment factors.
 \newpage
@@
 
-# Appendix {.unnumbered}
+# Appendix {#sec-model-details}
 
-# Additional data details
+## Model Verification
 
-# Models Validation {#sec-model-details}
+Both models were considerd due to their high accuracy which is shown in [@tbl-accuracy].
+Inclusion of weight and fat free mass, respectively, caused a higher accuracy but I suspect it is because of over-fitting of data.
 
-## Posterior predictive check
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-accuracy
+#| tbl-cap: "Accuracy of Models"
+# Predicted probabilities for each category
+accuracy_1 <- mean(cleaned_data_model$predicted_category == cleaned_data_model$fat_percentage_category)
 
-we compare the posterior with the prior. This shows...
+accuracy_2 <- mean(body_mass_model$predicted_category == body_mass_model$fat_percentage_category)
+
+accuracy_table <- data.frame(
+  Model = c("Measurements Model", "Body Mass Model"),
+  Accuracy = c(accuracy_1*100, accuracy_2*100)
+)
+
+# Generate a table in Quarto
+kable(accuracy_table)
+
+```
+
+## Surveys, sampling, and observational data
+
+To establish a more causational relationship between anthropocentric measurements and actual fat mass to indicate obesity, we can set up an interview based study.
+This would include interviewing the test subjects about their lifestyle, their fitness levels, additional hormonal fluctuations such as pregnancy or menopause.
+This would provide us more information to be able to differentiate between the circumference measurement.
+We could also interview the subjects about their ethnicity, age and if they currently have any metabolism enhancing or reducing treatments.
+
+This interview would provide the base for a more detailed comparative analysis of the establishment of bodily measurements and its correlation with body fat.
+The same set of people can be recorded for their anthropocentric measurements as well as body composition measurement.
+It would then make the results of the study more directly application to conclude a relationship.
 

### 2024-12-15 scripts/01-download_data.R

@@ -10,7 +10,7 @@
 library(tidyverse)
 library(arrow)
 
-# Download data
+# Instructions for obtaining the data is included in the READ.ME
 # Model 1 Data: AWS BodyM Training Data
 measurements_data <- read_csv("./data/raw_data/local-train/measurements.csv")
 measurements_subject_data <- read_csv("./data/raw_data/local-train/hwg_metadata.csv")
@@ -19,9 +19,6 @@ measurements_subject_data <- read_csv("./data/raw_data/local-train/hwg_metadata.
 # Data uses ; as delimiter
 body_mass_data <- read_csv2("./data/raw_data/DATASET FOR NEW BODY FAT MEASUREMENT.csv")
 
-# Testing Data: AWS BodyM Test A Data
-test_data <- read_csv("./data/raw_data/local-testA/measurements.csv")
-test_subject_data <- read_csv("./data/raw_data/local-testA/hwg_metadata.csv")
 
 #### Save data ####
 # Model 1
@@ -30,7 +27,3 @@ write_parquet(measurements_subject_data, "data/raw_data/measurements_subject_dat
 
 # Model 2
 write_parquet(body_mass_data, "data/raw_data/body_mass_data.parquet")
-
-# Test
-write_parquet(test_data, "data/raw_data/test_data.parquet")
-write_parquet(test_subject_data, "data/raw_data/test_subject_data.parquet")       
